### 2022-01-20 VarietyInNLS/VarietyInNonlinearLeastSquaresCodes.Rmd
more incremental

---
title: "Variety in the Implementation of Nonlinear Least Squares Program Codes"
author: 
   - John C Nash, retired professor, University of Ottawa
date: "16/02/2021 revised 20/01/2022"
output: 
    pdf_document:
        keep_tex: false
        toc: true
bibliography: ../BibSupport/ImproveNLS.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## require(bookdown) # language engine to display text??
```

# Abstract

There are many ways to structure a Gauss-Newton style nonlinear least squares
program code. In organizing and documenting the nearly half-century of programs
in the Nashlib collection associated with @jncnm79, the author realized that this
variety could be an instructive subject for numerical software designers.

# The Problem to be solved

Given a function that can be written as a sum of squared residuals

$$  f(x) = \sum_{i=1}^n {r_i(x)^2} = r^t r $$

where $x$ is a set of real valued parameters, we wish to find the minimum of $f(x)$
with respect to x. There may be additional conditions that add to the complexity of
the problem such as

- bounds (or box) constraints of the form  $$ l_i <= x_i <= u_i $$ alternatively
  written in vector form $$ l <= x <=u $$

- masks (or fixed parameters), for example, $$ x_i = value_i $$. Such constraints
  reduce the dimensionality of the nonlinear least squares problem, and we could
  specify the problem without the masked parameters. However, it is sometimes very
  convenient to be able to write our problem with such paramters that may later be
  altered in value or allowed to vary in an optimization of the sum of squares.

- inequality constraints, which in general may be written as 
  $$ C_k(x) >=0 $$ which can be written as a vector form $$ C(x) >= 0 $$
  A special and important case is that of **linear inequalities** $$ A x >= b $$
  that is, $$ C(x) = A x - b $$

# Underlying algorithms

## Gauss Newton method

An overview of this method is provided by Wikipedia: 
\url{https://en.wikipedia.org/wiki/Gauss%E2%80%93Newton_algorithm}
The method was described by C.F.Gauss in 1809 in *Theoria motus corporum 
coelestium in sectionibus conicis solem ambientum*.

In calculus we learn that a stationary point (local maximum, minimum
or saddle point) of a function $f$ occurs where its gradient (or first
derivative) is zero. In multiple dimensions, this is the same as
having the gradient $g$ stationary. The so-called **Newton method**
uses the second derivatives in the Hessian matrix $H$ defined as

$$  H_{i,j} = \partial^2{f}/ {\partial{x_i} \partial{x_j}}$$

where $f$ is a function of several variables $x_i$, $i=1,...,n$, written
collectively as $x$. The Newton equations are defined as

$$  H \delta = - g $$

where $H$ and $g$ are evaluated at a guess or estimate of $x$. $x$ is then updated
to

$$ x \leftarrow x + \delta $$

and we iterate until there is no change in $x$, assuming that the method does indeed
converge for the problem at hand. Newton (and his sometimes associated worker Raphson) 
used the algorithm we attribute
to them in a very special case and in a manner unrecognizable as that described
above.  (See \url{https://en.wikipedia.org/wiki/Newton%27s_method}).

For our present situation, the $f(x)$ can be written as a sum of squares as above. It is worth
repeating the vector form of this

$$  f(x) = r^t r $$

because it is less fussy to write and read. This is clear when we expand the evaluation
of the Hessian of this sum of squares function.

$$ H_{j,k} = 2 \sum_{i-1}^n { { [ (\partial r_i/\partial{x_j} ) } ({\partial r_i/\partial{x_k} ) + r_i (\partial^2 r_i/{\partial{x_j} \partial{x_k} ) ]}}} $$

If we define the **Jacobian** matrix

$$ J_{i,j} = \partial r_i/\partial{x_j} $$

then the gradient, in vector form, is

$$ g = 2 J^t r $$

and the first part of the Hessian is

$$  J^t J $$

Arguing that the residuals should be "small", Gauss proposed that the 
Newton equations could be approximated by ignoring
the second term (which have elements of the residual vector times second derivatives 
of the residual).  This gives the **Gauss-Newton**
equations (cancelling a factor 2)

$$ (J^t J)  \delta  =  - J^t r $$

We can use this in an iteration similar to the Newton iteration. Note that these
equations are equivalent to the linear least squares problem that can be written

$$  J \delta  \approx - r $$

Like most other treatments, we will skate over the thin ice that $J^t J$ is a 
good approximation of the Hessian and that the resulting method converges.

## Hartley's method

Conditions for convergence of the Newton iteration and the Gauss-Newton 
iteration are rather a nuisance to verify. In any case we want a solution 
to the original nonlinear least squares problem. 
@Hartley61 proposed that rather than use an iteration

$$ x \leftarrow x + \delta$$

one could do a search along the direction $\delta$. Ideally we would want to **minimize**

$$ f(x + \alpha \delta) $$

with respect to the (scalar) parameter $\alpha$. This one-dimensional minimization can
take quite a lot of computational effort, however, and typically a value of $\alpha$ that
permits a suitable reduction in the sum of squares $f(x)$ is accepted, and the iteration repeated
from

$$ x_{new} = x + \alpha * \delta $$

@Hartley61 recommends a quadratic (parabolic) line search where two "new" points
are evaluated, followed by a third try at the suggested minimum of the parabola fitted
to the existing and two new sum of squares function values. Tactics are needed to deal
with the case that the proposed point is not an improvement in the sum of squares.

A different approach is the R function `nls()`, which calls the C routine `nls.c` to compute 
the Gauss Newton direction <!-- (`incr` -- see nls-flowchart.txt) -->
after which a step-halving "line-search" is used.

In `nls.c` we also see that once a satisfactory "new" set of parameters has been 
obtained, we reset the $\alpha$ value to the minimum of $2 * \alpha$ and 1.0. Oddly, there
is no provision for the user to set the tactical choices of the maximum value, the reduction
or the increase in $\alpha$, despite our experience that other choices can be advantageous
for some problems. Moreover, $x_{new}$ is deemed satisfactory if its **deviance** (essentially
the weighted sum of squares of the residuals) is **no greater than** the current best
value. Our experience is that it may be better to insist on a decrease in the value,
or even to use an "acceptable point" approach that enforces a sufficient decrease using
a measure based on the gradient of the deviance projected on $\delta$.

<!-- ?? Should we explain more. Possibly incorporate into code? -->

Clearly there are many tactical choices that give rise to a variety of particular algorithms.
One concern is that the direction $\delta$ may give no lower value of the sum of squares, since
there is an approximation involved in using $J^t J$ rather than $H$. For example, $\delta$ could
be perpendicular to any downhill direction.

## Marquardt stabilized Gauss-Newton

@Marquardt1963 is perhaps one of the most important developments in nonlinear least squares
apart from the Gauss-Newton method. There are several ways to view the method.

First, considering that the $J^t J$ may be effectively singular in the computational
environment at hand, the Gauss-Newton method may be unable to compute a search
step that reduces the sum of squared residuals. The right hand side of the 
normal equations, that is, $g = - J^t r$ is the gradient for the sum of squares.
Thus a solution of

<!-- Note mathbf use below -->

$$    \mathbf{1}_n \delta = - g $$
is clearly a downhill version of the gradient. And if we solve

$$  \lambda  \mathbf{1}_n \delta =  - g $$

various values of $\lambda$ will produce steps along the **steepest descents**
direction. Solutions of the Levenberg-Marquardt equations

$$ (J^t J + \lambda  \mathbf{1}_n)  \delta  = C \delta =  - J^t r $$

can be thought of as yielding a step $\delta$ that merges the Gauss-Newton and 
steepest-descents direction. This approach was actually first suggested by 
@Levenberg1944, but it is my opinion that Levenberg 
likely never exercised them in practice, while Marquardt actually had a well-used
program. Before computers were commonly
available, many papers were written with computational
ideas that were given at most a cursory trial. (In 1984, I had the pleasure of showing
Marquardt my own BASIC program running on a Radio Shack TRS80 Model 100 and he  
commented on how the computations were converging to a solution as we watched
the iterations proceed.)

Note that we can define a matrix $K$ which is $J$ augmented below by a suitable diagonal
matrix that is the square root of $\lambda$ times the unit matrix $1_n$. Then the
Levenberg-Marquardt equations are equivalent to

$$  K \delta \approx - r_{augmented} $$

where we have added $n$ zeros to $r$ to make $r_{augmented}$. Similar adjustments
are used for scaled Marquardt methods mentioned below.

Practical Marquardt methods require us to specify an initial $\lambda$ and ways
to adjust its value. For example, if $$SS(x) = r^t r = f(x)$$ is the function 
we wish to minimize to "fit" our nonlinear model, one set of choices is:

- set an initial $\lambda$ of 0.0001

- **A** Compute $J$ and $g$,

- **B** set up Marquardt equations for current $\lambda$ and find $\delta$, 
   compute new set of parameters $$x_{new} = x + \delta $$

- if $x_{new} \ne x$, compute $SS(x_{new})$, else quit (**termination**)

- if $SS(x_{new}) < SS(x)$, replace $x$ with $x_{new}$, replace $\lambda$
with $0.4*\lambda$, then retry from **A**

- ELSE increase $\lambda$ to $10*\lambda$ and retry from **B** ($J$ and $g$ do
  NOT need to be recomputed)

In this scheme, apart from the many possibilities that "find $\delta$" allows, 
an algorithm is defined by the initial value, decrease factor 
(e.g., 0.4) and increase factor (e.g., 10) for $\lambda$. However, there are a number of 
further annoying computational 
details concerning underflow or overflow and how we measure equivalence of 
iterates. That is, we need to ensure $\lambda$ is not so "small" that multiplying
by 10 is meaningless, and we need to know what "not equal" means for floating
point vectors in the arithmetic in use.

An example of choices to deal with these details:

- if $\lambda$ is smaller than some pre-set tolerance, set it to the tolerance. 
  (Otherwise do not worry if it is tiny, as long as the sum of squares still gets
   reduced.)

- when comparing $x_{new}$ and $x$, we can use a modest number, e.g, 100.0 which
  we will call $OFFSET$. In the arithmetic of the system at hand, if the floating-point
  number $(x_{new} + OFFSET)$ matches bit for bit $(x + OFFSET)$, then we take the
  unmodified numbers as equal.

Many practitioners dislike using IF statements (comparisons) of floating point
numbers. Tolerances are often suggested to define "small", but this invites users
to make decisions that lead to poor outcomes and which ignore the precision and
other characteristics of the arithmetic at hand.
There are, of course, well-written codes that address the issues of what "small" means,
but such code is generally quite involved. The R function `all.equal()` presents
a well-considered approach. However,
I have never seen a case where the $OFFSET$ approach caused difficulties, and it
is trivial to code. When $x$ is tiny, we compare $OFFSET$ to 
itself and declare equality. When $x$ is very large, $OFFSET$ does not change it and
we compare $x$ with a value that is identical to it.

### Automatic scaling

In @Marquardt1963, it is shown that the use of the diagonal elements of $J^t J$
instead of $\mathbf{1}_n$ in the Marquardt stabilized Gauss-Newton equations is equivalent
to having a similar scale for all the parameters. Call this diagonal matrix $D$.
Generally this is the preferred
approach to Levenberg-Marquardt stabilizations. However, in low-precision 
arithmetic, elements of the diagonal may underflow to zero and cause issues of
singularity in the equations to be solved. A simple workaround (@jn77ima) is to slightly
modify the equations to

$$ (J^t J + \lambda (D + \phi 1_n)  \delta  =  - J^t r $$

where $\phi$ is a modest number such as 1.

<!-- ?? Check the statements about how things are done and redo tests -->
Note that we form $J^t J$ in `nlsr::nlfb()` to provide the $D$ matrix for 
the scaled stabilization, 
which is quite a lot of computational work. The same program, however, 
uses a QR approach to solving the linearized least squares sub-problem.
To avoid the formation of $D$, an experimental package
`nlsralt` was been built to add functions `nlxbx()` and `nlfbx()` that
use ONLY the fixed element (i.e., unit matrix) stabilization. A
preliminary timing (using @microbench13) with the Hobbs unscaled problem 
(@cnm79, Example 12.2) shows
a slight advantage for `nlsralt` in execution timing with an equivalent solution.

## Other approaches

While the above ideas are the dominant theme in nonlinear least squares, there have been many 
other proposed approaches. The following is a non-exhaustive sample.

### Spiral

@Jones70Spiral seems only to be mentioned in a single paper. There is a flowchart
given but I have never been able to find a code (supposedly Fortran V for the Univac
1108.) The motivations are to avoid both matrix equation solutions and one-dimensional
line-search minimization. Jones mentions matrix inversion, but it is not clear whether
he truly forms the inverse. The main idea in the paper (and presumably the code) is
to use a geometric spiral that mimics the Marquadt search trajectory as $\lambda$
is increased. It would be interesting to actually try some of the ideas, which do,
as far as I can determine, have a reasonable basis and could possibly offer some
efficiencies.

### Approaches to Newton's method for special models

The main attraction of the Gauss-Newton method is that the Jacobian only needs first
derivatives, which are generally much, much easier to compute than the full second
derivatives of the Hessian. However, there have been suggested
Newton-like approaches, possibly for special types of models where some short cut
to the Hessian is available. I regard these approaches as part of the
computational infrastructure of the special cases, as it is usually 
difficult to apply them outside of their context.

Two such examples are the R packages `maxLik` (@maxLik2011) and `trust` (@p-trust).

<!-- expand ??, include some BHHH stuff -->

### Hybrid methods

The central idea of Hartley's method is a line search and that of Marquardt a
parametrized stabilization that can be shown to reduce a sum of squares for a large
enough stabilization parameter. There is nothing to prevent use of a line
search additional to the Marquardt stabilization, and there are, of course, 
several reasonable choices for line search. Also we could solve the (possibly
stabilized) Gauss-Newton equations by several quite different numerical linear
algebra techniques, or even approximations. In certain situations, some of these
choices may be important. Unfortunately, the effort to benchmark the possibilities
seems higher than any benefit in replacing quite good existing methods.

# Sources of implementation variety

The sources of variety in implementation of nonlinear least-squares methods include:

- structuring of the algorithm, that is, how we set up and sequence the
  parts of the overall computation

- Linking of data to the problem. In particular, R allows **subsetting**
  of data at quite a high level, but this capability is almost totally
  hidden from the user or programmer, so that the feature could be dangerous to
  use. There appears, moreover, to be no example or test with `nls()` 
  to show how to use this facility, even though `subset` is an argument
  to the `nls()` function. This is discussed further below.

- programming language

- possible operating environment features

- solver for the least squares or linear equations sub-problems

- stucture of storage for the solver, that is, compact or full

- sequential or full creation of the Jacobian and residual, since
  this may be done in parts

- how the Jacobian is computed or approximated

- higher level presentation of the problem to the computer, as
  in R's `nls` versus packages `minpack.lm` and `nlsr`, or as
  a general optimization problem

- incorporation of constraints on the parameters, namely bounds,
  masks or inequalities.

## Algorithm structure

The R function `nls()` and also the package `minpack.lm` both set up the nonlinear
least squares computation for estimating a model specified as a **formula** (an
R object class) by building an `nlsModel` object, which we will label `m`. This 
object is a set of functions that provide the information needed to compute the 
residuals, jacobian, the search direction, and many other objects required for
the nonlinear least squares solution as well as ancillary information used in
the post-solution analysis and reporting.

In the package `nlsr` (@nlsr-manual), the function `nlxb()` calls `model2rjfun()`
to create functions `res()` and `jac()` that are used in `nlfb()` (the function
`nls.lm()` in `minpack.lm` is very similar) to find a
nonlinear least squares solution. However, post-solution information is built
explicitly AFTER a solution is found.

The essential difference is that `nls()` and `minpack.lm` build a toolbox, `m`,
while `nlsr` creates its required objects as they are needed for the Marquardt
computations.

### Issues with the `nls()` structure

A major criticism of the `nls()` approach is that the collection of information needed
at any one time -- the "environment" in which residuals and jacobians are calculated,
as well as the solution methods -- is not sufficiently well-documented. If the software is
to be updated or maintained, the location and purpose of the many elements that
reside in different environments is difficult to determine without effort.
In my opinion, this renders the structure prone to collapse if there are infrastructure
or other changes. Documentation would go a long way to correcting this problem, but
even better would be clearer structuring. Separation of the problem from the method
would also be useful. At the time of writing of this document, the Gauss-Newton 
METHOD is embedded in the functions created by `nlsModel()` for a given PROBLEM.

## Programming language

We have a versions of the Nashlib Algorithm 23 in BASIC, Fortran, Pascal, and R, 
with Python pending. There may be dialects of these programming languages also,
giving rise to other variations.

## Data linkage to software

Quite reasonably, linking particular data to our problem is mediated by software,
and hence is influenced by the programming language. This is particularly true
for R, since R allows us to **subset** data. Essentially this provides an index
to the rows of data, most easily thought of in terms of a data frame structure.

### Data subsets

The `nls()` function has an argument `subset` and a test included in the 
`Croucher-example1.R` file of this project showed this works as expected. 
However, nowhere have we found a published example of the use of the 
`subset` argument with `nls()`. Worse, it clearly must be applied to
many of the functions used by `nls()` but makes no explicit appearance. Thus it
is VERY difficult to ensure subsetting is correctly applied when we wish to modify
the code, for instance, to provide for newer features especially more reliable
solvers.

Note that the `subset` ARGUMENT to `nls()` is not the same as the `subset()` function
that is usable to restrict the scope of an R object.

OUR HYPOTHESIS: `nls()` uses `subset` to automatically adjust the data to which all
the relevant functions are applied. We note that `minpack.lm::nlsLM` also has a `subset` argument,
but package `nlsr` does not at time of writing (2021-7-12), but will use the **weights** mechanism
of the next subsection of this document at some future release.

### Other approaches to subsetting

In order to provide a sane and well-understood mechanism for subsetting, there
are alternatives for implementation.

- **via weights**: We can weight each "row" that we do NOT wish included as 0.
  This does require us to compute the number of observations as the count of
  non-zero weights e.g, `sum(weights > 0). Incidentally, this underlines the 
  need to ensure weights are non-negative. ??now in nlsj, need to add to nlsr code.??

- **via preprocessing**: We could prepare a function to pre-process the problem and
  create a data collection that is appropriately subset. This may offend against
  the spirit of R/S, but it has the merit that we can easily reassure ourselves
  that we have worked with the correct information, correctly used in our methods.
  ?? Should add this as an option.

PROPOSAL: use the "weights" approach if `nlsj` is called with `subset` active, 
else recommend the preprocessing approach. The preprocessor could also do a lot
of checking, which we might leave out of `nlsj()` in some cases. As of 2021-7-12,
the weights mechanism is working in a trial `nlsj()` and should be transferrable
to package `nlsr` fairly easily.

## Operating environment

Generally, for modern computers, the operating system and its localization do not
have much influence on the way in which we set up nonlinear least squares computations,
though the actual performance of the computations may well be influenced by such factors.
I will comment below about some issues where speed and size of data storage may favour
some approaches over others. However, such issues are less prominent today than in
the past.

One aspect of the computational environment that has proved important to me has
been that of low-precision arithmetic. In the 1970s, it was common to have only
the equivalent of 6 or 7 (decimal) digits of precision. In such environments, 
where, moreover, guard digits and underflow and overflow behaviour were highly
variable features, it was quite common that arguments to functions like exp()
and log() were inadmissible for reasons of computability in the local system
rather than underlying mathematics. In such environments, where there were also
storage limitations, it was important to use very robust algorithms implemented
in ways that were simple but "safe". This led to a number of choices in the
development of the codes in @jncnm79. However, it has turned out that these
choices fare rather well even in more capable computational systems.

## Solver for the least squares or linear equations sub-problems

There are many choices for solver of the Gauss-Newton or related equations
in a nonlinear least squares program. We will consider some of the options.
Note that the choice to set up the normal equations will direct part of the
algorithm selection.

### Solution of the linear normal equations

The Gauss-Newton or Marquardt eqations are a set of linear equations. Moreover, 
the coefficient matrix is non-negative definite and symmetric. Thus it permits of
both general and specialized solution methods for linear equations. Furthermore, one can 
often set up the solution to a linear least squares problem without forming a coefficient matrix $C$ 
by using several matrix decomposition methods. Thus there are many possible procedures.

- Gauss elimination with partial or complete pivoting: The choice of partial or
  complete pivoting changes greatly the program code. However, neither of these
  choices uses the underlying structure and non-negative definite nature of 
  the matrix $A = J^t J$ or stabilized versions $C$ thereof.

- Variants of Gauss elimination that build matrix decompositions: We can apply
  elimination to produce an LU decomposition of $C$, and this may offer some 
  computational advantages, though in the present case such approaches do not
  seem to be ideal.

- Gauss-Jordan inversion: Here we compute the inverse of $C$ and then work with
  this inverse. That is, we use

  $$\delta = - C^{-1} g$$

  This uses much more computational effort than simply solving for $\delta$, 
  though there is an interesting, and very tiny, method by @BauerReinsch71.
  The sheer ingenuity of this code was magnetic, but it has proved to be of
  limited value.

- Cholesky Decomposition and back substitution: There are a number of varieties
  of Cholesky decomposition, but they all take advantage of the symmetric and
  non-negative definite structure of $C$. A particularly compact form was used
  in @Nash1987a, though it is debatable whether the saving in storage was not
  heavily outweighed by the complexity of the code and its documentation.

- Eigendecompositions of the sum of squares and cross-products (SSCP) 
  matrix: the eigenvalues of real, symmetric matrices
  such as $C$ can be computed in many ways and the eigendecomposition used to 
  solve the Gauss-Newton equations. The computations are much more costly than
  simple solution of the equations, but the knowledge of the eigenvalues allows
  us to avoid, or at least be aware of, conditioning issues.

### Solution of the least squares sub-problem by matrix decomposition

Here we propose to solve the linear least squares sub-problem 
using the matrix $J$ or an augmented form $K$
directly. There are a number of approaches that can be used, which we
will divide into two classes, QR and singular value decomposition (SVD) methods.

The QR approach aims to decompose the working matrix (call it $K$,
with $m$ rows and $n$ columns)

$$ K = Q R $$

where $Q$ is $m$ by $n$ orthogonal so

$$ Q^t Q = I_n $$

and $R$ is upper triangular and $n$ by $n$. Some variants have $Q$ an $m$ by $m$ orthogonal
matrix and $R$ upper triangular with $m - n$ rows of zeros appended below.

There are many ways of computing such decompositions:

- Householder reflections with or without pivoting options;
- Givens plane rotations, again with pivoting variants;
- Gram-Schmidt orthogonalization of the columns of $K$, either directly (one
  column at a time orhogonalized to all previous columns) or in a modified
  fashion that proceeds row-wise. There are a number of detailed choices.

SVD approaches compute a decompostion

$$ K =  U S V^t$$

where U and V have various orthogonality properties and S is a diagonal matrix
with elements called the singular values (the square roots of the eigenvalues
of $K^t K$ actually). There are many variations depending on whether $U$ is 
a full $m$ by $m$ matrix, limited to $m$ by $n$ or even truncated to $m$ by $r$ where 
$r$ is the rank of $K$.

A detail in choosing our approach is the mechanism of inclusion of the
Marquardt (or Marquardt-Nash) stabilization. It turns out to be remarkably
easy, if a little tedious. We simply augment $J$ by a diagonal matrix that
has elements that the square roots of  $\lambda (D + \phi 1_n)$. 
Then $K^t K$ is the appropriate $C$ for the Marquardt version
of the Gauss-Newton. Clearly it is equivalent to augment $J$ by **both** $\lambda D$
and $\lambda \phi 1_n$.

Unfortunately, if we decide to use the stabilization involving $D$, then 
we must perform at least a partial computation of the SSCP matrix,
which is relatively costly. For this reason, it may be preferred to use
the unscaled stabilization using $1_n$ only. It may then be important to
provide an explicit (and fixed) parameter scaling.

?? manual scaling -- is it part of nlsr??

### Avoiding re-computation of the Jacobian

The "original" Gauss-Newton iteration offers no guidance in how to deal with a
situation where $x_{new}$ has a poorer fit than $x$. @Hartley61 and @Marquardt1963,
along with most other practical methods, need to recompute the (weighted) residuals
and consequent loss function (sum of squares or deviance) for each value of a step-length
or stabilization parameter ($\lambda$) until either a "better" point $x_{new}$ is found
or we terminate the search. However, as we have seen, Hartley does not have to adjust
the search direction.

Thus it could be helpful to separate the computation of residuals from that of the
Jacobian. However, the structure of R is such that it is often much easier to have the Jacobian
as an **attribute** of the residuals object. Indeed, it is the "gradient" attribute for 
historical reasons. This is the setup for both the finite difference approximations to the 
Jacobian in 
`nlspkg::numericDeriv()` which is the Jacobian approximation routine in base R for nls() in file
`./src/library/stats/R/nls.R`. It is also the structure of `rjfun()` generated by `nlsr::model2rjfun()`. 
On the other hand, 
the computation of Jacobian elements can share many calculations with the computation of the residuals. 
Evidence is needed as to which de-duplication efforts give worthwhile savings, and in what
situations, but is likely of importance only to users with extremely time consuming problems.

## Storage of algorithm objects

If the choice of approach to Gauss-Newton or Marquardt is to build the normal
equations and hence the sum of squares and cross products (SSCP) matrix, we
know by construction that matrix is symmetric and also non-negative definite.
(Generally, unless we have a strict linear dependence between the columns of
the Jacobian, we can say "positive definite".)
If we use an SSCP form, we can apply algorithms that specifically take advantage of both these
properties, including programs that store the working matrix in special formats. 
Some of these are Algorithms 7, 8 and 9 of Nashlib (@jncnm79). Algorithms 7 and 8 are the
Cholesky decomposition and back-solution using a vector of length `n*(n+1)/2`
to store just the lower triangle of the SSCP matrix. Algorithm 9 inverts this
matrix *in situ*.

The original @jncnm79 Algorithm 23 (Marquardt nonlinear least squares solution) computes
the SSCP matrix $J^t J$  and solves the Marquardt-Nash augmented normal equations
with the Cholesky approach. This was continued in the Second Edition @nlacatvn1060620 and
in @jnmws87. Even storing the full $A$ matrix uses only $n^2$ numbers.

However, in the now defunct @jnnlmrt12 and successor @nlsr-manual, the choice
has been to use a QR decomposition of an augmented Jacobian. Similarly, `nls()` uses a QR decomposition
of the Jacobian (referred to as the "gradient") without augmentation. Clearly the $K$ matrix is
at least $m$ by $n$, which for most nonlinear modelling problems is much larger than the $C$ matrix.
The particular QR calculations in `nls()` and its packaged version are quite well-hidden in internal
code, complicating comparisons of storage, complexity and performance.

### Storage in Nash/Walker-Smith (1987) BASIC code

In @Nash1987a, each major iteration generates a $A = J^t J$ matrix and the right hand side $-g = J^t r$
at parameters $x$ using functions for the residuals and Jacobian.
The lower triangle of $A$ is stored as a vector of $n (n+1)/2$ elements.
The diagonal elements are then multiplied by $(1 + \lambda)$ and $\lambda * \phi$ is added to each,
then saved in matrix $C$. $\lambda$ is our Marquardt parameter and $\phi$ is the adjustment suggested in
@jn77ima to
overcome potential underflow to zero of the accumulation of the diagonal elements in low-precision
arithmetic.

Solving $C \delta = -g$, we compute the new sum of squares at $x + \delta$. If this value is greater than
or equal to the current best (i.e., lowest) sum of squares, we increase $\lambda$ and generate a new
$C$ matrix. We do not, however, need to build a new SSCP matrix or gradient $g$, just make the diagonal
adjustment to $A$ to get a new $C$.

The code in @Nash1987a is much more complicated than this description, since we should address

- bounds and masks on the parameters

- tests for "non-computability", where we allow the user to provide an explicit flag that the function cannot
  be computed at the trial parameters. This was included to avoid some of the weaknesses in evaluation of 
  special functions, e.g., `log()` for small arguments or `exp()` for large ones, or for problem specific
  situations where some parameter combinations might be considered inadmissible. (Readers may note that
  the codes were written before @IEEE754-1985 was applied to most BASICs. The author was a member of
  that standardization committee.)

- situations where rounding and truncation errors make the Cholesky decomposition return with an indication
  that $C$ is "singular", at least in the particular computational arithmetic.

### Approach in package `nlsr`

In @nlsr-manual, the approach to the augmented Gauss-Newton equations is nominally equivalent to that in
@Nash1987a,
except that we perform the solution using a QR decomposition, in fact a modification of the DQRDC routine from 
@lapack99, which uses Householder transformations.

One of the issues with creating the SSCP matrix $A$ is that we lose information in finite arithmetic 
(See @jncnm79, Chapter 5). A QR approach to the traditional Gauss-Newton codes is to solve

$$  J \delta = -r $$

That is, we assume a generalized inverse of $J$ can be applied to both sides of the Gauss-Newtx

To move to the Marquardt variant of the Gauss-Newton, we need to augment $J$. Let us call this augmented matrix
$K$. We will also need to augment $r$ by an appropriate number of zeros. 
For the traditional Marquardt method, we
need to use the square roots of the diagonal elements of $J^t J$, that is, an $n$ by $n$ 
diagonal matrix $Z1$ where

$$ Z1_{k,k}  = \sqrt(\lambda) \sqrt(J^t J)_{k,k}  $$

and $r$ is extended by $n$ zeros. For the Nash modification, we add a further $n$ zeros to $r$ and augment
the $K$ matrix with another $n$ by $n$ block $Z2$ which is a diagonal matrix where the diagonal elements are

$$ Z2_{k,k}  = \sqrt(\lambda) \sqrt(\phi)  $$

Unfortunately, there is a need in each major iteration to compute at least the diagonal elements of the
SSCP matrix. It may, in fact, be more efficient to omit the $Z1$ augmentation. Package `nlsralt`, as
mentioned, does this with functions `nlxbx()` (for formula-specified models) and `nlfbx()` (for problems
specified with `res()` and `jac()` functions). This gives the following preliminary example using the 
Hobbs weed problem.

<!-- https://bookdown.org/yihui/rmarkdown-cookbook/source-script.html -->
<!-- ```{r, include=FALSE} -->
<!-- source("your-script.R", local = knitr::knit_global()) -->
<!-- # or sys.source("your-script.R", envir = knitr::knit_global()) -->
<!-- ``` -->

??FIXME??

```{r hobbtime1}
source("HobbsTiming.R", echo=TRUE)
```

### Other storage approaches.

?? do we want to report any other approaches here?
In particular, should we look at minpack.lm to see how the Marquardt stabilization 
is applied? This is a very complicated program, since it calls C to call Fortran,
and the Fortran is layered in a way that each part is well-documented but the
general flow and linkage of the parts is non-obvious.

## Sequential or full Jacobian computation

We could compute a row of the Jacobian plus the corresponding residual element
and process this before computing the next row etc. This means the full Jacobian
does not have to be stored. In Nashlib, Algorithms 3 and 4, we used row-wise data 
entry in linear least squares via Givens' triangularization (QR decompostition),
with the possibility of extending the QR to a singular value decomposition.
Forming the SSCP matrix can also be generated row-wise as well. We do not 
anticipate that any of these approaches will offer advantages that are so
great that we would choose them.

## Analytic or approximate Jacobian

`nls()` and `minpack.lm` use finite difference approximations to compute the Jacobian.
In the case of `minpack.lm` it seems that the standard `numericDeriv()` (from the 
set of routines in file `nls.R`) is used initially for the "gradient", but within
the iteration, the Fortran routine lmdif.f computes its own approximations for
the Jacobian, adding another layer of potential confusion.

The `nlsr` package, however, attempts to compute analytic derivatives using symbolic
and analytic derivative possibilities from other parts of R. A great advantage of
these is that they are for a given point, and do not suffer the issue that a step
must be taken, thereby chancing the violation of a constraint or else an inadmissible
set of arguments for special functions. In my experience, the use of analytic derivatives
is not much faster than the finite difference approximations, and generally finite
difference methods iterate as quickly as the "exact" derivative methods to the 
neighbourhood of a solution. It is in allowing good estimates of convergence 
criteria that the analytic derivatives seem to show their advantage i.e., knowing
we are at a solution rather than the speed in getting there.

## Interfacing to problems

R allows the nonlinear least squares problem to be presented via a formula
for the model. This is generally very attractive to users, since they can
specify their problem in ways that often accord with how they think about them.
Many (most?) researchers do not write program code, so preparing residual and
Jacobian functions is foreign to them.

From the point of view of package or system developers, however, specifications as
formulas pose some particular issues:

- translating a formula to a function for its actual evaluation may be tricky,
  since we must essentially parse the formula. R has some tools for this that
  we use.

- even if we can translate a formula to a function, we need to check that we
  have all the data and parameter inputs to enable the evaluation. Moreover,
  we need to check if the inputs are valid or admissible.

- checking for correctness of inputs and translations can be considered to
  take up most of the code in a reliable and robust package.

# Saving storage

The obvious ways to reduce storage are:

- use a row-wise generation of the Jacobian in either a Givens' QR or SSCP approach.
  This saves space for the Jacobian as well as as well as the working matrices of
  the Gauss-Newton or Marquardt iterations;

- if the number of parameters to estimate is large enough, then a normal equations
  approach using a compact storage of the lower triangle of the SSCP matrix. 
  However, the scale of the saving is really very small in comparison to 
  the size of most programs.

# Measuring performance

Do we plan to use only timing? Or size of code and memory used? ??

# Test problems

```{r settest15}
# set parameters
set.seed(123456)
a <- 1
b <- 2
c <- 3
np <- 15
tt <-1:np
yy <- 100*a/(1+10*b*exp(-0.1*c*tt))
plot.new()
plot(tt, yy, type='l')
set.seed(123456)
ev <- runif(np)
ev <- ev - mean(ev)
y1 <- yy + ev
points(tt,y1,type='p', col="green")
y2 <- yy + 5*ev
points(tt,y2,type='p', col="blue")
y3 <- yy + 10*ev
lg3d15 <- data.frame(tt, yy, y1, y2, y3)
points(tt,y3,type='p', col="red")
library(nlsr)
sol0 <- nlxb(yy ~ a0/(1+b0*exp(-c0*tt)), data=lg3d15, start=list(a0=1, b0=1, c0=1))
print(sol0)
sol1 <- nlxb(y1 ~ a1/(1+b1*exp(-c1*tt)), data=lg3d15, start=list(a1=1, b1=1, c1=1))
print(sol1)
sol2 <- nlxb(y2 ~ a2/(2+b2*exp(-c2*tt)), data=lg3d15, start=list(a2=1, b2=1, c2=1))
print(sol2)
sol3 <- nlxb(y3 ~ a3/(3+b3*exp(-c3*tt)), data=lg3d15, start=list(a3=1, b3=1, c3=1))
print(sol3)
```

The following is a larger dataset version of this test.

```{r settest150, eval=FALSE}
np <- 150
tt <- (1:np)/10
yy <- 100*a/(1+10*b*exp(-0.1*c*tt))
set.seed(123456)
ev <- runif(np)
ev <- ev - mean(ev)
y1 <- yy + ev
y2 <- yy + 5*ev
y3 <- yy + 10*ev
lg3d150 <- data.frame(tt, yy, y1, y2, y3)
np <- 1500
tt <- (1:np)/100
yy <- 100*a/(1+10*b*exp(-0.1*c*tt))
set.seed(123456)
ev <- runif(np)
ev <- ev - mean(ev)
y1 <- yy + ev
y2 <- yy + 5*ev
y3 <- yy + 10*ev
lg3d1500 <- data.frame(tt, yy, y1, y2, y3)
f0 <- yy ~ a0/(1+b0*exp(-c0*tt))
f1 <- y1 ~ a1/(1+b1*exp(-c1*tt))
f2 <- y2 ~ a2/(2+b2*exp(-c2*tt))
f3 <- y3 ~ a3/(3+b3*exp(-c3*tt))
```

# Implementation comparisons

Here want to explore the ideas. First we will examine the sub-problem of solving
the Gauss-Newton equations or their Marquardt variants.

## Linear least squares and storage considerations

Without going into too many details, we will present the linear least squares 
problem as

$$  A x \tilde = b  $$

In this case $A$ is an $m$ by $n$ matrix with $m \ge n$ and $b$ a vector of lenght $m$.
We write **residuals** as

$$  r  =  A x - b  $$
or as

$$  r_1  = b - A x  $$

Then we wish to minimize the sum of squares $r^t r$. This problem does not necessarily
have a unique solution, but the **minimal length least squares solution** which is 
the $x$ that has the smallest $x' x$ that also minimizes $r' r$ is unique.

### Example setup and run in lm()

Let us set up a simple problem in R:

```{r ls1, echo=TRUE}
# simple linear least squares examples
v <- 1:6
v2 <- v^2
vx <- v+5
one <- rep(1,6)
Ad <- data.frame(one, v, v2)
A <- as.matrix(Ad)
print(A)
Ax <- as.matrix(data.frame(one, v, vx, v2))
print(Ax)
y <- -3 + v + v2
print(y)
set.seed(12345)
ee <- rnorm(6)
ee <- ee - mean(ee)
ye <- y + 0.5*ee
print(ye)
sol1 <- lm.fit(A, y)
print(sol1)
cat("Residual SS=",as.numeric(crossprod(sol1$residuals)),"\n")
sol1e <- lm.fit(A,ye)
print(sol1e)
crossprod(sol1e$residuals)
sol2<-lm.fit(Ax,y)
# Note the NA in the coefficients -- Ax is effectively singular
print(sol2)
S2 <- sol2$coefficients
J <- which(is.na(S2))
S2[J]<-0
crossprod(S2)
```

The above uses the intermediate code in function `lm.fit()`. This uses a QR solver, but
it is written as a wrapper in C calling a Fortran routine (in the Fortran 77 dialect).

?? Should we put in the structure and where the code is located?

I believe that the structure of lm() predates the availability of the family
of `qr.xxx()` functions for R. These allow us to access the QR approach 
directly.

```{r ls2, echo=TRUE}
x <- qr.solve(A,y)
print(x)
xe<-qr.solve(A, ye)
print(xe)
# But then we get an error when we try to solve the singular system
# This was NOT caught above.
xx <- try(qr.solve(Ax,y))
print(xx)
xxe<-try(qr.solve(Ax, ye))
print(xxe)
```

## Traditional Normal equations approach

The historically traditional method for solving the **linear** least squares problem was
to apply calculus to set the partial derivatives of the sum of squares with respect to
each parameter to zero. This forms the **normal equations**

$$ A^t A x  = A^t b $$

This was attractive to early computational workers, since while $A$ is $m$ by $n$, $A^t A$
is only $n$ by $n$. Unfortunately, this **sum of squares and cross-products** (SSCP) matrix
can make the solution less reliable, and this is discussed with examples in @jncnm79 and
@nlacatvn1060620.

```{r lm5, echo=TRUE}
AtA <- t(A)%*%A
print(AtA)
Aty <- t(A)%*%y
print(t(Aty))
x<-solve(AtA,Aty)
print(t(x))
```

Let us try this with the extended matrix `Ax`

```{r lm6, echo=TRUE}
AxA <- t(Ax)%*%Ax
print(AxA)
Axy <- t(Ax)%*%y
print(t(Axy))
xx<-try(solve(AxA,Axy))
print(t(xx))
```

### Dealing with singularity

The `lm.fit()` function has a parameter `singular.ok` which defaults to TRUE.
By setting this to FALSE, we get.

```{r lm3, echo=TRUE}
sol2b<-try(lm.fit(Ax,y, singular.ok=FALSE))
print(sol2b)
```

We've already seen above that the solution `sol2` has a size of

```{r num1}
eval(as.numeric(crossprod(S2)))
```

The solution `sol2` is, it turns out, not unique, since the variables `v` and
`vx` and `one` are related, namely

```{r num2, echo=TRUE}
print(vx - (v+5*one))
```
Thus we can find a "new" solution as follows and show (essentially) the same 
residuals

```{r num3, echo=TRUE}
res2<-Ax%*%S2-y
print(t(res2))
print(S2)
S2b<-S2+c(10,2,-2,0)
res2b<-Ax%*%S2b-y
print(t(res2b))
cat("Sum of squares of S2b=", as.numeric(crossprod(S2b)),"\n")
```

### Approximate solution -- Ridge regression

An approximate solution via the normal equations can be found by adding a small 
diagonal matrix to the sum of squares and
cross products `AxA`.

```{r lm7, echo=TRUE}
AxA <- AxA + diag(rep(1e-10,4))
print(AxA)
xxx<-try(solve(AxA,Axy))
print(t(xxx))
print(t(Ax %*% xxx - y))
```

We note that this solution is rather different from `S2` or `S2b`.
There is a large (and often misguided and confusing) literature 
about this sort of approach under the title **ridge regression**.
Note that the Marquardt stabilization of the Gauss-Newton equations
uses the same general idea.

### Dealing with singularity in QR solutions

We already saw that we got errors when trying to use `qr.solve()`, but there
are ways to use the QR decomposition that overcome the singularity. Here
are some illustrations. Note that the size of the solution as measured by
the sum of squares of the coefficients (ignoring the NA usng `na.rm=TRUE`)
are different. Moreover, they are different from the minimum length solution
in the next subsection.

```{r qrsing1, echo=TRUE}
## This fails
xx <- try(qr.solve(Ax,y))
print(xx)
## So does this
xxe<-try(qr.solve(Ax, ye))
print(xxe)
## Let us compute a QR decomposition, using LINPACK then LAPACK
qrd1<-qr(Ax, LAPACK=FALSE)
qrd2<-qr(Ax, LAPACK=TRUE)
# and get the solutions
xx1 <- try(qr.coef(qrd1,y))
print(xx1)
xx2 <- try(qr.coef(qrd2,y))
print(xx2)
# and computer the sum of squares of the coefficients (size of solution)
try(sum(xx1^2))
try(sum(xx1^2, na.rm=TRUE))
try(sum(xx2^2))
```
### Minimum length least squares solution

The minimum length least squares solution, which is unique, 
is found using the Moore-Penrose inverse of `Ax`
(the article https://en.wikipedia.org/wiki/Moore%E2%80%93Penrose_inverse gives a quite good
overview) which can be computed from the Singular Value Decomposition. R has a function `svd()`
which is adequate to our needs here.

```{r lm4, echo=TRUE}
Z<-svd(Ax) # get the svd
D1 <- 1/Z$d # invert S, the singular values diagonal matrix
print(D1) 
D1[4]<-0 # to remove linear dependency (small singvals are set to zero)
# D1 is diagonal of S+ now
# minimum length LS solution A+ = V S+ t(U)
minsol2 <- Z$v %*% (diag(D1) %*% (t(Z$u) %*% y))
print(minsol2)
cat("SS of minsol2=",as.numeric(crossprod(minsol2)),"\n")
resminsol<-Ax%*%minsol2-y
print(t(resminsol))
```

### QR decomposition variants

There are a number of ways to compute a QR decomposition.

- Householder transformations have good vector-computation performance and properties.
  They use matrix transformations that can be called **reflections** to be used to
  build the Q matrix as a product of these transformation matrices as the initial
  rectangular matrix $A$ is rendered upper triangular, or $R$

- Givens transformations (@Bindel2002Givens) are **plane rotations** that zero one element at a time of
  the current working matrix that starts out as $A$. Once again, $Q$ is a product of
  the transformations and $R$ the result of zeroing out sub-diagonal elements.

- The Gram-Schmidt method, which has several forms, builds the Q matrix by 
  subtracting a multiple of the first column of $A$ from the second and then 
  normalizing those columns so their sum of squares is 1. The process is then 
  applied to the third column by subtracting multiples of the first two. The $R$
  matrix is formed from the norms and subtraction multipliers in the process. 
  There are row- and column-wise approaches. 
  See https://en.wikipedia.org/wiki/Gram%E2%80%93Schmidt_process

### Using a QR approach

Let us try to form a QR decomposition of $A$, for example with Givens rotations.

$$ A =  Q R$$

where $Q$ is orthogonal (by construction for plane rotations) and $R$ is upper triangular.
We can rewrite our original form of the least squares problem as

$$  Q^t A = Q^t Q R = R \tilde= Q^t b$$

$R$ is an upper triangular matrix $R_n$ stacked on an $m-n$ by $n$ matrix of zeros. But 
$z = Q^t b$ can be thought of as $n$-vector $z_1$ stacked on ($m-n$)-vector $z_2$. 
It can easily be shown (we won't do so here) that a least squares solution is the
rather easily found (by back-substitution) solution of

$$  R_n x = z_1$$

and the minimal sum of squares turns out to be the cross-product $z_2^t z_2$. Sometimes
the elements of $z_2$ are called **uncorrelated residuals**. The solution for $x$ can
actually be formed in the space used to store $z_1$ as a further storage saving,
since back-substitution forms the elements of $x$ in reverse order.

All this is very nice, but how can we use the ideas to both avoid forming the SSCP
matrix and keep our storage requirements low?

Let us think of the row-wise application of the Givens transformations, and use a 
working array that is $n+1$ by $n+1$. (We can actually add more columns if we have
more than one $b$ vector.)

Suppose we put the first $n+1$ rows of a merged $A | b$ working matrix into this
storage and apply the row-wise Givens transformations until we have an $n$ by $n$
upper triangular matrix in the first $n$ rows and columns of our working array.
We further want row $n+1$ to have $n$ zeros (which is possible by simple transformations)
and a single number in the $n+1$, $n+1$ position. This is the first element of $z_2$.
We can write it out to external storage if was want to have it available, or else
we can begin to accumulate the sum of squares.

We then put row $n+2$ of [$A | b$] into the bottom row of our working storage and
eliminate the first $n$ columns of this row with Givens transformations. This
gives us another element of $z_2$. Repeat until all the data has been processed.

We can at this point solve for $x$. Algorithm 4 of @jncnm79, however, applies the one-sided
Jacobi method to get a singular value decomposition of $A$ allowing of a minimal
length least squares solution as well as some useful diagnostic information about
the condition of our problem. This was also published as @LefkovitchNash1976.

### QR approach to the Marquardt stabilization

We can arrange to solve the normal equations approximately as under Ridge Regression
above by adding a unit matrix scaled by e.g., 1e-5, and padding the RHS with zeros.
The solution is "almost but not quite" the same because the formation of the sum of squares and
cross products matrix explicitly in the normal equations can introduce errors that
are partially avoidable in the QR approach below.

```{r lm7a, echo=TRUE}
Dx <- diag(rep(1e-5,4))
print(Dx)
Ap<-rbind(Ax, Dx)
print(Ap)
## we need to pad y with zeros
yp<-c(y, rep(0,4))
xxp<-qr.solve(Ap,yp)
print(xxp)
## Previous solution via normal equations
print(as.vector(xxx))
```

### Choices when solving the normal equations

Whether the normal equations arise from the original Gauss-Newton or Marquardt
approaches to nonlinear least squares, we have several choices of how to solve
them.

Traditional solutions to the solution of linear equations used elimination 
techniques under names such as **Gaussian Elimination**, **Gauss-Doolittle**,
**Crout's Method**, **Gauss-Jordan** and many others. All these can be considered
as LU decompositions. That is, they can be expressed as a transformation of the
matrix $A = J^t J$ or its Marquardt stabilized variant into a product of a lower
triangular and an upper triangular matrix, possibly with permutations of rows and/or
columns. Row permutations of A give

$$ P A = L U$$

The origin of our matrix $A$ means that it is at least non-negative definite. 
This means that we can form a special LU decomposition that does NOT need
permutations and has

$$ A = L L^t$$

which is called the Cholesky decomposition. In practice, we need $A$ computationally
non-singular (not just theoretically so), but the Marquardt stabilization can be used
to ensure this.

If we are very worried about storage space, we can note that $A$ is symmetric, so
that we only need to store the lower (or upper) triangle, including diagonals. 
And some bright programmers realized we could do our storage in a single array,
so that a matrix of order $n$ could be stored in a vector of length $n (n+1)/2$.
@Nash1987a uses this Cholesky decomposition structure in BASIC code.

Similarly, a Gauss Jordan variant could be programmed using such a format for
$A$ and with only a temporary storage of size $n$, matrix $A$ could be overwritten
with its inverse. This is the code of @BauerReinsch71 which is Algorithm 9 of
@jncnm79.

Both these "single vector" storage programs are challenging to check and maintain.
While I embraced them in the age when 16K of storage was the norm for personal
computers, they are simply too finicky to follow easily, and highly prone to
coding and maintenance errors.

# Implementation choices for an R package

??2021 update to nlsr and ideas for the "Improvements to nls()" project.

# References