### 2022-06-14 MachineSummary/MachineSummary.Rmd
Fix boxplots

---
title: "A Machine Profile Summary"
author: 
   - John C. Nash \thanks{ retired professor, Telfer School of Management, University of Ottawa}
   - Arkajyoti Bhattacharjee \thanks{Department of Mathematics and Statistics, Indian Institute of Technology, Kanpur}
date: "14/06/2021 revised 18/01/2022"
bibliography: ../BibSupport/ImproveNLS.bib
output: 
  pdf_document:
    keep_tex: false
link-citations: yes
linkcolor: blue
---

<!-- date: "`r format(Sys.time(), '%d %B, %Y')`" -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Many statistical calculations draw on system resources of memory, storage, arithmetic,
instructional processing and communications. A given computational environment will
have a particular collection of such resources. With workers now commonly working 
separately and sharing their work virtually, we should consider how to characterize
the systems under which calculations are carried out in order that we can diagnose
differences in results or performance of the computing methods. Our objective here
is to consider how a succinct labeling may provide a **machine profile summary**
to assist such analysis and comparisons. To focus our work, we limit our attention
to R as the software environment, but the ideas are more generally applicable.

# Characterizing a particular computing environment for R

In 2021, we undertook an investigation of tools in R for nonlinear least squares
calculations under the Google Summer of Code initiative. In the process, we recognized 
that we would be comparing calculations
performed on **different computing environments** at our disposal. Some of these were
based on different hardware, while others were different combinations of given hardware
with alternative operating software. In one case a machine failed and had to be replaced,
underlining the need to be able to compare results from the defunct system with its
replacements.

A useful context for our ideas is provided by the Wikipedia article on Computer Performance
(https://en.wikipedia.org/wiki/Computer_performance). We seek a way to provide succinct
measures and identifiers of the particular computing environment at hand, leading to a
**profile** for a given computing environment for which we will provide a 
short but meaningful label unique within our project context. This article
considers some possibilities available in the R statistical software and language
that assist this effort, but the ideas are extensible to other programming systems.

# Desired information

There may be multiple terms used to describe some of the information we want, as
well as different interpretations of some of the units of measure. For example,
storage for disks or memory has decimal and binary schemes of units. Thus 
(see https://en.wikipedia.org/wiki/Gigabyte) there is a movement to use, for example, Gigabyte or GB to mean usits of 1000^3 bytes, while the Gigibyte or Gib is 1024^3.

## Temporal information

For tracking work, it is important to have 
information on when the calculations were performed. 
Most computations should have a time/date stamp that is unlikely to be confused with any
other, at least in combination with a machine name and/or other tags. Here is one
suggestion based on the R `Sys.time()` function. Moreover, as we were working on opposite
sides of the globe, it could be important to either specify time zone or to work with a
single "clock" such as UTC. R provides tools to allow this, as in the following example.

```{r tdstamp}
#get a timestamp
tsstr <- format(Sys.time(), "%Y/%m/%d|%H:%M") # tsstr == time stamp string in form YYYYmmddHHMM
cat("Date and Time stamp:", tsstr, "\n")
# and here is a way to get the UTC time
tutc <- as.POSIXlt(Sys.time(), tz = "UTC")
print(tutc)
utctstmp <- format(tutc, "%Y/%m/%d|%H:%M")
print(utctstmp)
```

## Computing software environment

Some of the important information elements concerning the computing environment that 
we need for reporting tests are

- a machine name. While most systems (and Linux in particular) offer to let the user
  provide a machine name, there are generally defaults that many people accept, and
  these are often uninformative and may be non-unique. We note that VirtualBox with
  a Windows 10 guest machine gave the name DESKTOP-HF4CKVA. Where this name was 
  generated we are not sure. Settings / System allows "rename this PC". Clearly 
  we may want to extend the name if particular software is involved, as follows.

- operating system and version. For Linux systems, we note that besides the 
  distribution variant, e.g., Linux Mint 20.3 MATE 64 bit, it is necessary to add 
  the Linux kernel identifier, e.g.,  5.13.0-25-generic, from the `uname` command.

- compiler or interpreter version. For our needs, it is obvious that the R version
  will be important. However, if any code is compiled or linked to libraries, we
  would like to know that. In particular, versions of compilers (e.g., gfortran, 
  gcc) or BLAS or LAPACK libraries used will affect performance. Linux generally
  displays the BLAS and LAPACK **filenames** in `sessionInfo()`, but to get the
  version information, one needs to dig deeper, for example, using operating
  system commands. 
  <!-- ??For the present, we will omit such extra information unless -->
  <!-- we can extract it easily within R. -->

- specific variations, if any, that should be noted. Here we may want to note if
  a machine is running background tasks, or if some special steps have been 
  taken to speed up or slow down the operation e.g., overclocking.

## Hardware information

Some of the factors relating to hardware that could be important are:

- cpu (i.e., processor)
- number of cores
- operating cycle speed (this is sometimes specific to the cpu specified )

?? how do we reconcile?

<!-- From get_sys_details() -->
<!-- $cpu$model_name -->
<!-- [1] "Intel(R) Core(TM) i5-10400 CPU @ 2.90GHz" -->

<!-- From inxi -C -->
<!-- john@M21:~$ inxi -C -->
<!-- CPU:       Topology: 6-Core model: Intel Core i5-10400 bits: 64 type: MT MCP L2 cache: 12.0 MiB  -->
<!--            Speed: 800 MHz min/max: 800/4300 MHz Core speeds (MHz): 1: 800 2: 800 3: 800 4: 800 5: 800 6: 800 7: 800 8: 800  -->
<!--            9: 800 10: 800 11: 800 12: 800  -->

?? also noted that get_sys_details shows 16 cores on my laptop, but inxi shows 8??

- RAM size total
- RAM size available, possibly indicating swap space and usage
- RAM speed (which could be variable) 
  ?? this is, as I understand the output, in inxi (for Linux). Not as far as I know in benchmarkme.

- GPU (i.e, additional processors) information. 
  ?? is there a way to get this? I don't have a GPU on my desktop machine.

## Software information

Clearly it is important to identify which software packages are active, and the
R `sessionInfo()` command provides this. However, there may be additional 
features that should be mentioned if the user has modified the software in any
way, or is using customized packages.

The use of tools to carry out computations in parallel, using multiple cores or
GPUs, is clearly an issue. In our opinion, such possibilities are not well supported
by common R functions, or indeed in other computing languages, though we would very
much like to hear of tools to provide such information.

# R tools for machine information

## sessionInfo()

```{r sessionInfox}
si <- sessionInfo()
si <- as.vector(si)
si
```

## Machine precision and .Machine

Until the IEEE 754 standard (@institute1985ieee) became available in the 1980s, 
computers used multiple forms of floating point arithmetic. With
some ingenuity, it is possible to detect many aspects of the 
floating point properties. In particular, we may be interested in the
**machine precision**, that is, the smallest number `eps` such that 
`1 + eps` is greater than `1`. R performs computations in IEEE "double"
precision, and provides the R variable `.Machine` that has "information on 
the numerical characteristics of the machine R is running on", including
the machine precision as `.Machine$double.eps`.

The information in `.Machine` is documented as based on @Cody1988Alg665, which is 
an extension of Mike Malcolm's ENVRON (@Malcolm72). A port of this to R is
presented below. It computes the 
machine precision, radix and number of radix digits. A more detailed 
tool was W. Kahan's PARANOIA (@Karpinski85). The rise of IEEE arithmetic has
set aside the routine use of such tools, but recent innovations in so-called
"half-precision", of which there is more than one variety, may result in 
renewed interest. See https://en.wikipedia.org/wiki/Half-precision_floating-point_format.

```{r envrontest}
envron <- function(){
D1 <- as.numeric(1) # use variable to avoid constants when double precision invoked
E5 <- 10 # arbitrary scaling for additive equality tests
B9 <- 1E+35 # big number, not necessarily biggest possible
E6 <- 1 # initial value for radix 
E9 <- 1 # initial value for machine precision
D0 <- E6 
repeat{
  E9 <- E9 / 2 # start of loop to decrease estimated machine precision
  D0 <- E6 + E9 # force storage of sum into a floating-point scalar
  if (D0 <= E6) break
} # repeat reduction while (1+E9) > 1
E9 <- E9 * 2 # restore smallest E9 which gives (1 + E9) > 1
repeat {
   E6 <- E6 + 1 # try different radix values
   D0 <- E6 + E9
   if (D0 <= E6) break
} # until a shift is observed
J1 <- 1 # initial count of radix digits in mantissa
E9 <- 1 # use radix for exact machine precision
repeat {  
  E9 <- E9 / E6 # loop while dividing by radix
  J1 <- J1 + 1 # increment counter
  D0 <- D1 + E9 # add tp 1
  if (D0 <= D1) break # test and repeat until equality
}
E9 <- E9 * E6 # recover last value of machine precision
J1 <- J1 - 1 # and adjust the number of digits
mpvals<-list(eps = E9, radix = E6, ndigits = J1)
}
mp <- envron()
E9 <- mp$eps
E6 <- mp$radix
J1 <- mp$ndigits
cat("E9 = ",E9,"  -- the machine precision,  E9=MIN (X  1+X>1)\n")
cat( "E9*1E+16=", E9 * 1E+16, "\n")
cat( "E6 = ", E6,"  -- the radix of arithmetic", "\n")
cat( "J1 = ", J1,"  -- the number of radix digits in mantissa of FP numbers", "\n")
cat( "E6^(-J1+1) * 1E+16=", 1E+16*(E6^(-J1+1)), "\n")
cat(".Machine$double.eps=", .Machine$double.eps, "\n")
cat("From R  .Machine:\n")
cat("double.eps =", .Machine$double.eps, "\n", 
    "double.base =", .Machine$double.base, "\n", 
    "double.digits =", .Machine$double.digits, "\n")
```

## Sys.info()

This R function returns information on the operating system and the generic type of the 
processor.

```{r sysinfox}
Sys.info()
```

## benchmarkme

The CRAN package `benchmarkme` (@benchmarkme) provides a wide range of
information on system in which it is running, indeed a superset of 
previous tools mentioned. It is primarily designed to 
benchmark the capabilities of a particular computing system, so highly appropriate
to our current effort, though its output is not succinct. Note that we found a 
minor bug in the `get_ram()`
function of this package for some Windows 10 versions. Our thanks for 
Colin Gillespie, the author, for making  corrections in version 1.0.8, based on communications with one of us (AB).

?? Need to comment on

?? AB: I found the plots of the benchmarks difficult to read. Which end of the
plots are "better"?

<!-- - uses `proc.time()` for timings -->
<!-- - has single core and parallel timings -->
<!-- - ?? does it do GPUs -->
<!-- - provides far too much for most needs i.e., a succinct summary of -->
<!--   how capable a computing system should be -->

<!--  ```{r benchmarkmex}  -->
<!-- # library(benchmarkmeData) -->
<!-- # library(benchmarkme) -->
<!-- # ls(package:benchmarkmeData) -->
<!-- # ls(package:benchmarkme) -->
<!-- # lsf.str("package:benchmarkme") # list functions in -->
<!-- # lsf.str("package:benchmarkmeData") -->
<!-- ## This next line takes a lot of time to run, so is commented out here -->
<!-- # benchmark_std() -->
<!-- # get_byte_compiler() -->
<!-- # gla <- get_linear_algebra() -->
<!-- # gsd <- get_sys_details() -->
<!--  ``` -->

## Some other performance examples

There are many performance examples used for testing the speed of computers. A 
well-known example is the @linbench03 LINPACK test used for ranking supercomputers.
These generally employ linear algebra calculations. The example below, by
contrast, focuses on the special function computations. Variations on this
simple performance test have been used by one of the authors for nearly half
a century. We used the `microbenchmark` timing package (@microbench13),
which allows us to see the variation in timings, which can be the result of
other activities and processes in our computing systems.

Note that it appears that the first cycle of the loop is very slow. Fortunately 
`microbenchmark` seems aware of this, and we can specify a number of "warmup"
cycles. In the present case, we tried 2 cycles. We also tried byte compiling
the loopesc function with the `compiler` package. Note that rerunning the 
code several times gave slightly different boxplots, but the general characteristics
were preserved.

```{r loopesc, echo=TRUE}
loopesc <- function(nn){
  ss <- 0
  for (i in 1:nn) {
    xx <- exp(sin(cos(1.0*i)))
    ss <- ss + xx
  }
  xx
}
```

```{r looptime, code=readLines("loopbox.R"), echo=FALSE}
```

## Missing tools

While we cannot be certain that they do not exist, there do not appear to be tools
readily available to allow us to characterize the following features of R computations
that might be relevant to understanding machine performance.

- Some users will want to know key information about (pseudo-)random number generation
  used in their calculations.
  We suspect that users will need to custom-code ways to get such information
  about their computations. For example, it is useful to have the specification
  of the generator and the seed(s) used. For parallel computations, it is important
  in some applications to use the same set of numbers for each thread, while in
  others we want to use separate segments of the stream of pseudo-random numbers. 
  From the specification of the generator, it should be possible to know the 
  period (how many numbers are generated before the sequence repeats) and 
  properties concerning correlations between generated numbers. However, in our
  experience, such information requires a lot of research to uncover.

- R allows for byte compilation of code, and there are several ways in which this
  can be done. `benchmarkme` returns some information about the general setting used,
  but details would need user intervention if locally written code is used.

- Compilation of code written in other programming languages and how the resultant
  routines are called will similarly require user attention. Fortunately, some 
  information is available for heavily used software such as the BLAS and LAPACK,
  for which `sessionInfo()` reports the libraries used. There are, however, several
  alternatives, some of which claim high-performance.

<!-- # Tools for accessing and clearing environments and dataframes -->

<!-- ?? should this be here -- probably to list things in the workspace? -->

<!-- ``` -->
<!-- sys.frame() -->
<!-- sys.frames() -->
<!-- sys.status() -->
<!-- sys.on.exit() -->
<!-- sys.parents() -->
<!-- sys.calls() -->
<!-- ``` -->

# A modest profile

For most uses in recording tests of R functions and packages for optimization and nonlinear 
least squares, which has been our focus, it seems that the `benchmarkme` function 
`get_sys_details()` provides more than sufficient information for most of our needs. 
A workable profile for our project is captured in the following script,
which allows a 1-line compact sketch of the system in use.

```{r code=xfun::read_utf8('MachID.R')}
```

<!-- Note that we are reading from the file `MachID.R` so that this file is usable elsewhere -->
<!-- and the current document does not become out of sync with the working MachID. -->

For other applications, it is clear that users will make other choices to meet their particular needs.

Our efforts noted in this article underline the following principles.

- Relatively comprehensive tools such as package `benchmarkme` are needed to 
  make possible machine profiles to allow sensible comparison of computational 
  performance across different platforms.

- If a project uses a common software environment such as R, then the tools for
  capturing profile data should be available from that environment, and run 
  under different operating systems and hardware. In that respect, `benchmarkme`
  has been very helpful to us.

- No tool covers all the possibilities, and users may have to do extra work to 
  capture information they need. Moreover, the challenges of cross-platform 
  computing ensure that some of the results of different tools will be inaccurate.

- Summarizing capability information for computing environments to provide a 
  short-form profile is a challenge that requires care and attention. Indeed,
  it is much easier to gather a large body of information than to devise a 
  compact but useful precis of it.

We welcome conversation on how machine capability may be summarized and reported
to allow sensible comparison of computational performance of algorithms and 
systems.

## References