### 2023-02-03 RefactoringNLS/JNAB230203.Rmd
Add another section Jacobian singular

---
title: "Refactoring `nls()`: a constructive critique"
date: "2023-02-03"
abstract: >
   Based on work for a Google Summer of Code project "Improvements to `nls()`",
   we consider the features and limitations of the R function `nls()`
   in the context of improving and rationalizing R tools for nonlinear regression.
   Important considerations are the usability and  maintainability of
   the code base that provides the functionality `nls()` claims to offer.
   Our work suggests that the
   existing code makes maintenance and improvement very difficult,
   with legacy applications and examples blocking some important updates.
   Discussion of these matters is relevant to improving R generally as
   well as its nonlinear estimation tools.

draft: true
author:  
  # see ?rjournal_article for more information
    - name: John C. Nash
      affiliation: retired professor, University of Ottawa
      address:
      - Telfer School of Management
      - Ottawa ON Canada K1N 6N5
      orcid: 0000-0002-2762-8039
      email: profjcnash@gmail.com
    - name: Arkajyoti Bhattacharjee
      affiliation: Indian Institute of Technology
      address:
      - Department of Mathematics and Statistics
      - Kanpur
      email: arkastat98@gmail.com
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: ../BibSupport/ImproveNLS.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# The `nls()` function: strengths and shortcomings

`nls()` is the tool in base R, the primary software software distribution
from the Comprehensive R Archive Network (*https://cran.r-project.org*),
for estimating nonlinear statistical models. The function dates to the 1980s and the work
related to @bateswatts in S (see *https://en.wikipedia.org/wiki/S_%28programming_language%29*).

The `nls()` function has a remarkable and comprehensive set of capabilities for estimating
nonlinear models that are expressed as **formulas**. In particular, we note that it

- handles formulas that include R functions, even ones which call calculations in
  other programming languages
- allows data to be weighted or subset
- can estimate bound constrained parameters
- provides a mechanism for handling partially linear models
- permits parameters to be indexed over a set of related data
- produces measures of variability (i.e., standard error estimates) for the estimated parameters
- has related profiling capabilities for exploring the likelihood surface as parameters are changed
- links to a number of pre-coded (`selfStart`) models

With such a range of features and long history, the code has become untidy 
and overly patched. It is, to
our mind, essentially unmaintainable. Moreover, its underlying methods can and should be improved, 
as we suggest below.

## Feature: Convergence and termination tests (FIXED)

A previous issue with `nls()` that prevented it from providing parameter estimates for zero-residual
(i.e., perfect fit) data was corrected thanks to suggestions by one of us.

In the manual page for `nls()` in R 4.0.0 there is the warning:

>  **Do not use `nls` on artificial "zero-residual" data.**

>  The `nls` function uses a relative-offset convergence criterion
  that compares the numerical imprecision at the current parameter
  estimates to the residual sum-of-squares.  This performs well on data of
  the form

>  $$  y = f(x, \theta) + eps $$

>  (with `var(eps) > 0`).  It fails to indicate convergence on data of the form

>  $$  y = f(x, \theta)  $$

>  because the criterion amounts to comparing two components of the round-off 
  error.

>  If you wish to test `nls` on artificial data please add a noise component, 
  as shown in the example below.

This amounted to admitting R cannot solve well-posed problems unless data is polluted
with errors.

This issue can be easily resolved. The "termination test" for the **program** rather than for 
"convergence" of the underlying **algorithm** is the Relative Offset Convergence Criterion 
(see @BatesWatts81). This projects the proposed step in the parameter vector on the gradient 
and estimates how much the sum of squares loss function should decrease. This estimate is divided 
by the current size of the loss function to avoid scale issues.
When we have "converged", the estimated
decrease is very small. However, with small residuals,
the sum of squares loss function is (almost) zero and we get the possibility of a 
zero-divide failure.

Adding a small quantity to the loss function before dividing avoids trouble. 
In 2021, one of us (J. Nash) proposed that `nls.control()` have an additional parameter `scaleOffset` 
with a default value of zero. Setting it to a small number -- 1.0 is 
a reasonable choice --
allows small-residual problems (i.e., near-exact fits) to be dealt with easily. We call this the
**safeguarded relative offset convergence criterion**. The default value gives the legacy behaviour.
We are pleased to note that this improvement is now in the R distributed code since version 4.1.0.

### More general termination tests

The convergence criterion above, the principal termination control of `nls()`, leaves out 
some possibilities that could
be useful for some problems. The package `nlsr` (@nlsr2019manual) already offers both 
the safeguarded relative offset test (**roffset**) as well as a **small sum of 
squares** test (**smallsstest**) that compares the latest evaluated sum of squared
(weighted) residuals to `e4` times the initial sum of squares, where
`e4 <- (100*.Machine$double.eps)^4` is approximately 2.43e-55.

We note that `nls()` uses a termination test to stop after `maxiter` "iterations". 
Unfortunately, the meaning of "iteration" varies among programs and requires 
careful examination of the
code. We prefer to use the number of times the residuals or the jacobian have
been computed and put upper limits on these. Our codes exit (terminate) when these limits
are reached. Generally we prefer larger limits than the default `maxiter = 50` of `nls()`,
but that may reflect the more difficult problems we have encountered because 
users consult us when standard tools have given unsatisfactory results.

## Feature: Failure when Jacobian is computationally singular

This refers to the infamous "singular gradient" termination message of `nls()`. 
A Google search of

```
R nls "singular gradient"
```

gets over 4000 hits that are spread over some years. This could be because the 
Jacobian is poorly approximated (see **Jacobian computation** below). However,
it is common in nonlinear least squares computations that the Jacobian is very close
to singular for some values of the model parameters. In such cases, we need to find
an alternative algorithm to the Gauss-Newton iteration of `nls()`. The most common
work-around is the Levenberg-Marquardt stabilization (see @Marquardt1963, @Levenberg1944, 
@jn77ima), and versions of it have been implemented in packages `minpack.lm` and `nlsr`.
In our work, we prepared experimental prototypes of `nls` that incorporate
stabilization, but integration with all the features of `nls()` has proved difficult.

Let us consider a 'singular gradient' example 
*https://stats.stackexchange.com/questions/13053/singular-gradient-error-in-nls-with-correct-starting-values*.
`nlsr::nlxb()` displays the singular values of the Jacobian, which confirm that the `nls()` error message is correct,
but this is not helpful in obtaining a solution.

```{r SESingGrad}
reala <- -3; realb <- 5; realc <- 0.5; realr <- 0.7; realm <- 1 # underlying parameters
x <- 1:11 # x values; 11 data points
y <- reala + realb * realr^(x - realm) + realc * x # linear + exponential function
testdat <- data.frame(x, y) # save in a data frame
strt <- list(a = -3, b = 5, c = 0.5, r = 0.7, m = 1) # give programs a good start
jform <- y ~ a + b * r^(x - m) + c * x # Formula
library(nlsr)
linexp2 <- try(nlxb(jform, data = testdat, start = strt, trace = FALSE))
linexp2 # Note singular values of Jacobian in rightmost column
# We get an error with nls()
linexp <- try(nls(jform, data = testdat, start = strt, trace = FALSE))
```

The Jacobian is essentially singular as shown by its singular
values. Note that these are **displayed** by package `nlsr` in a single column in
the output to provide a compact layout, but the values do **NOT** correspond to 
the individual parameters in 
whose row they appear; they are a property of the whole problem.

## Feature: Jacobian computation

`nls()`, with the `numericDeriv()` function, computes the Jacobian as the "gradient"
attribute of the residual vector. This is implemented as a mix of R and C code, but
we have created a rather more compact version entirely in R. 
See *https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS*.
<!-- Code project. See the companion document **Jacobian Calculations for nls()** -->
<!-- (?? need citation -- currently **DerivsNLS.pdf**). -->

As far as we can understand the logic in `nls()`, the Jacobian computation during
parameter estimation is carried out entirely within the called C-language program,
and the R code function `numericDeriv()`, part of `./src/library/stats/R/nls.R`
in the R distribution source code. This is used to provide Jacobian information in 
the `nlsModel()` and `nlsModel.plinear()` functions, which are **not** exported for 
general use.

The Jacobian used by `nlsr::nlxb()` is, by default, computed from analytic expressions, 
while that
of `nls()` mostly uses a numerical approximation. Some `selfStart` models do provide
"gradient" information. `minpack.lm::nlsLM()` invokes `numericDeriv()` in its local 
version of `nlsModel()`, but it appears to use an internal approximate Jacobian code from 
the original Fortran `minpack` code, namely, `lmdif.f`. Such differences in approach
can lead to different behaviour, though in our experience differences are usually minor.

- A pasture regrowth problem (@Huet2004, page 1, based on @Ratkowsky1983) has a poorly
  conditioned Jacobian and `nls()` fails with "singular gradient". 
  Worse, numerical approximation to the Jacobian give the error
  "singular gradient matrix at initial parameter estimates" for `minpack.lm::nlsLM`
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of `nlsr::nlxb` give a solution.

- Karl Schilling (private communication) provided an example where a model specified
  with the formula `y ~ a * (x ˆ b)` causes `nlsr::nlxb` to fail because the partial
  derivative w.r.t. `b` is `a * (x^b * log(x))`. If there is data for which `x = 0`,
  this is undefined. In such cases, we observed that `nls()` and `minpack.lm::nlsLM`
  found a solution, though it can be debated whether such lucky accidents can be
  taken as an advantage.

Note that the `selfStart` models in the base R `./src/library/stats/R/zzModels.R` file 
and in package `nlraa` (@MiguezNLRAA2021) may
provide the Jacobian in the "gradient" attribute of the "one-sided" formula that defines
each model, and these Jacobians may be the analytic forms. The `nls()` function, after
computing the "right hand side" or `rhs` of the residual, then checks to see if the
"gradient" attribute is defined, and, if not, uses `numericDeriv` to compute a Jacobian
into that attribute. This code is within the `nlsModel()` or `nlsModel.plinear()`
functions. The use of analytic Jacobians
almost certainly contributes to the good performance of `nls()` on `selfStart` models.

# References

?? to sort out
- note that AB now at Ohio State?
- ORCID for AB?