### 2023-03-02 RefactoringNLS/JNAB230203.tex
adding files needed to develop first R-Journal submisison

% !TeX root = RJwrapper.tex
\title{Refactoring \texttt{nls()}: a constructive critique}
\author{by John C. Nash and Arkajyoti Bhattacharjee}

\maketitle

\abstract{%
Based on work for a Google Summer of Code project ``Improvements to \texttt{nls()}'', we consider the features and limitations of the R function \texttt{nls()} in the context of improving and rationalizing R tools for nonlinear regression. Important considerations are the usability and maintainability of the code base that provides the functionality \texttt{nls()} claims to offer. Our work suggests that the existing code makes maintenance and improvement very difficult, with legacy applications and examples blocking some important updates. Discussion of these matters is relevant to improving R generally as well as its nonlinear estimation tools.
}

\hypertarget{the-nls-function-strengths-and-shortcomings}{%
\section{\texorpdfstring{The \texttt{nls()} function: strengths and shortcomings}{The nls() function: strengths and shortcomings}}\label{the-nls-function-strengths-and-shortcomings}}

\texttt{nls()} is the tool in base R, the primary software software distribution
from the Comprehensive R Archive Network (\emph{\url{https://cran.r-project.org}}),
for estimating nonlinear statistical models. The function dates to the 1980s and
the work related to Bates and Watts (1988) in S (see \emph{\url{https://en.wikipedia.org/wiki/S_\%28programming_language\%29}}).

The \texttt{nls()} function has a remarkable and comprehensive set of capabilities for
estimating nonlinear models that are expressed as \textbf{formulas}. In particular, we note that it

\begin{itemize}
\tightlist
\item
  handles formulas that include R functions, even ones which call calculations in
  other programming languages
\item
  allows data to be weighted or subset
\item
  can estimate bound constrained parameters
\item
  provides a mechanism for handling partially linear models
\item
  permits parameters to be indexed over a set of related data
\item
  produces measures of variability (i.e., standard error estimates) for the
  estimated parameters
\item
  has related profiling capabilities for exploring the likelihood surface as
  parameters are changed
\item
  links to a number of pre-coded (\texttt{selfStart}) models
\end{itemize}

With such a range of features and long history, the code has become untidy
and overly patched. It is, to
our mind, essentially unmaintainable. Moreover, its underlying methods can and
should be improved, as we suggest below.

\hypertarget{feature-convergence-and-termination-tests-fixed}{%
\subsection{Feature: Convergence and termination tests (FIXED)}\label{feature-convergence-and-termination-tests-fixed}}

A previous issue with \texttt{nls()} that prevented it from providing parameter
estimates for zero-residual
(i.e., perfect fit) data was corrected using a suggestion by one of us.

In the manual page for \texttt{nls()} in R 4.0.0 there is the warning:

\begin{quote}
\textbf{Do not use \texttt{nls} on artificial ``zero-residual'' data.}
\end{quote}

\begin{quote}
The \texttt{nls} function uses a relative-offset convergence criterion
that compares the numerical imprecision at the current parameter
estimates to the residual sum-of-squares. This performs well on data of
the form
\end{quote}

\begin{quote}
\[  y = f(x, \theta) + eps \]
\end{quote}

\begin{quote}
(with \texttt{var(eps)\ \textgreater{}\ 0}). It fails to indicate convergence on data of the form
\end{quote}

\begin{quote}
\[  y = f(x, \theta)  \]
\end{quote}

\begin{quote}
because the criterion amounts to comparing two components of the round-off
error.
\end{quote}

\begin{quote}
If you wish to test \texttt{nls} on artificial data please add a noise component,
as shown in the example below.
\end{quote}

This amounted to admitting R cannot solve well-posed problems unless data is polluted
with errors.

This issue can be easily resolved. The ``termination test'' for the \textbf{program} rather than for
``convergence'' of the underlying \textbf{algorithm} is the Relative Offset
Convergence Criterion (see Bates, Douglas M. and Watts, Donald G. (1981)). This projects the proposed step
in the parameter vector on the gradient
and estimates how much the sum of squares loss function should decrease.
This estimate is divided
by the current size of the loss function to avoid scale issues.
When we have ``converged'', the estimated
decrease is very small. However, with small residuals,
the sum of squares loss function is (almost) zero and we get the possibility of a
zero-divide failure.

Adding a small quantity to the loss function before dividing avoids trouble.
In 2021, one of us (J. Nash) proposed that \texttt{nls.control()} have an additional
parameter \texttt{scaleOffset}
with a default value of zero. Setting it to a small number -- 1.0 is
a reasonable choice -- allows small-residual problems (i.e., near-exact fits)
to be dealt with easily. We call this the
\textbf{safeguarded relative offset convergence criterion}. The default value
gives the legacy behaviour. We are pleased to note that this improvement is
now in the R distributed code since version 4.1.0.

\hypertarget{more-general-termination-tests}{%
\subsubsection{More general termination tests}\label{more-general-termination-tests}}

The convergence criterion above, the principal termination control of \texttt{nls()},
leaves out some possibilities that could be useful for some problems. The
package \texttt{nlsr} (John C Nash and Duncan Murdoch (2019)) already offers both
the safeguarded relative offset test (\textbf{roffset}) as well as a \textbf{small sum of
squares} test (\textbf{smallsstest}) that compares the latest evaluated sum of squared
(weighted) residuals to \texttt{e4} times the initial sum of squares, where
\texttt{e4\ \textless{}-\ (100*.Machine\$double.eps)\^{}4} is approximately 2.43e-55.

We note that \texttt{nls()} uses a termination test to stop after \texttt{maxiter} ``iterations''.
Unfortunately, the meaning of ``iteration'' varies among programs and requires
careful examination of the
code. We prefer to use the number of times the residuals or the jacobian have
been computed and put upper limits on these. Our codes exit (terminate) when
these limits are reached. Generally we prefer larger limits than the default
\texttt{maxiter\ =\ 50} of \texttt{nls()},
but that may reflect the more difficult problems we have encountered because
users consult us when standard tools have given unsatisfactory results.

\hypertarget{feature-failure-when-jacobian-is-computationally-singular}{%
\subsection{Feature: Failure when Jacobian is computationally singular}\label{feature-failure-when-jacobian-is-computationally-singular}}

This refers to the infamous ``singular gradient'' termination message of \texttt{nls()}.
A Google search of

\begin{verbatim}
    R nls "singular gradient"
\end{verbatim}

\noindent
gets over 4000 hits that are spread over some years. This could be because the
Jacobian is poorly approximated (see \textbf{Jacobian computation} below). However,
it is common in nonlinear least squares computations that the Jacobian is very close
to singular for some values of the model parameters. In such cases, we need to find
an alternative algorithm to the Gauss-Newton iteration of \texttt{nls()}. The most common
work-around is the Levenberg-Marquardt stabilization (see Marquardt (1963),
Levenberg (1944), Nash (1977)), and versions of it have been implemented in
packages \texttt{minpack.lm} and \texttt{nlsr}.
In our work, we prepared experimental prototypes of \texttt{nls} that incorporate
stabilization, but integration with all the features of \texttt{nls()} has proved difficult.

Let us consider a `singular gradient' example
\emph{\url{https://stats.stackexchange.com/questions/13053/singular-gradient-error-in-nls-with-correct-starting-values}}.
\texttt{nlsr::nlxb()} displays the singular values of the Jacobian, which confirm that
the \texttt{nls()} error message is correct, but this is not helpful in obtaining a
solution.

\begin{verbatim}
#> residual sumsquares =  0  on  11 observations
#>     after  1    Jacobian and  1 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> a                     -3            NA         NA         NA           0       26.49  
#> b                      5            NA         NA         NA           0       9.615  
#> c                    0.5            NA         NA         NA           0       2.466  
#> r                    0.7            NA         NA         NA           0      0.1098  
#> m                      1            NA         NA         NA           0   9.798e-17
\end{verbatim}

\begin{verbatim}
#> Error in nlsModel(formula, mf, start, wts, scaleOffset = scOff, nDcentral = nDcntr) : 
#>   singular gradient matrix at initial parameter estimates
\end{verbatim}

The Jacobian is essentially singular as shown by its singular
values. Note that these are \textbf{displayed} by package \texttt{nlsr} in a single column in
the output to provide a compact layout, but the values do \textbf{NOT} correspond to
the individual parameters in
whose row they appear; they are a property of the whole problem.

\hypertarget{feature-jacobian-computation}{%
\subsection{Feature: Jacobian computation}\label{feature-jacobian-computation}}

\texttt{nls()}, with the \texttt{numericDeriv()} function, computes the Jacobian as the ``gradient''
attribute of the residual vector. This is implemented as a mix of R and C code, but
we have created a rather more compact version entirely in R.
See \emph{\url{https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS}}.

As far as we can understand the logic in \texttt{nls()}, the Jacobian computation during
parameter estimation is carried out entirely within the called C-language program,
and the R code function \texttt{numericDeriv()}, part of \texttt{./src/library/stats/R/nls.R}
in the R distribution source code. This is used to provide Jacobian information in
the \texttt{nlsModel()} and \texttt{nlsModel.plinear()} functions, which are \textbf{not} exported for
general use.

The Jacobian used by \texttt{nlsr::nlxb()} is, by default, computed from analytic
expressions, while that of \texttt{nls()} mostly uses a numerical approximation. Some
\texttt{selfStart} models do provide ``gradient'' information. \texttt{minpack.lm::nlsLM()}
invokes \texttt{numericDeriv()} in its local version of \texttt{nlsModel()}, but it appears
to use an internal approximate Jacobian code from the original Fortran
\texttt{minpack} code, namely, \texttt{lmdif.f}. Such differences in approach can lead to
different behaviour, though in our experience differences are usually minor.

\begin{itemize}
\item
  A pasture regrowth problem (Huet et al. (2004), page 1, based on Ratkowsky (1983)) has a poorly
  conditioned Jacobian and \texttt{nls()} fails with ``singular gradient''.
  Worse, numerical approximation to the Jacobian give the error
  ``singular gradient matrix at initial parameter estimates'' for \texttt{minpack.lm::nlsLM}
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of \texttt{nlsr::nlxb} give a solution.
\item
  Karl Schilling (private communication) provided an example where a model specified
  with the formula \texttt{y\ \textasciitilde{}\ a\ *\ (x\ ˆ\ b)} causes \texttt{nlsr::nlxb} to fail because the partial
  derivative w.r.t. \texttt{b} is \texttt{a\ *\ (x\^{}b\ *\ log(x))}. If there is data for which \texttt{x\ =\ 0},
  this is undefined. In such cases, we observed that \texttt{nls()} and \texttt{minpack.lm::nlsLM}
  found a solution, though it can be debated whether such lucky accidents can be
  taken as an advantage.
\end{itemize}

Note that the \texttt{selfStart} models in the base R \texttt{./src/library/stats/R/zzModels.R}
file and in package \texttt{nlraa} (Miguez (2021)) may
provide the Jacobian in the ``gradient'' attribute of the ``one-sided'' formula that
defines each model, and these Jacobians may be the analytic forms. The \texttt{nls()}
function, after computing the ``right hand side'' or \texttt{rhs} of the residual, then
checks to see if the ``gradient'' attribute is defined, and, if not, uses
\texttt{numericDeriv} to compute a Jacobian
into that attribute. This code is within the \texttt{nlsModel()} or \texttt{nlsModel.plinear()}
functions. The use of analytic Jacobians almost certainly contributes to the good
performance of \texttt{nls()} on \texttt{selfStart} models.

\hypertarget{feature-weights-on-observations}{%
\subsection{Feature: Weights on observations}\label{feature-weights-on-observations}}

\texttt{nls()}, \texttt{nlsr::nlxb()} and \texttt{minpack.lm::nlsLM()} all have a \texttt{weights} argument that
specifies a vector of weights the same length as the number of residuals. Each
residual is multiplied by the square root of the corresponding weight. The values
returned by the \texttt{residuals()} function are weighted, and the \texttt{fitted()} or
\texttt{predict()} function are used to compute raw residuals.

\hypertarget{feature-subsetting}{%
\subsection{Feature: Subsetting}\label{feature-subsetting}}

\texttt{nls()} accepts an argument \texttt{subset}. This acts through the mediation of
\texttt{model.frame} and is not clearly obvious in the source code files
\texttt{/src/library/stats/R/nls.R} and \texttt{/src/library/stats/src/nls.C}.

While the implementation of subset at the level of the call to \texttt{nls()} has a certain
attractiveness, it does mean that the programmer of the solver needs to be aware of
the source (and value) of objects such as the data, residuals and Jacobian. This
is overly complicated. By preference,
we would implement subsetting by means of zero-value weights, with observation counts
(and degrees of freedom) computed via the numbers of non-zero weights. Alternatively,
we would extract a working dataframe from the relevant elements in the original.

\hypertarget{feature-na.action}{%
\subsection{Feature: na.action}\label{feature-na.action}}

\texttt{na.action} is an argument to the \texttt{nls()} function, but it does not appear obviously
in the source code, often being handled behind the scenes after referencing the
option \texttt{na.action}. A useful, but possibly dated, description is given in:
\emph{\url{https://stats.idre.ucla.edu/r/faq/how-does-r-handle-missing-values/}}.

The typical default action, which can be seen by using the command
\texttt{getOption("na.action")}
is \texttt{na.omit}. This option essentially omits from computations any observations
containing missing values (i.e.~any row of a data frame containing an NA).
\texttt{na.exclude} does much of the same for computations, but keeps the rows with NA
elements so that predictions are in the correct row position. We recommend that
workers actually test output to verify the behaviour is as wanted.

A succinct description of the issue is given in: \emph{\url{https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r}}
where the ``Answer'' states

\begin{quote}
The only benefit of \texttt{na.exclude} over \texttt{na.omit} is that the former will retain the
original number of rows in the data. This may be useful where you need to retain the
original size of the dataset - for example it is useful when you want to compare
predicted values to original values. With \texttt{na.omit} you will end up with fewer
rows so you won't as easily be able to compare.
\end{quote}

\texttt{na.pass} simply passes on data ``as is'', while \texttt{na.fail} will essentially stop
if any missing values are present.

Our concern with \texttt{na.action} is that users may be unaware of the effects of an option
setting they may not even be aware has been set. Should \texttt{na.fail} be the default?

\hypertarget{feature-model-frame}{%
\subsection{Feature: model frame}\label{feature-model-frame}}

\texttt{model} is an argument to the \texttt{nls()} function, which is documented as:

\begin{quote}
\textbf{model} logical. If true, the model frame is returned as part of the object. Default is FALSE.
\end{quote}

Indeed, the argument only gets used when \texttt{nls()} is about to return its result
object, and the element \texttt{model} is NULL unless the calling argument \texttt{model} is
TRUE. (Using the same name could be confusing.) However, the model frame is
used within the function code in the form of the object \texttt{mf}. We feel that
users could benefit from more extensive documentation and examples of its use.

\hypertarget{feature-sources-of-data}{%
\subsection{Feature: Sources of data}\label{feature-sources-of-data}}

\texttt{nls()} can be called without specifying the \texttt{data} argument. In this case, it will
search in the available environments (i.e., workspaces) for suitable data objects.
We do NOT like this approach, but it is ``the R way''. R allows users to leave many
objects in the default (\texttt{.GlobalEnv}) workspace. Moreover, users have to actively
suppress saving this workspace (\texttt{.RData}) on exit; otherwise, any such file in the
path, when R is launched, will be loaded. The overwhelming proportion of R users
in our acquaintance avoid saving the workspace because of the danger of lurking
data and functions which may cause unwanted results.

Nevertheless, to provide compatible behaviour with \texttt{nls()}, competing programs
need to ensure equivalent behaviour, but users should test that the
operation is as they intend.

\hypertarget{feature-missing-start-vector-and-self-starting-models}{%
\subsection{Feature: missing start vector and self-starting models}\label{feature-missing-start-vector-and-self-starting-models}}

Nonlinear estimation algorithms are almost all iterative and need a set of starting
parameters. \texttt{nls()} offers a special class of modeling functions called
\textbf{\texttt{selfStart}} models. There are a number of these in base R
(\texttt{./src/library/stats/R/zzModels.R})
and others in R packages such as CRAN package \texttt{nlraa} (Miguez (2021)), as well as
the now-archived package \texttt{NRAIA}. Unfortunately, some \textbf{\texttt{selfStart}} codes
entangle the calculation of starting values for parameters
with the particulars of the \texttt{nls()} code. Though there is a \texttt{getInitial()} function,
this is not easy to use to simply compute the initial parameter estimates outside of
\texttt{nls()}, in part
because it may call that function. Such circular references are, in our view,
dangerous. Moreover, we believe that it would be helpful to have \texttt{selfStart}
models that allow users to explicitly provide values other than those suggested
for the starting parameters.

Other concerns are revealed by the example below. Here, the \texttt{SSlogis}
\texttt{selfStart} function is used to generate a set
of initial parameters for a 3-parameter logistic curve. The form used by \texttt{SSlogis}
is \(y \,\sim\, Asym\,/\, (1 \,+\, exp((xmid\,-\,tt)\,/\,scal))\), but we show
how the starting parameters for this model can be transformed to those of
another form of the model, namely, \(y \,\sim\, b1\,/\,(1 \,+\, b2\,*\,exp(-b3\,*\,t))\).

The code for \texttt{SSlogis()} is in \texttt{./src/library/stats/R/zzModels.R}.
This R function includes analytic expressions for the Jacobian (``gradient'').
These could be useful to R users, especially if documented. Moreover, we
wonder why the programmers have chosen to save so many quantities in ``hidden''
variables, i.e., with names preceded by ``.''. These are then not displayed by the
\texttt{ls()} command, making them difficult to access.

In the event that a \texttt{selfStart} model is not available, \texttt{nls()} sets all the
starting parameters to 1. This is, in our view, tolerable, but could be improved
by using a set of values that are all slightly different, which, in the case of
the example model \(y \,\sim\, a \,*\, exp(-b \,*\, x) + c\,*\,exp(-d \,*\, x)\)
would avoid a singular Jacobian because \(b\) and \(d\) were equal in value.
A sequence like 1.0, 1.1, 1.2, 1.3 for the four parameters
could be provided quite simply instead of all 1's.

\hypertarget{strategic-issues-in-selfstart-models}{%
\subsubsection{\texorpdfstring{Strategic issues in \texttt{selfStart} models}{Strategic issues in selfStart models}}\label{strategic-issues-in-selfstart-models}}

By providing starting values that are likely to be reasonable and by including
Jacobian code in the \texttt{selfStart} model code, some of the deficiencies of the
Gauss-Newton algorithm are mitigated. However, creating such functions is a lot
of work, and their
documentation (file \texttt{./src/library/stats/man/selfStart.Rd}) is quite complicated.
We believe that the focus should be placed on getting good initial parameters,
that is \texttt{getInitial()} function, though avoiding the current calls back to \texttt{nls()}.
Interactive tools, such as ``visual fitting'' (Nash and Velleman (1996)) might be
worth considering.

We also note that the introduction of \texttt{scaleOffset} in R 4.1.1 to deal with the
convergence test for small residual problems now requires that the \texttt{getInitial()}
function have dot-arguments (\texttt{...}) in its argument list. This illustrates the
entanglement of many features in \texttt{nls()}.

\hypertarget{issue-returned-output-of-nls-and-its-documentation}{%
\subsection{Issue: returned output of nls() and its documentation}\label{issue-returned-output-of-nls-and-its-documentation}}

The output of \texttt{nls()} is an object of class ``nls'' which has the following structure:

\begin{quote}
A list of
\end{quote}

\begin{quote}
\texttt{m} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~~ an \texttt{nlsModel} object incorporating the model.
\end{quote}

\begin{quote}
\texttt{data} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~the expression that was passed to \texttt{nls} as the data argument. The actual \newline
\(~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\) data values are present in the environment of the \texttt{m} components, e.g., \newline \(~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\) \texttt{environment(m\$conv)}.
\end{quote}

\begin{quote}
\texttt{call} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ the matched call with several components, notably \texttt{algorithm}.
\end{quote}

\begin{quote}
\texttt{na.action} ~ ~ ~ ~ ~ ~ ~ ~ ~ the \texttt{"na.action"} attribute (if any) of the model frame.
\end{quote}

\begin{quote}
\texttt{dataClasses} ~ ~ ~ ~ ~ ~ ~ \(~\)the \texttt{"dataClasses"} attribute (if any) of the ``\texttt{terms}'' attribute of the \newline
\(~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\) model frame.
\end{quote}

\begin{quote}
\texttt{model} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~if \texttt{model\ =\ TRUE}, the model frame.
\end{quote}

\begin{quote}
\texttt{weights} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ if \texttt{weights} is supplied, the weights.
\end{quote}

\begin{quote}
\texttt{convInfo} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~~ a list with convergence information.
\end{quote}

\begin{quote}
\texttt{control} ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ the control \texttt{list} used, see the \texttt{control} argument.
\end{quote}

\begin{quote}
\texttt{convergence,\ message} ~for an \texttt{algorithm\ =\ "port"} fit only, a convergence code (\texttt{0} for \newline
\(~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\)convergence) and message.
\end{quote}

\begin{quote}
~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ To use these is \emph{deprecated}, as they are available from \texttt{convInfo} now.
\end{quote}

The \texttt{nls} object contains some elements that are awkward to produce by other
algorithms, but some information that would be useful is not presented in a
clear manner. Moreover, the complexity of the object is a challenge to users.

In the following, we use \texttt{result} as the returned object from \texttt{nls()}.

The \texttt{data} return element is an R symbol. To actually access the data from this
element, we need to use the syntax:

\begin{verbatim}
eval(parse(text=result$data))
\end{verbatim}

However, if the call is made with \texttt{model=TRUE}, then there is a returned element
\texttt{model} which contains the data, and we can list its contents using:

\begin{verbatim}
ls(result$model)
\end{verbatim}

and if there is an element called \texttt{xdata}, it can be accessed as \texttt{result\$model\$xdata}.

Let us compare the \texttt{nls()} result with that from
\texttt{nlsr::nlxb()}, which ostensibly solves the same problem:

\begin{quote}
\texttt{coefficients} A named vector giving the parameter values at the supposed solution.
\end{quote}

\begin{quote}
\texttt{ssquares} \(~~~~~\) The sum of squared residuals at this set of parameters.
\end{quote}

\begin{quote}
\texttt{resid} \(~~~~~~~~~~\) The residual vector at the returned parameters.
\end{quote}

\begin{quote}
\texttt{jacobian} \(~~~~~\) The jacobian matrix (partial derivatives of residuals w.r.t. the
parameters) at the \newline
\(~~~~~~~~~~~~~~~~~~~~\) returned parameters.
\end{quote}

\begin{quote}
\texttt{feval} \(~~~~~~~~~~\) The number of residual evaluations (sum of squares computations) used.
\end{quote}

\begin{quote}
\texttt{jeval} \(~~~~~~~~~~\) The number of Jacobian evaluations used.
\end{quote}

However, actually looking at the structure of a returned result gives a list of 11
items. The extra 5 are:

\begin{verbatim}
 $ lower       : num [1:3] -Inf -Inf -Inf
 $ upper       : num [1:3] Inf Inf Inf
 $ maskidx     : int(0) 
 $ weights     : NULL
 $ formula     :Class 'formula'  language y ~ Asym/(1 + exp((xmid - tt)/scal))
  .. ..- attr(*, ".Environment")=<environment: R_GlobalEnv> 
 - attr(*, "class")= chr "nlsr"
\end{verbatim}

The result object from \texttt{nlsr::nlxb()} is still much smaller than the one \texttt{nls()}
returns. Moreover, \texttt{nlxb}
explicitly returns the sum of squares as well as the residual vector and Jacobian.
The counts of evaluations are also returned.
(Note that the singular values of the Jacobian are
actually computed in the \texttt{print} and \texttt{summary} methods for the result.)
As we made our comparisons, we noted several potential updates to the \texttt{nlsr}
documentation as well as that for \texttt{nls()}.

\hypertarget{weights-in-returned-functions-from-nls}{%
\subsubsection{Weights in returned functions from nls()}\label{weights-in-returned-functions-from-nls}}

As already noted, the function \texttt{resid()} (an alias for \texttt{residuals()}) is
WEIGHTED, as are those in the structured object \texttt{m} returned by \texttt{nls()} or
\texttt{minpack.lm::nlsLM}, e.g., \texttt{ans\$m\$resid()}. The function \texttt{nlsModel()}, which
we have had to extract from the base R code and explicitly \texttt{source()} because
it is not exported to the working namespace, allows us to compute residuals
for particular coefficient sets.

\begin{verbatim}
weed <- c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443,
          38.558, 50.156, 62.948, 75.995, 91.972)
tt <- 1:12
weeddf <- data.frame(tt, weed)
wts <- 0.5^tt # simple weights
frmlogis <- weed ~ Asym/(1 + exp((xmid - tt)/scal))
Asym<-1; xmid<-1; scal<-1
nowt<-nls(weed ~ SSlogis(tt, Asym, xmid, scal)) # UNWEIGHTED
nowt
\end{verbatim}

\begin{verbatim}
#> Nonlinear regression model
#>   model: weed ~ SSlogis(tt, Asym, xmid, scal)
#>    data: parent.frame()
#>   Asym   xmid   scal 
#> 196.19  12.42   3.19 
#>  residual sum-of-squares: 2.59
#> 
#> Number of iterations to convergence: 0 
#> Achieved convergence tolerance: 1.12e-06
\end{verbatim}

\begin{verbatim}
nowt$m$resid() # This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.
\end{verbatim}

\begin{verbatim}
#>  [1] -0.011900  0.032755 -0.092030 -0.208782 -0.392634  0.057594  1.105728
#>  [8] -0.715786  0.107647  0.348396 -0.652592  0.287569
#> attr(,"gradient")
#>           Asym     xmid    scal
#>  [1,] 0.027117  -1.6229  5.8103
#>  [2,] 0.036737  -2.1769  7.1111
#>  [3,] 0.049596  -2.8997  8.5628
#>  [4,] 0.066645  -3.8266 10.1000
#>  [5,] 0.089005  -4.9881 11.6015
#>  [6,] 0.117921  -6.3988 12.8762
#>  [7,] 0.154635  -8.0418 13.6607
#>  [8,] 0.200186  -9.8498 13.6432
#>  [9,] 0.255106 -11.6901 12.5267
#> [10,] 0.319083 -13.3660 10.1313
#> [11,] 0.390688 -14.6444  6.5083
#> [12,] 0.467334 -15.3139  2.0038
\end{verbatim}

\begin{verbatim}
usewt <- nls(weed ~ SSlogis(tt, Asym, xmid, scal), weights=wts)
usewt
\end{verbatim}

\begin{verbatim}
#> Nonlinear regression model
#>   model: weed ~ SSlogis(tt, Asym, xmid, scal)
#>    data: parent.frame()
#>   Asym   xmid   scal 
#> 199.86  12.50   3.19 
#>  weighted residual sum-of-squares: 0.0196
#> 
#> Number of iterations to convergence: 2 
#> Achieved convergence tolerance: 8.29e-06
\end{verbatim}

\begin{verbatim}
usewt$m$resid() # WEIGHTED. Does NOT take coefficients.
\end{verbatim}

\begin{verbatim}
#>  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
#>  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
#> attr(,"gradient")
#>           Asym     xmid    scal
#>  [1,] 0.026498  -1.6157  5.8228
#>  [2,] 0.035901  -2.1679  7.1333
#>  [3,] 0.048474  -2.8890  8.6006
#>  [4,] 0.065153  -3.8149 10.1617
#>  [5,] 0.087046  -4.9775 11.6983
#>  [6,] 0.115387  -6.3933 13.0222
#>  [7,] 0.151425  -8.0483 13.8709
#>  [8,] 0.196222  -9.8787 13.9297
#>  [9,] 0.250362 -11.7553 12.8918
#> [10,] 0.313611 -13.4827 10.5608
#> [11,] 0.384641 -14.8251  6.9663
#> [12,] 0.460954 -15.5631  2.4357
\end{verbatim}

\begin{verbatim}
source("nlsModel.R")
nmod0 <- nlsModel(frmlogis, data=weeddf, start=c(Asym=1, xmid=1, scal=1), wts=wts)
nmod0$resid() # Parameters are supplied in nlsModel() `start` above.
\end{verbatim}

\begin{verbatim}
#>  [1] 3.3998 3.2545 3.0961 2.9784 2.8438 2.7748 2.6910 2.3474 2.1724 1.9359
#> [11] 1.6572 1.4214
\end{verbatim}

\begin{verbatim}
nmod <- nlsModel(frmlogis, data=weeddf, start=coef(usewt), wts=wts)
nmod$resid()
\end{verbatim}

\begin{verbatim}
#>  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
#>  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}

\hypertarget{interim-output-from-the-port-algorithm}{%
\subsubsection{Interim output from the ``port'' algorithm}\label{interim-output-from-the-port-algorithm}}

As the \texttt{nls()} \textbf{man} page states, when the ``port'' algorithm is used with the
\texttt{trace} argument TRUE, the iterations display the objective function value which
is 1/2 the sum of squares (or deviance). It is likely that the trace display is
embedded in the Fortran of the \texttt{nlminb} routine that is called to execute the
``port'' algorithm, but the factor of 2 discrepancy is nonetheless
unfortunate for users.

\hypertarget{failure-to-return-best-result-achieved}{%
\subsubsection{Failure to return best result achieved}\label{failure-to-return-best-result-achieved}}

If \texttt{nls()} reaches a point where it cannot continue but has not found a point
where the relative offset convergence criterion is met, it may simply exit,
especially if a ``singular gradient'' (singular Jacobian) is found. However, this
may occur AFTER the function has made considerable progress in reducing the sum
of squared residuals.

Here is an abbreviated example:

\begin{verbatim}
#> 61216.    (3.56e+03): par = (-2 0.25 150 50)
#> 2.1757    (2.23e+01): par = (-1.9991 0.31711 2.6182 -1.3668)
#> 1.6211    (7.14e+00): par = (-1.9605 -2.6203 2.5753 -0.55599)
#> Error in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE) : 
#>   singular gradient
\end{verbatim}

\begin{verbatim}
#> [1] "Error in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE) : \n  singular gradient\n"
#> attr(,"class")
#> [1] "try-error"
#> attr(,"condition")
#> <simpleError in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE): singular gradient>
\end{verbatim}

Note that the sum of squares has been reduced from 61216 to 1.6211, but
unless \texttt{trace} is invoked, the user will not get any information about this.
This almost trivial change to the \texttt{nls()} function could be
useful to R users.

\hypertarget{feature-partially-linear-models-and-their-specification}{%
\subsection{Feature: partially linear models and their specification}\label{feature-partially-linear-models-and-their-specification}}

The variable projection method (Golub and Pereyra (1973), O'Leary and Rust (2013)) is generally much more
effective than general approaches in finding good solutions to nonlinear least
squares problems when some of the parameters appear linearly. However, setting
up the calculations, that is, identifying which parameters are linear, is not a
trivial task.

Moreover, specifying a model to a solver should, ideally, use the same syntax across
solver tools. Unfortunately, R allows multiple approaches within different
modeling tools, and within \texttt{nls()} itself. This is highlighted by the case of
partially linear models.

The nonlinear modeling specifications are, of course, a development of linear ones.
Unfortunately, the explicit model \texttt{y\ \textasciitilde{}\ a\ *\ x\ +\ b} does not work with the linear
modeling function \texttt{lm()}, which requires this model to be specified as \texttt{y\ \textasciitilde{}\ x}.

Within \texttt{nls()}, consider the following FOUR different specifications for the same
problem, plus an intuitive choice, labelled \texttt{fm2a}, that does NOT work. In this
failed attempt, putting the \texttt{Asym} parameter in the model causes the \texttt{plinear}
algorithm to try to add another term to the model. We believe this is
unfortunate, and would like to see a consistent syntax. At the time of writing,
we do not foresee a resolution for this issue. In the example, we have NOT
evaluated the commands to save space.

\begin{verbatim}
DNase1 <- subset(DNase, Run == 1) # select the data
## using a selfStart model - do not specify the starting parameters
fm1 <- nls(density ~ SSlogis(log(conc), Asym, xmid, scal), DNase1)
summary(fm1)

## using conditional linearity - leave out the Asym parameter
fm2 <- nls(density ~ 1 / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(xmid = 0, scal = 1),
                 algorithm = "plinear")
summary(fm2)

## without conditional linearity
fm3 <- nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1,
                 start = list(Asym = 3, xmid = 0, scal = 1))
summary(fm3)

## using Port's nl2sol algorithm
fm4 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(Asym = 3, xmid = 0, scal = 1),
                 algorithm = "port"))
summary(fm4)

## using conditional linearity AND Asym does NOT work
fm2a <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)), 
                 data = DNase1, start = list(Asym=3, xmid = 0, scal = 1),
                 algorithm = "plinear", trace = TRUE))
summary(fm2a)
\end{verbatim}

\hypertarget{feature-indexed-parameters}{%
\subsection{Feature: indexed parameters}\label{feature-indexed-parameters}}

Indexed parameters -- those whose names have subscripts on some element that
is part of the coefficient list --
are a useful idea to simplify model specification in some cases. However, they
are also a focus of trouble.

The \textbf{man} file for \texttt{nls()} includes the following example of a situation in which
parameters are indexed. It also uses the ``plinear'' option as an added complication.

Here we use a truncated version of the example to save display space.

\begin{verbatim}
#> > with(mm, table(Strip))
#> Strip
#> S01 S02 S03 
#>   4   4   4 
#> > nl <- nlevels(mm$Strip)
#> > musc.1 <- nls(Length ~ cbind(1, exp(-Conc/th)), mm, start = list(th = 1), 
#> +     algorithm = "plinear")
#> > summary(musc.1)
#> 
#> Formula: Length ~ cbind(1, exp(-Conc/th))
#> 
#> Parameters:
#>       Estimate Std. Error t value Pr(>|t|)    
#> th       0.624      0.222    2.81   0.0203 *  
#> .lin1   25.684      1.441   17.82  2.5e-08 ***
#> .lin2  -26.631      6.147   -4.33   0.0019 ** 
#> ---
#> Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
#> 
#> Residual standard error: 2.96 on 9 degrees of freedom
#> 
#> Number of iterations to convergence: 5 
#> Achieved convergence tolerance: 4.71e-07
#> 
#> > b <- coef(musc.1)
#> > musc.2 <- nls(Length ~ a[Strip] + b[Strip] * exp(-Conc/th), data = mm, 
#> +     start = list(a = rep(b[2], nl), b = rep(b[3], nl), th = b[1]))
#> > summary(musc.2)
#> 
#> Formula: Length ~ a[Strip] + b[Strip] * exp(-Conc/th)
#> 
#> Parameters:
#>    Estimate Std. Error t value Pr(>|t|)    
#> a1  22.9277     0.5154   44.49  1.1e-07 ***
#> a2  27.8606     0.5150   54.09  4.1e-08 ***
#> a3  28.3426     1.0771   26.31  1.5e-06 ***
#> b1 -44.1558    12.3900   -3.56    0.016 *  
#> b2 -43.8084    12.3148   -3.56    0.016 *  
#> b3 -32.8421     2.1553  -15.24  2.2e-05 ***
#> th   0.5548     0.0808    6.86    0.001 ** 
#> ---
#> Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
#> 
#> Residual standard error: 0.798 on 5 degrees of freedom
#> 
#> Number of iterations to convergence: 6 
#> Achieved convergence tolerance: 2.03e-06
\end{verbatim}

Running the example reveals that the answers for the parameters are NOT indexed.
That is, we do not see \texttt{a{[}1{]},\ a{[}2{]},\ a{[}3{]}} but \texttt{a1,\ a2,\ a3}.
This is no doubt because programming for indexed parameters is extremely challenging.

\hypertarget{feature-bounds-on-parameters}{%
\subsection{Feature: bounds on parameters}\label{feature-bounds-on-parameters}}

There are many situations where the context of a problem constrains the values of
parameters. For example, one of us was asked many years ago to estimate a model
where one parameter estimate was negative. The client complained ``But that is
supposed to be the number of grain elevators in Saskatchewan''. The number
should have been a positive integer, and likely less than a few thousand.

\texttt{nls()} can impose bounds on parameters, but only if the \texttt{port} algorithm is used.
Unfortunately, the manual states

\begin{quote}
The \texttt{algorithm\ =\ "port"} code appears unfinished, and does not even check that the starting
value is within the bounds. Use with caution, especially where bounds are supplied.
\end{quote}

This is clearly unsatisfactory for general use, but does find bounded solutions.
Inclusion of bounds on parameters in nonlinear least squares computations is
relatively straightforward, and it is part of the default method
for package \texttt{nlsr}. Package \texttt{minpack.lm} also includes provision for bounds, but
the following script
shows that it does not find a good solution for the scaled Hobbs test problem.

\begin{verbatim}
#> Using nlsr::nlxb():
\end{verbatim}

\begin{verbatim}
#> residual sumsquares =  9.4726  on  12 observations
#>     after  12    Jacobian and  12 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> b1                     2U           NA         NA         NA           0       23.35  
#> b2               4.43325            NA         NA         NA   -2.16e-07           0  
#> b3                     3U           NA         NA         NA           0           0
\end{verbatim}

\begin{verbatim}
#> using minpack.lm::nlsLM():
\end{verbatim}

\begin{verbatim}
#> Nonlinear regression model
#>   model: y ~ 100 * b1/(1 + 10 * b2 * exp(-0.1 * b3 * tt))
#>    data: wf
#> b1 b2 b3 
#>  2  6  3 
#>  residual sum-of-squares: 881
#> 
#> Number of iterations to convergence: 2 
#> Achieved convergence tolerance: 1.49e-08
\end{verbatim}

\begin{verbatim}
#> Using nls with 'port':
\end{verbatim}

\begin{verbatim}
#> Nonlinear regression model
#>   model: y ~ 100 * b1/(1 + 10 * b2 * exp(-0.1 * b3 * tt))
#>    data: wf
#>   b1   b2   b3 
#> 2.00 4.43 3.00 
#>  residual sum-of-squares: 9.47
#> 
#> Algorithm "port", convergence message: both X-convergence and relative convergence (5)
\end{verbatim}

\hypertarget{philosophical-considerations}{%
\subsubsection{Philosophical considerations}\label{philosophical-considerations}}

Bounds on parameters raise some interesting and difficult questions about how
uncertainty in parameter estimates should be computed or reported. That is, the
traditional ``standard errors'' are generally taken to imply symmetric intervals
about the point estimate in which the parameter may be expected to be found with
some probability under certain assumptions.
Bounds change those assumptions. Hence, \texttt{nlsr::nlxb()} does not compute standard
errors nor their derived statistics when bounds are active.

\hypertarget{issue-code-structure-and-documentation-for-maintenance}{%
\subsection{Issue: code structure and documentation for maintenance}\label{issue-code-structure-and-documentation-for-maintenance}}

The \texttt{nls()} code is structured in a way that inhibits both maintenance and
improvement. In particular, the iterative setup is such that introduction
of Marquardt stabilization is not easily available. To obtain performance
when the code was written, a large portion was written in C with consequent
calls and returns that severely complicate the code. Over time, R has become
much more efficient on modern computers, and the need to use compiled C and
Fortran is less critical. Moreover, the burden for maintenance
could be reduced by moving code entirely to R.

While R and its packages are generally very well documented for usage, that
documentation rarely extends as far as it might for code maintenance.
The paucity of such documentation is exacerbated by the mixed R/C/Fortran code base,
where seemingly trivial differences in things like representation of numbers or
base index for arrays lead to traps for unwary programmers.

\hypertarget{goals-and-progress-of-our-effort-to-improve-nls}{%
\section{\texorpdfstring{Goals and progress of our effort to improve \texttt{nls()}}{Goals and progress of our effort to improve nls()}}\label{goals-and-progress-of-our-effort-to-improve-nls}}

\hypertarget{code-rationalization-and-documentation}{%
\subsection{Code rationalization and documentation}\label{code-rationalization-and-documentation}}

When we started \emph{Improvements to nls()} for the 2021 Google Summer of Code we wanted:

\begin{itemize}
\item
  to provide a packaged version of \texttt{nls()} (call it \texttt{nlsalt}) coded entirely in R
  that matches the features in base R or what is packaged in \texttt{nlspkg} as described
  in Nash and Bhattacharjee (2022). The R version can either serve as the operational code or
  as a reference version to facilitate the maintenance of C, Fortran, or other
  versions of the software.
\item
  to streamline the overall \texttt{nls()} infrastructure. By this we
  mean a re-factoring of the routines so they are better suited to maintenance of both
  the existing \texttt{nls()} methods and features as well as new features we or others
  would like to add.
\item
  to explain what we do, either in comments, examples, or separate maintainer
  documentation. Since we are complaining about the lack of explanatory material
  for the current code, we feel it is incumbent on us to provide such material
  for our own work, and if possible for the existing code.
\end{itemize}

These goals echo themes in the recent discussion by Vidoni (2021) and accompanying
commentary.

\hypertarget{rationalization-of-formula-specifications}{%
\subsection{Rationalization of formula specifications}\label{rationalization-of-formula-specifications}}

We have seen that
\texttt{nls()} uses a different formula specification from the default if the
\texttt{plinear} algorithm is used. This is unfortunate, since the user cannot then simply
add \texttt{algorithm="plinear"} to the standard call, making errors more likely.

Nonlinear models, apart from those using the ``plinear'' algorithm, need all the
parameters specified.
That is, the model formula should be written down as it would be computed.
Thus a linear model (which is valid input for nonlinear estimation) should
be specified \texttt{y\ \textasciitilde{}\ a\ *\ x\ +\ b} rather than \texttt{y\ \textasciitilde{}\ x}.
Moreover, we want to avoid the partially linear parameters appearing
in the result object as \texttt{.lin1}, \texttt{.lin2}, etc., where the actual position
of parameters in the model is not explicit. We are unclear why these parameters
begin with a dot, that is, that they have the ``hidden'' indicator.

A possible workaround to allow consistency would be to specify the
partially linear parameters when the \texttt{plinear} option is specified. For example,
we could use \texttt{algorithm="plinear(Asym)"} while keeping the general model formula
from the default specification. This would allow
for the output of different \texttt{algorithm} options to be consistent. However, we have
not yet tried to code such a change, and welcome discussion from those working with
other packages using model formulas.

A further complication of model specification in R is that some formulas require
the user to surround a term with \texttt{I()} to inhibit interpretation of arithmetic
operators. We would prefer that formulas be usable without this complication.

\hypertarget{rationalization-of-indexed-models}{%
\subsection{Rationalization of indexed models}\label{rationalization-of-indexed-models}}

Indexed models clearly have a place in some areas of research.
We need to be able to process models such as
\texttt{Length\ \textasciitilde{}\ a{[}Strip{]}\ +\ b{[}Strip{]}\ *\ exp(-Conc\ /\ th)}, where \texttt{Strip} is the index. Data
\texttt{Length} and \texttt{Conc} are available for all observations and parameter \texttt{th} applies
to every fitted point, but \texttt{a{[}Strip{]}} and \texttt{b{[}Strip{]}} apply only to data indexed, which
could be an experimental run number, or some other categorization.

Given that \texttt{{[}\ {]}} provides the R method to index arrays, the presence of square
brackets in a model
formula signals indexing. The question is how to perform the correct calculations.

Our view is that inclusion of the indexing \textbf{within} \texttt{nls()} or similar tools
introduces complexity that could be avoided by a suitable wrapper function. The
modeling formula must be expanded to encompass all the index cases, likely with
indexed parameters renamed e.g., \texttt{a{[}2{]}} becomes \texttt{a2} as at present with \texttt{nls()}.
However, a wrapper could process output to reconstitute the indexing of parameters.

\hypertarget{streamlining-code}{%
\subsection{Streamlining code}\label{streamlining-code}}

We are pessimistic that the overall structure of \texttt{nls()} can be streamlined due
to the entanglement of so many features with the complicated mix of R, C and
Fortran. Indeed, despite digging into the code over some months, we do not feel
confident that we sufficiently understand it nor that we could maintain it. On
the other hand, we do believe equivalent functionality can be built, but with
at least some differences in how the features will be accessed.

\hypertarget{tests-and-use-case-examples}{%
\subsection{Tests and use-case examples}\label{tests-and-use-case-examples}}

Maintainers of packages need suitable tests and use-case examples in order:

\begin{itemize}
\tightlist
\item
  to ensure packages work properly, in particular, giving
  results comparable to or better than the functions they are to replace;
\item
  to test individual solver functions to ensure they work across the range of
  calling mechanisms, that is, different ways of supplying inputs to the solver(s);
\item
  to pose ``silly'' inputs to see if these bad inputs are caught by the programs.
\end{itemize}

Such goals align with the aims of \textbf{unit testing}
(e.g., \emph{\url{https://towardsdatascience.com/unit-testing-in-r-68ab9cc8d211}}, Wickham (2011), Wickham et al. (2021)
and the conventional R package testing tools).

\hypertarget{a-testing-package-nlscompare}{%
\subsection{A testing package: NLSCompare}\label{a-testing-package-nlscompare}}

One of us has developed a working prototype package at
\emph{\url{https://github.com/ArkaB-DS/nlsCompare}} .
A primary design objective of this is to allow the summarisation of multiple tests in
a compact output. The prototype has a vignette to illustrate its use.

\hypertarget{documentation-and-resources}{%
\subsection{Documentation and resources}\label{documentation-and-resources}}

In our investigation, we built several resources, which are now part of
the repository \emph{\url{https://github.com/nashjc/RNonlinearLS/}}. Particular items
are:

\begin{itemize}
\item
  A BibTex bibliography for use with all documents in this project, but which
  has wider application to nonlinear least squares projects in general
  (\emph{\url{https://github.com/nashjc/RNonlinearLS/blob/main/BibSupport/ImproveNLS.bib}}).
\item
  \texttt{MachID.R} offers a suggested concise summary function
  to identify a particular computational system used for tests. A discussion
  of how it was built and the resources needed across platforms is given in
  at \emph{\url{https://github.com/nashjc/RNonlinearLS/tree/main/MachineSummary}}.
\item
  Nash and Bhattacharjee (2022) is an explanation of the construction of the \texttt{nlspkg} from the
  \texttt{nls()} code in R-base.
\item
  As the 2021 Summer of Code period was ending, one of us (JN) was invited to prepare
  a review of optimization in R. Ideas from the present work have been instrumental
  in the creation of Nash (2022).
\end{itemize}

\hypertarget{strategic-choices-in-nonlinear-model-estimation}{%
\section{Strategic choices in nonlinear model estimation}\label{strategic-choices-in-nonlinear-model-estimation}}

It is possible that \texttt{nls()} will remain more or less as it has been for the past
two decades. Given its importance to R, the focus of discussion should be the
measures needed to secure its continued operation for legacy purposes and how
that may be approached.
We welcome an opportunity to participate in such conversations.

To advance the stability and maintainability of R, we believe the program objects (R
functions) that are created by tools such as \texttt{nls()} should have minimal
cross-linkages and side-effects.

The aspects of \texttt{nls()} that have given us the most trouble:

\begin{itemize}
\item
  The functions that compute the residuals and jacobians often presume
  the data and parameters needed are available in a particular environment. As long
  as the correct environment is used, this provides a compact syntax to invoke the
  calculations. The danger is that the wrong data is accessed if the internal
  search finds a valid name that is not the object we want.
\item
  Weights, subsets, and various contextual controls (such as that for the
  \texttt{na.action} option) are similarly taken from the first available source,
  making for a very simple invocation of calculations, but the context is
  hidden from the user. Furthermore, it can be difficult for those of us
  trying to maintain or improve the code to be certain we have the
  context correct.
\item
  The mixing of R, C and Fortran code was necessary for computational performance
  in the past. Lacking decent developer documentation, programmers now face a
  lot of work to fix or improve code. We believe in having at least a working
  reference version of code in a single programming language. If necessary, by
  measuring (``profiling'') the code, we can find bottlenecks and substitute for
  those slower parts of the reference code.
\item
  A design that isolates the setup, solution, and post-solution parts of
  for complicated calculations reduces the number of objects that must be kept in
  alignment. On the other hand, it may mean users have to add more commands to
  complete their computations.
\item
  All iterative codes should return the best solution they have found so far,
  even if there are untoward conditions reached, such as a singular Jacobian.
  This modification could be made to \texttt{nls()} in the short term, correcting the
  issue identified above under ``Failure to return best result achieved''.
\item
  Given the existence of examples of good practices such as analytic derivatives,
  stabilized solution of Gauss-Newton equations and bounds-constrained parameters,
  base R tools should be moving to incorporate them.
\end{itemize}

We anticipate that tools like \texttt{nlsr} will become more attractive as users discover
the increased robustness and general capabilities, along with the more recent
documentation. From a design point of view, a key difference in approach between
\texttt{nls()} and \texttt{nlsr::nlxb()} is that \texttt{nls()}
builds a large infrastructure, especially the \texttt{nlsModel()} function, from
which the Gauss-Newton
iteration can be executed and other statistical information such as profiles can be
computed, while \texttt{nlxb()} returns quite limited information, and
computes what is needed on an ``as and when'' basis. This follows a path that one
of us (JN) established almost 50 years ago with the software that became Nash (1979),
separating setup, solver, and post-solution analysis phases of computations.
There is, however, the wrapper \texttt{wrapnlsr()} that solves the nonlinear estimation problem with the solver of \texttt{nlsr} then calls \texttt{nls()} to build the object of
class \texttt{nls} for which profile log-likelihood information can be computed.

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

Hans Werner Borchers has been helpful in developing the proposal for this project
and in comments on this and related work. Heather Turner co-mentored the project
and helped guide the progress of the work. Exchanges with Fernando Miguez helped to
clarify aspects of \texttt{selfStart} models and improve package \texttt{nlsr} in that respect.
Colin Gillespie (package \texttt{benchmarkme}) has been helpful in guiding our attempts
to succinctly summarize computing environments.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{references-1}{%
\section*{References}\label{references-1}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-bateswatts}{}}%
Bates, D. M., and D. G. Watts. 1988. \emph{Nonlinear Regression Analysis and Its Applications}. Wiley.

\leavevmode\vadjust pre{\hypertarget{ref-BatesWatts81}{}}%
Bates, Douglas M., and Watts, Donald G. 1981. {``A Relative Offset Orthogonality Convergence Criterion for Nonlinear Least Squares.''} \emph{Technometrics} 23 (2): 179--83.

\leavevmode\vadjust pre{\hypertarget{ref-Golub1973}{}}%
Golub, G. H., and V. Pereyra. 1973. {``The Differentiation of Pseudo-Inverses and Nonlinear Least Squares Problems Whose Variables Separate.''} \emph{SIAM Journal of Numerical Analysis} 10 (2): 413--32.

\leavevmode\vadjust pre{\hypertarget{ref-Huet2004}{}}%
Huet, S., A. Bouvier, M.-A. Poursat, and E. Jolivet. 2004. \emph{Statistical Tools for Nonlinear Regression: A Practical Guide with {S-PLUS} Examples, 2nd Edition}. Berlin \& New York: Springer-Verlag.

\leavevmode\vadjust pre{\hypertarget{ref-nlsr2019manual}{}}%
John C Nash, and Duncan Murdoch. 2019. \emph{{nlsr: Functions for Nonlinear Least Squares Solutions}}.

\leavevmode\vadjust pre{\hypertarget{ref-Levenberg1944}{}}%
Levenberg, Kenneth. 1944. {``A Method for the Solution of Certain Non-Linear Problems in Least Squares.''} \emph{Quarterly of Applied Mathematics} 2: 164-\/-168.

\leavevmode\vadjust pre{\hypertarget{ref-Marquardt1963}{}}%
Marquardt, Donald W. 1963. {``{An Algorithm for Least-Squares Estimation of Nonlinear Parameters}.''} \emph{SIAM Journal on Applied Mathematics} 11 (2): 431--41.

\leavevmode\vadjust pre{\hypertarget{ref-MiguezNLRAA2021}{}}%
Miguez, Fernando. 2021. \emph{{nlraa: Nonlinear Regression for Agricultural Applications}}. \url{https://CRAN.R-project.org/package=nlraa}.

\leavevmode\vadjust pre{\hypertarget{ref-jn77ima}{}}%
Nash, John C. 1977. {``Minimizing a Nonlinear Sum of Squares Function on a Small Computer.''} \emph{Journal of the Institute for Mathematics and Its Applications} 19: 231--37.

\leavevmode\vadjust pre{\hypertarget{ref-cnm79}{}}%
---------. 1979. \emph{Compact Numerical Methods for Computers: Linear Algebra and Function Minimisation}. Bristol: Adam Hilger.

\leavevmode\vadjust pre{\hypertarget{ref-NashWires22}{}}%
---------. 2022. {``{Function minimization and nonlinear least squares in R}.''} \emph{WIREs Computational Statistics}, no. e1580. https://doi.org/\url{https://doi.org/10.1002/wics.1580}.

\leavevmode\vadjust pre{\hypertarget{ref-PkgFromRbase22}{}}%
Nash, John C., and Arkajyoti Bhattacharjee. 2022. {``Making a Package from Base {R} Files.''} \emph{R-Bloggers}, January. \url{https://www.r-bloggers.com/2022/01/making-a-package-from-base-r-files/}.

\leavevmode\vadjust pre{\hypertarget{ref-nash1996nonlinear}{}}%
Nash, John C., and Paul Velleman. 1996. {``Nonlinear Estimation Combining Visual Fitting with Optimization Methods.''} In \emph{Proceedings of the Section on Physical and Engineering Sciences of the American Statistical Association}, 256--61. American Statistical Association.

\leavevmode\vadjust pre{\hypertarget{ref-OlearyRust13}{}}%
O'Leary, Dianne P., and Bert W. Rust. 2013. {``Variable Projection for Nonlinear Least Squares Problems.''} \emph{Computational Optimization and Applications} 54 (3): 579--93.

\leavevmode\vadjust pre{\hypertarget{ref-Ratkowsky1983}{}}%
Ratkowsky, David A. 1983. \emph{Nonlinear Regression Modeling: A Unified Practical Approach}. New York; Basel: Marcel Dekker Inc.

\leavevmode\vadjust pre{\hypertarget{ref-Vidoni21}{}}%
Vidoni, Melina. 2021. {``{Software Engineering and R Programming: A Call for Research}.''} \emph{{The R Journal}} 13 (2): 6--14. \url{https://doi.org/10.32614/RJ-2021-108}.

\leavevmode\vadjust pre{\hypertarget{ref-HWtestthat11}{}}%
Wickham, Hadley. 2011. {``{testthat: Get Started with Testing}.''} \emph{The R Journal} 3: 5--10. \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-HWdevtools21}{}}%
Wickham, Hadley, Jim Hester, Winston Chang, and Jennifer Bryan. 2021. \emph{{devtools: Tools to Make Developing R Packages Easier}}. \url{https://CRAN.R-project.org/package=devtools}.

\end{CSLReferences}

\bibliography{../BibSupport/ImproveNLS.bib}

\address{%
John C. Nash\\
retired professor, University of Ottawa\\%
Telfer School of Management\\ Ottawa ON Canada K1N 6N5\\
%
%
\textit{ORCiD: \href{https://orcid.org/0000-0002-2762-8039}{0000-0002-2762-8039}}\\%
\href{mailto:profjcnash@gmail.com}{\nolinkurl{profjcnash@gmail.com}}%
}

\address{%
Arkajyoti Bhattacharjee\\
Indian Institute of Technology\\%
Department of Mathematics and Statistics\\ Kanpur\\
%
%
%
\href{mailto:arkastat98@gmail.com}{\nolinkurl{arkastat98@gmail.com}}%
}