### 2023-05-14 RefactoringNLS/ComparingNLStools.Rmd

---
title: "A Comparison of R Tools for Nonlinear Least Squares Modeling"
author:
- John C. Nash \thanks{ retired professor, Telfer School of Management, University
  of Ottawa}
- Arkajyoti Bhattacharjee \thanks{Department of Mathematics and Statistics, Indian
  Institute of Technology, Kanpur}
date: "2023-3-2"
output:
  pdf_document:
    keep_tex: yes
    toc: no
  word_document:
    toc: no
bibliography: ./ImproveNLS.bib
link-citations: yes
linkcolor: red
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## require(bookdown) # language engine to display text - does not seem necessary
```

## Abstract

Our Google Summer of Code project "Improvements to `nls()`" considered the features and limitations of the R function `nls()` with the aim 
of improving and rationalizing R tools for nonlinear regression.
The rich features of `nls()` are weakened by several deficiencies and
inconsistencies such as a lack of stabilization of the Gauss-Newton solver.
Further considerations are the usability and  maintainability of
the code base that provides the functionality `nls()` claims to offer.
Various packages, including our `nlsr`, provide alternative capabilities.
We consider the differences in goals, approaches, and features of
different tools for nonlinear least squares modeling in R. Discussion of 
these matters is relevant to improving R generally as well as its 
nonlinear estimation tools.

## The `nls()` function

`nls()` is the tool for estimating nonlinear statistical models in base R, the 
primary software distribution
from the Comprehensive R Archive Network (*https://cran.r-project.org*).
The function dates to the 1980s and the work related to @bateswatts in S (see 
*https://en.wikipedia.org/wiki/S_%28programming_language%29*).

The `nls()` function has a remarkable and comprehensive set of capabilities for 
estimating nonlinear models that are expressed as **formulas**. In particular, we
note that it

- handles formulas that include R functions, even ones that call calculations in
  other programming languages
- allows data to be weighted or subset
- can estimate bound-constrained parameters
- provides a mechanism for handling partially linear models
- permits parameters to be indexed over a set of related data
- produces measures of variability (i.e., standard error estimates) for the 
estimated parameters
- has related profiling capabilities for exploring the likelihood surface as 
parameters are changed
- links to many pre-coded (`selfStart`) models that do not require
  initial parameter values.

With such a range of features and long history, the code has become untidy 
and overly patched, difficult to maintain, and its underlying methods could
be improved. Various workers have developed packages to overcome these 
concerns, and we will address some of these here.

## Scope of our comparison

The tools we will consider are the base-R `nls()` function 
and some packages that are available on the CRAN repository.
Of these, we will pay particular attention to **nlsr** (@nlsr2023manual),
**minpack.lm** (@minpacklm12), and **gslnls** (@gslnls23) which are
general nonlinear least-squares solvers.

While we will provide capsule comments for some other CRAN packages, 
we will not pursue those in the Bioconductor (@Bioconductor) 
collection, nor those on repositories such as 
GitHub (https://github.com) and Gitlab (https://about.gitlab.com).

## Some other CRAN packages for nonlinear modeling

`onls` (@R-onls-manual) -- The usual optimization in estimating nonlinear models
is the vertical difference between the dependent variable and the functional
model. `onls` minimizes the sum of **orthogonal** residuals. The objective is
therefore different and involves nontrivial extra calculation. A vignette with
the package and the blog article 
https://www.r-bloggers.com/2015/01/introducing-orthogonal-nonlinear-least-squares-regression-in-r/ 
give some description with illustrative graphs. `onls` appears to be limited to 
problems with one independent and one dependent variable. The Wikipedia article
https://en.wikipedia.org/wiki/Total_least_squares presents an overview of some
ideas, with references to the literature. The approach needs a wider 
discussion and tutorial examples to allow its merits to be judged than can be
included here.

`crsnls` (@R-crsnls-manual) -- This package allows nonlinear estimation by 
controlled random search via two methods. There is unfortunately no vignette.
A modest trial we carried out showed `nlsr::nlxb()` gave the same results in 
a small fraction of the time required by either of the methods in `crsnls`. The
method discussed in @tvrdik2007 claims better reliability in finding solutions
than a Levenberg-Marquardt code (actually from Matlab), but the tests were conducted
on the extreme NIST examples mentioned next.

`NISTnls` (@NISTnls-manual) -- This package provides R code and data for a set of
(numerically ill-conditioned) nonlinear least squares problems from the U.S. National 
Institute for Standards and Technology. These may not represent real-world situations.

`nlshelper` (@nlshelper-manual) -- This package, which unfortunately lacks a
vignette, provides a few utilities for summarizing, testing, and plotting 
non-linear regression models estimated with nls(), nlsList() or nlme() that
are linked or grouped in some way.

`nlsic` (@nlsic-manual) -- This solves nonlinear least squares problems with optional
equality and/or inequality constraints. It is clearly **not** about modeling, and
the input and output are quite different from class `nls` methods. However, there do
not appear to be other R packages with these capabilities.

`nlsMicrobio` (@nlsMicrobio-manual) -- Data sets and nonlinear regression models 
dedicated  to predictive microbiology, including a vignette, by authors of the
`nlstools` package.

`nlstools` (@p-nlstools) -- This package provides several tools for aiding the estimation 
of nonlinear models, particularly using `nls()`. The vignette is actually a journal
article,  and the authors have considerable experience in the subject.

`nlsmsn` (@m-nlsmsn) -- Fit univariate non-linear scale mixture of skew-normal(NL-SMSN) 
regression, with details in Garay, Lachos, and Abanto-Valle (2011). The problem 
here is to minimize an objective that is modified from the traditional sum of squared
residuals.

`nls.multstart` (@nlsmultstart) -- Non-linear least squares regression using
AIC scores with the Levenberg-Marquardt algorithm using multiple starting values 
for increasing the chance that the minimum found is the global minimum.

`nls2` (@nls2-22) -- Nonlinear least squares by brute force has similar
motivations to `nls.multstart`, but uses `nls()` within multiple trials. The author
has extensive expertise in R.

`nlstac` (@m-nlstac) -- A set of functions implementing the algorithm described in
Fernandez Torvisco et al. (2018) for fitting separable nonlinear regression curves. The 
special class of problem for which this package is intended is an important and 
difficult one. No vignette is provided, unfortunately.

`easynls`  -- Fit and plot some nonlinear models. Thirteen models are treated, but 
there is minimal documentation and no vignette. Package `nlraa` is to be preferred.

`nlraa`  (@MiguezNLRAA2021) -- a set of nonlinear *selfStart* models, primarily from
agriculture. Most include analytic Jacobian code.

`optimx` (@p-optimx) -- This provides 
optimizers that can be applied to minimize a nonlinear function which could be
a nonlinear sum of 
squares. NOT generally recommended if nonlinear 
least squares programs can be easily used, but provides a check and alternative
solvers.

## An illustrative example

The Hobbs weed infestation problem (@cnm79 [page 120]) is a growth curve modeling 
task that seems straightforward but is quite nasty. Its very succinct statement 
provides the "short reproducible example" much requested on R mailing lists. 
This is a real problem from a field researcher. The data and graph follow.

```{r ex01, echo=TRUE}
weed <- c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443,
          38.558, 50.156, 62.948, 75.995, 91.972)
tt <- 1:12
weeddf <- data.frame(tt, weed)
plot(weeddf, main="Hobbs weed infestation data")
```

Three suggested models for this data are (with names to allow for easy reference)

**Logistic3U**:
$$  y \approx  b_1 / (1 + b_2 * exp(- b_3 * t)) $$

**Logistic3S**:
$$ y \approx  100 * c_1 / (1 + 10 * c_2 * exp(- 0.1 * c_3 * t)) $$

**Logistic3T**:
$$ y \approx Asym / (1 + exp((xmid - t)/scal)) $$

where we will use `weed` for $y$ and `tt` for $t$. 
The functions above are equivalent, but the first is generally more awkward
to solve numerically due to its poor scaling. The parameters of the three forms
are related as follows:

$$   Asym =  b_1 = 100 * c_1 $$
$$ exp(xmid/scal)  =  b_2 = 10 * c_2 $$
$$ 1/scal  =  b_3 $$
To allow for simpler discussion, let us say that the parameters form a (named) vector
$p$ and the model function is called $model(p)$. The residuals can be written
either as $r_1 = y - model(p)$ or $r = model(p) - y$
since their sum of squares has an identical value. The second form
avoids a potential sign error if we need to evaluate derivatives.

We wish to minimize the sum of squared residuals, which is our loss (or objective) function. 
Starting with some guess for the parameters, we aim to alter these parameters to obtain a smaller
loss function. We then iterate until we can make no further progress.

Let us consider there are $n$ parameters and $m$ residuals. 
The loss function is 
$$ S(p) = r' r = \sum_{i=1}^m { r_i^2 }$$
The gradient of $S(p)$ is
$$ 2 * J' r$$
where the Jacobian $J$ is given by elements
$$ J_{i,j} = \partial r_i / \partial p_j $$
and the Hessian is defined by elements
$$ H_{i,j} = \partial ^2 S(p) / {\partial p_i \partial p_j} $$
If we expand the Hessian for nonlinear least squares problems, we find
$$  H_{i,j} = \sum_{k = 1}^m{ J_{k,i} J_{k,j}}  + \sum_{k = 1}^m {r_k * \partial{r_k}/{\partial p_i \partial p_j}}$$
Let us use $D_{i,j}$ for the elements of the second term of this expression.
What is generally called **Newton's method** for function minimization tries
to set the gradient to zero (to find a stationary point of the function $S(p)$).
This leads to **Newton's equation**
$$ H \delta = -g $$
Given a set of parameters $p$, we solve this equation for $\delta$, adjust $p$
to $p + \delta$ and iterate, hopefully to converge on a solution. 
Applying this to a sum of squares problem gives
$$ H \delta = (J'J + D) \delta = - J' r$$
In this expression, only the elements of $D$ have second partial derivatives. 
Gauss, attempting to model planetary orbits, had small residuals, and noted
that these multiplied the second partial derivatives of $r$, so he approximated
$$ H \approx J' J$$
by assuming $D \approx 0$. This results in the Gauss-Newton method where we
solve 
$$ J' J \delta = - J' r$$
though we can avoid some loss of accuracy by NOT forming the inner product matrix
$J' J$ and solving the linear least squares matrix problem 
$$ J \delta \approx -r $$
by one of several matrix decomposition methods.

In reality, there are many problems were $D$ should not be ignored, but the work
to compute it precisely is considerable. Many work-arounds have been 
proposed, of which the Levenberg-Marquardt stabilization 
(@Levenberg1944, @Marquardt1963) is
the most commonly used. For convenience, we will use "Marquardt", as we believe he
first incorporated the ideas into a practical computer program.

The usual suggestion is that $D$ be replaced by a multiple of the unit matrix or else
a multiple of the diagonal part of $J' J$. In low precision, some elements of 
$J' J$ could underflow to zero (@jn77ima), and a linear combination of both
choices is an effective compromise. Various choices for
$D$, as well as a possible line search along the
direction $\delta$ rather than a unit step (@Hartley1961), 
give rise to several variant
algorithms. "Marquardt's method" is a family of methods. Fortunately, most choices
work well.

### Problem setup

Let us specify in R the three model formulas and set some starting
values for parameters. These starting points are NOT equivalent and
are deliberately crude choices. Workers performing many calculations 
of a similar nature should try to provide good starting points to reduce
computation time and avoid finding a false solution.

```{r ex02set, echo=TRUE}
# model formulas
frmu <- weed ~ b1/(1+b2*exp(-b3*tt))
frms <- weed ~ 100*c1/(1+10*c2*exp(-0.1*c3*tt))
frmt <- weed ~ Asym /(1 + exp((xmid-tt)/scal))
#
# Starting parameter sets
stu1<-c(b1=1, b2=1, b3=1)
sts1<-c(c1=1, c2=1, c3=1)
stt1<-c(Asym=1, xmid=1, scal=1)
```

One of the useful features of `nls()` is the possibility of a `selfStart` model,
where starting parameter values are not required. However, if a `selfStart` model 
is not available, `nls()` sets all the starting parameters 
to 1. This is tolerable, but could be improved by using a set of values
that are all slightly different, which, in the case of the example 
model $y \,\sim\, a \,*\, exp(-b \,*\, x) + c\,*\,exp(-d \,*\, x)$
would avoid a singular Jacobian because $b$ and $d$ were equal in value. 
Program modifications to give a sequence like  1.0, 1.1, 1.2, 1.3 for the four 
parameters are fairly obvious.

It is also possible to provide R functions for the residual and Jacobian. 
This is usually more work for the user if the formula setup is possible.
To illustrate, we show the functions for the unscaled 3 parameter logistic.

```{r ex02fn, echo=TRUE}
# Logistic3U
hobbs.res  <-  function(x){ # scaled Hobbs weeds problem -- residual
  # This variant uses looping
  if(length(x) != 3) stop("hobbs.res -- parameter vector n!=3")
  y  <-  c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443, 
           38.558, 50.156, 62.948, 75.995, 91.972)
  tt  <-  1:12
  res  <-  x[1]/(1+x[2]*exp(-x[3]*tt)) - y
}

hobbs.jac  <-  function(x) { # scaled Hobbs weeds problem -- Jacobian
  jj  <-  matrix(0.0, 12, 3)
  tt  <-  1:12
  yy  <-  exp(-x[3]*tt)
  zz  <-  1.0/(1+x[2]*yy)
  jj[tt,1]   <-   zz
  jj[tt,2]   <-   -x[1]*zz*zz*yy
  jj[tt,3]   <-   x[1]*zz*zz*yy*x[2]*tt
  attr(jj, "gradient") <- jj
  jj
}
```

## Estimation of models specified as formulas

Using a formula specification was a principal advantage made
with `nls()` when it became available in S sometime in the 1980s. It uses a
Gauss-Newton (i.e., unstabilized) iteration with a step reduction line
search. This works very efficiently as long as $J$ is not ill-conditioned.
Below we see `nls()` does poorly on the example problem. To save page space,
we use 1-line result display functions from package `nlsr`, namely `pnls()` 
and `pshort()`.

```{r shortforms, echo=FALSE}
library(nlsr)
library(minpack.lm)
```

```{r ex02nls}
unls1<-try(nls(formula=frmu, start=stu1, data=weeddf))
snls1<-try(nls(formula=frms, start=sts1, data=weeddf))
tnls1<-try(nls(formula=frmt, start=stt1, data=weeddf))
```

Here we see the infamous "singular gradient" termination message of `nls()`.

### Solution attempts with nlsr

```{r ex02nlsr}
unlx1<-try(nlxb(formula=frmu, start=stu1, data=weeddf))
print(unlx1) 
snlx1<-try(nlxb(formula=frms, start=sts1, data=weeddf))
pshort(snlx1) # a short-form output
tnlx1<-try(nlxb(formula=frmt, start=stt1, data=weeddf))
pshort(tnlx1) # alternatively print(tnlx1)
```

Though we have found solutions, the Jacobian is essentially singular as 
shown by its singular
values. Note that these are **displayed** by package `nlsr` in a single column in
the output to provide a compact layout, but the values do **NOT** correspond to 
the individual parameters in 
whose row they appear; they are a property of the whole problem.

### Solution attempts with minpack.lm

```{r ex02minpack}
unlm1<-try(nlsLM(formula=frmu, start=stu1, data=weeddf))
pnls(unlm1)  # Short form of output
snlm1<-try(nlsLM(formula=frms, start=sts1, data=weeddf))
pnls(snlm1)
tnlm1<-try(nlsLM(formula=frmt, start=stt1, data=weeddf))
pnls(tnlm1) # short form to give sum of squares, else use summary(tnlm1)
```

### Solution attempts with gslnls

```{r ex02gslnls}
library(gslnls)
ugslnls1<-try(gsl_nls(fn = frmu, data = weeddf,  start = stu1))
pnls(ugslnls1) # to get sum of squares
sgslnls1<-try(gsl_nls(fn = frms, data = weeddf,  start = sts1))
pnls(sgslnls1) # Use summary() to get display
tgslnls1<-try(gsl_nls(fn = frmt, data = weeddf,  start = stt1))
pnls(tgslnls1) 
```

### Comparison notes for formula-setup solutions

`nlsr::nlxb()` uses `print()` to output standard errors and 
singular values of the Jacobian (for diagnostic purposes). By contrast, 
`minpack.lm::nlsLM()` and `nls()`
use `summary()`, which does NOT display the sum of squares, while `print()` 
gives the sum of squares, but not the standard error of the residuals.

The singular values allow us to gauge how "nearly singular" the 
Jacobian is at the solution, and the ratio of the smallest to largest of the 
singular values is a simple but effective measure. The ratios are
4.6641e-05 for Logistic3U, 0.021022 for Logistic3S, and 0.001055 for
Logistic3T, so Logistic3S is the "least singular".

The results from `nlsLM` and `gsl_nls` for the transformed model Logistic3T have a very 
large sum of squares, which may suggest that these programs have failed. 
Since `nls()`, `nlsLM()`, and `gsl_nls()` do not 
offer singular values, we need to extract the Jacobian and compute its
singular values. The following script shows how to do this, using as Jacobian what is called
the `gradient` element in the returned solution for these solvers.

```{r ex04singval, echo=TRUE, eval=TRUE}  
# for nlsLM
if (inherits(tnlm1, "try-error"))  {
   print("Cannot compute solution -- likely singular Jacobian")
 } else {  
   JtnlsLM <- tnlm1$m$gradient() # actually the Jacobian
   svd(JtnlsLM)$d # Singular values
}   
# for gsl_nls
if (inherits(tgslnls1, "try-error")) {
   cat("Cannot compute solution -- likely singular Jacobian")
} else {  
   JtnlsLM <- tgslnls1$m$gradient()
   svd(JtnlsLM)$d # Singular values
}   
```

We see that there are differences in detail, but the more important result is that
two out of three singular values are essentially 0. Our Jacobian is singular, and no 
method of the Gauss-Newton type should be able to continue. Indeed, from the parameters
reported at this saddle point, `nlsr::nlxb()` cannot proceed.

```{r ex05, echo=TRUE}  
stspecial<- c(Asym = 35.532,  xmid = 43376,  scal = -2935.4)
badstart<-nlxb(formula=frmt, start=stspecial, data=weeddf)
print(badstart)
```

## Functional specification of problems

We illustrate how to solve nonlinear least squares problems using
a function to define the residual. Note that `gsl_nls()` requires
a vector `y` that is the expected value of what we have called the
residual, but in this case, is actually the model. `gsl_nls` uses 
a numerical approximation for the Jacobian if the argument `jac` 
is missing. Note function `nlsr::pnlslm()` for a 1-line display of
the results of `minpack.lm::nls.lm()`.

```{r exhobbsfn, eval=TRUE}
hobnlfb<-nlfb(start=stu1, resfn=hobbs.res, jacfn=hobbs.jac)
pshort(hobnlfb) # use print(hobnlfb) for more detail
hobnlm<-nls.lm(par=stu1, fn=hobbs.res, jac=hobbs.jac)
pnlslm(hobnlm)  
hobgsln<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12))
pnls(hobgsln)
hobgsl<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12), jac=hobbs.jac)
pnls(hobgsl) # using analytic Jacobian
```

## Design goals, termination tests, and output objects

The output object of `nlxb()` is smaller than the class `nls` object returned
by `nls()`, `nlsLM()`, and `gsl_nls()`. Package `nlsr` emphasizes the solution 
of the nonlinear least squares problem rather than the estimation of a nonlinear 
model that fits or explains the data. The object of class `nls` allows for 
a number of specialized modeling and diagnostic extensions. For compatibility, 
the `nlsr` package has the function `wrapnlsr()`, for which `nlsr()` is an alias. 
This uses `nlxb()` to find good parameters, then calls `nls()` to return the class 
`nls` object. Unless particular modeling features are needed, the use of 
`wrapnlsr()` is unnecessary and wasteful of resources.

The design goals of the different tools may also be revealed in the so-called
"convergence tests" for the iterative solvers. In the manual page for `nls()` 
in R 4.0.0 there was the warning:

>  *Do not use `nls` on artificial "zero-residual" data.*

with suggested addition of small perturbations to the data. This admits 
`nls()` could not solve well-posed problems unless data is polluted
with errors. Zero-residual problems are not always artificial,
since problems in function approximation and nonlinear equations can be 
approached with nonlinear least squares. Fortunately, a small adjustment
to the "termination test" for the **program**, rather than for the
"convergence" of the underlying **algorithm**, fixes the defect. The test
is the Relative Offset Convergence
Criterion (see @BatesWatts81). This scales an estimated reduction in the loss
function by its current value. If the loss function is very small, we are close
to a zero-divide. Adding a small quantity to the divisor avoids trouble. 
In 2021, one of us (J. Nash) proposed that `nls.control()` have an 
additional parameter `scaleOffset` with a default value of zero. Setting it to a 
small number -- 1.0 is a reasonable choice -- allows small-residual problems 
(i.e., near-exact fits) to be dealt with easily. We call this the
**safeguarded relative offset convergence criterion**, and it has been in `nlsr`
since it was introduced. The default value 
gives the legacy behaviour. This improvement has been in the R distributed code 
since version 4.1.0.

Additional termination tests can be used. `nlsr` has a **small sum of 
squares** test (**smallsstest**) that compares the latest evaluated sum of squared
(weighted) residuals to `e4` times the initial sum of squares, where
`e4 <- (100*.Machine$double.eps)^4` is approximately 2.43e-55.

Termination after what may be considered excessive computation is also important.
`nls()` stops after `maxiter` "iterations". The meaning of "iteration" may require an 
examination of the code. We terminate execution when the number of residual or 
Jacobian evaluations exceed set limits. Generally, we prefer larger limits than the 
default `maxiter = 50` of `nls()` to avoid stopping early.

### Returned results of nls() and other tools

As mentioned, the output of `nls()` is an object of class "nls" which has a quite rich structure
described in the manual file or revealed by applying the `str()` function to the
result of `nls()`.

The complexity of this object is a challenge to users. Let us use the
`result` as the returned object from `nls()`, `minpack.lm::nlsLM()` or,
`gslnls::gsl_nls()`. For example, the `data` return element is an R symbol. To
actually access the data from this element, we need to use the syntax:
```
eval(parse(text=result$data))
```
However, if the call is made with `model=TRUE`, then there is a returned element
`model` which contains the data, and we can list its contents using:
```
ls(result$model)
```
and if there is an element called `xdata`, then it can be accessed as
`result$model$xdata`.

By contrast, `nlsr::nlxb()` returns a much simpler structure of 11 items
in one level. Moreover, `nlxb` explicitly returns the sum of squares, the residual vector, Jacobian, and counts of evaluations.

### When to compute ancillary information

Tools that produce a class `nls` output object create a rich set of functions
and structures that are then used in a variety of modeling tasks, including the
least squares solution. By contrast,
`nlsr` computes quantities as they are requested or needed, with additional
features in separate functions. For example, the singular values of the Jacobian 
are actually computed in the `print` and `summary` methods for the result. 
These two approaches
lead to different consequences for performance and how features are
provided. `nlsr` has antecedents in the methods of @cnm79, where storage for
data and programs was at a ridiculous premium in the small computers of the era.
Thus the code in `nlsr` is likely of value for workers to copy and modify
for customized tools.

## Jacobian calculation

Gauss-Newton/Marquardt methods all need a Jacobian matrix at each iteration. 
By default, `nlsr::nlxb()` will try to evaluate this using analytic 
expressions using symbolic and automatic differentiation tools. 
When using a formula specification of the model, `nls()`, `minpack.lm::nlsLM()`
and `gslnls::gsl_nls()` use a finite difference approximation to compute the
Jacobian, though `gsl_nls()` does have an option to attempt symbolic expressions.
Package `nlsr` provides, via appropriate calling syntax, four numeric approximation 
options for the Jacobian, with a further control `ndstep` for the size of the
step used in the approximation.

Using the "gradient" attribute of the output of the Jacobian function to hold
the Jacobian matrix lets us embed this in the **residual** function as well,
so that the call to `nlsr::nlfb()` can be made with the same name used for 
both residual and Jacobian function arguments. This programming trick saves a 
lot of trouble for the package developer, but it can be a nuisance for users 
trying to understand the code.

As far as we can understand the logic in `nls()`, the Jacobian computation during
parameter estimation is carried out within the called C-language program
and its wrapper R code function `numericDeriv()`, part of
`./src/library/stats/R/nls.R`
in the R distribution source code. This is used to provide Jacobian information in 
the `nlsModel()` and `nlsModel.plinear()` functions, which are **not** exported for 
general use. `gsl_nls()` also appears to use `numericDeriv()`.

`numericDeriv()` uses a 
simple forward difference approximation of derivatives, though a central 
difference approximation can be specified in control parameters. 
We are unclear why `numericDeriv()` in base R calls `C_numeric_deriv`,
as we were easily able to create a more compact version entirely in R. 
See *https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS*.

`minpack.lm::nlsLM()` invokes `numericDeriv()` in its local 
version of `nlsModel()`, but it appears to use an internal approximate Jacobian 
code from 
the original Fortran `minpack` code, namely, `lmdif.f`. Such differences in approach
can lead to different behaviour, usually minor, but sometimes annoying with
ill-conditioned problems.

- A pasture regrowth problem (@Huet2004, page 1, based on @Ratkowsky1983) has a
  poorly conditioned Jacobian and `nls()` fails with "singular gradient". 
  Worse, numerical approximation to the Jacobian gives the error
  "singular gradient matrix at initial parameter estimates" for `minpack.lm::nlsLM`
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of `nlsr::nlxb` give a solution.

- Karl Schilling (private communication) provided an example where a model specified
  with the formula `y ~ a * (x ˆ b)` causes `nlsr::nlxb` to fail because the partial
  derivative w.r.t. `b` is `a * (x^b * log(x))`. If there is data for which `x = 0`,
  the derivative is undefined, but the model can be computed. In such cases, 
  we observed that `nls()` and 
  `minpack.lm::nlsLM` found a solution, though this seems to be a lucky accident.

### Jacobian code in selfStart models

Analytic Jacobian code can be provided to all the solvers discussed.
Most `selfStart` models that automatically provide starting parameters
also include such code. There is documentation in R
of `selfStart` models, but their construction is non-trivial. A number 
of such models are included with base R in `./src/library/stats/R/zzModels.R`,
with package `nlraa` (@MiguezNLRAA2021) providing a richer set. 
There are also some in the now-archived package `NRAIA`.
These provide the Jacobian in the "gradient" attribute of the "one-sided" formula
that defines each model, and these Jacobians are often the analytic forms.

The `nls()` function, after
computing the "right-hand side" or `rhs` of the residual, checks to see if the
"gradient" attribute is defined, otherwise using `numericDeriv()` to compute a 
Jacobian into that attribute. This code is within the `nlsModel()` or
`nlsModel.plinear()` functions. The use of analytic Jacobians
almost certainly contributes to the good performance of `nls()` on `selfStart` 
models.

The use of `selfStart` models with
`nlsr` is described in the "Introduction to nlsr" vignette. However, since `nlsr`
generally can use very crude starting values, we have rarely needed them, though
it should be pointed out that our work is primarily diagnostic. If we were carrying
out a large number of similar estimations, such initial parameters are 
critical to efficiency.

In considering `selfStart` models, we noted that the base-R function `SSlogis`
is intended to solve problem **Logistic3T** above.  When this function is used
via `getInitial()` to  find
starting values, it actually calls `nls()` with the 'plinear' algorithm 
and finds a (full) solution. It then passes the solution coefficients to the default 
algorithm unnecessarily. Moreover, the implicit double call is, in our view, prone to creating 
errors in code maintenance. 
To provide simpler starting parameters, the function `SSlogisJN`
is now part of the package `nlsr`, but is most useful for `nls()`.

Users may also want to profit from the Jacobian code of `selfStart` models 
but supply explicit starting values other than those suggested by `getInitial()`.
This does not appear to be possible with `nls()`. `nlsr::nlxb()` always requires 
starting parameters, and can either use `getInitial()` to find them from the `selfStart`
model or provide explicit values, but the `selfStart` model is used in the formula 
for the `nlxb()` call.

<!-- tssx <-try(nls(weed ~ SSlogis(tt, Asym, xmid, scale), start=stt1, data=weeddf)) -->
<!-- tssx <-try(nls(weed ~ SSlogis(tt, Asym=1, xmid=1, scale=1),  data=weeddf)) -->
<!-- tssx <-try(nls(weed ~ SSlogis(tt, Asym, xmid, scale),  data=weeddf))summary(tssx) -->

We are also surprised that the 
analytic expressions for the Jacobian ("gradient") in the `SSLogis` function 
and others save quantities in "hidden"
variables, i.e., with names preceded by ".". 
These are then not displayed by the `ls()`
command, making them difficult to access by users who may wish to create
their own selfStart model via copy and edit.
Interactive tools, such as "visual fitting" (@nash1996nonlinear) might be
worth considering as another way to find starting parameters, but we know of no
R capability of this type.

As a side note, the introduction of `scaleOffset` in R 4.1.1 to deal with the 
convergence test for small residual problems now requires that the `getInitial()`
function have dot-arguments (`...`) in its argument list. This illustrates the
entanglement of many features in `nls()` that complicate its maintenance and
improvement.

## Bounds constraints on parameters

For many problems, we know that parameters cannot be bigger or smaller than some
externally known limits. Such limits should be built into models, but there 
are some important details for using the tools in R.

- `nls()` can only impose bounds if the `algorithm="port"` argument is 
used in the call. Unfortunately, the documentation warns us:

  *The algorithm = "port" code appears unfinished, and does not even 
   check that the starting value is within the bounds. Use with 
   caution, especially where bounds are supplied.*

- `gsl_nls()` does not offer bounds.

- bounds are part of the default method for package `nlsr`.

- `nlsLM()` includes bounds in the standard call, but we have observed cases
   where it fails to get the correct answer. From an examination of the code,
   we believe the authors have not taken into account all possibilities, though
   all programs have some weakness regarding constrained optimization in that
   programmers have to work with assumptions on the scale of numbers, 
   and some problems will be outside the scope envisaged.

```{r ex10, echo=TRUE} 
# Start MUST be feasible i.e. on or within bounds
anlshob1b <- nls(frms, start=sts1, data=weeddf, lower=c(0,0,0),
             upper=c(2,6,3), algorithm='port')
pnls(anlshob1b) #  check the answer (short form)
# nlsLM seems NOT to work with bounds in this example
anlsLM1b <- nlsLM(frms, start=sts1, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pnls(anlsLM1b)
# also no warning if starting out of bounds, but gets a good answer!!
st4<-c(c1=4, c2=4, c3=4)
anlsLMob <- nlsLM(frms, start=st4, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pnls(anlsLMob)
# Try nlsr::nlxb()
anlx1b <- nlxb(frms, start=sts1, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pshort(anlx1b)
```

### Philosophical considerations

Bounds on parameters raise some interesting questions about how 
uncertainty in parameter estimates should be computed or reported. That is, the
traditional "standard errors" are generally taken to imply symmetric intervals about
the point estimate in which the parameter may be expected to be found with some
probability under certain assumptions.
Bounds change those assumptions. Hence, `nlsr::nlxb()` does not compute standard
errors nor their derived statistics when bounds are active.

### Fixed parameters (masks)

Let us try to fix (mask) the first parameter in the first two example problems.

```{r ex10m, echo=TRUE}
# Hobbsmaskx.R -- masks with formula specification of the problem
require(nlsr); require(minpack.lm); traceval<-FALSE
stu <- c(b1=200, b2=50, b3=0.3) # a default starting vector (named!)
sts <- c(c1=2, c2=5, c3=3) # a default scaled starting vector (named!)
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, 3), trace=traceval))
print(anxbmsk1)
anlM1 <- try(nlsLM(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
			upper=c(200, 60, 3), trace=traceval))
pnls(anlM1)
anlsmsk1 <- try(nls(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
		upper=c(200, 60, 3),  algorithm="port", trace=traceval))
pnls(anlsmsk1)
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- nlxb(frms, start=sts, data=weeddf, lower=c(2,0,0),
                  upper=c(2,6,30))
print(anlxmsks1)
anlshmsk1 <- nls(frms, start=sts, trace=traceval, data=weeddf, lower=c(2,0,0),
             upper=c(2,6,30), algorithm='port')
pnls(anlshmsk1)
anlsLMmsks1 <- nlsLM(frms, start=sts, data=weeddf, lower=c(2,0,0),
                 upper=c(2,6,30))
pnls(anlsLMmsks1)

# Test with all parameters masked
anlxmskall<- try(nlxb(frms, start=sts, data=weeddf, lower=sts, upper=sts))
print(anlxmskall)
```

`nlsr` has an output format that indicates the constraint status of the parameter
estimates. For `nlsr`, we have **chosen** to suppress the calculation of 
approximate standard errors in the parameters when constraints are active because
their meaning under constraints is unclear, though we believe this 
policy worthy of discussion and further investigation.

## Stabilization of Gauss-Newton computations

All four major tools illustrated solve some variant of the Gauss-Newton 
equations. `nls()` uses a modification of an approach suggested by @Hartley1961,
while `nlsr`, `gslnls`, and `minpack.lm` use flavours of @Marquardt1963. `gslnls`
offers an accelerated Marquardt method and three alternative methods which we
have not investigated.
Control settings for `nlxb()` or `nlfb()` allow exploration of Hartley and Marquardt 
algorithm variants. In general, the Levenberg-Marquardt stabilization 
is important in obtaining 
solutions in methods of the Gauss-Newton family, as `nls()` terminates 
too frequently and unnecessarily with `singular gradient` errors.

### Programming language

An important choice made in developing `nlsr` was to code entirely within the R 
programming language. `nls()` uses a mix of R, C, and Fortran, as does `minpack.lm`. 
`gslnls` is an R wrapper to various C-language routines in the Gnu Scientific 
Library (@GSL-manual).
Generally, keeping to a single programming language can allow for 
easier maintenance and upgrades. It also avoids some work when there are changes 
or upgrades to libraries for the non-R languages. 
R is usually considered slower 
than most computing environments because it keeps track of objects and 
because it is usually interpreted. In recent years, the 
performance penalty for using code entirely in R has been much reduced
with the just-in-time compiler and other improvements. All-R computation 
may now offer acceptable performance. In `nlsr`, the use of R may be a
smaller performance cost than the aggressive approach to a solution,
which can cause more iterations to be used.

## Data sources for problems

`nls()` can be called without specifying the `data` argument. In this case, it will
search in the available environments (i.e., workspaces) for suitable data objects. 
We do NOT like this approach, but it is "the R way". R allows users to leave many 
objects in the default (`.GlobalEnv`) workspace. Moreover, users have to actively 
suppress saving this workspace (`.RData`) on exit; otherwise, any such file in 
the path will be loaded on startup. R users in our
acquaintance avoid saving the workspace because of lurking data and
functions that may cause unwanted results.

### Feature: Subsetting

`nls()` and other class `nls` tools accept an argument `subset`. This acts through 
the mediation of `model.frame`, which not obvious in the 
source code files `/src/library/stats/R/nls.R` and `/src/library/stats/src/nls.C`. 
Having `subset` at the level of the call to a function like
`nls()` saves effort, but it does mean that the programmer of the 
solver needs to be aware of the
origin (and value) of objects such as the data, residuals and Jacobian. 
By preference, 
we would implement subsetting by zero-value weights, with observation counts
(and degrees of freedom) computed via the numbers of non-zero weights. Alternatively,
we would extract a working dataframe from the relevant elements in the original.

### Feature: na.action (missing value treatment)

`na.action` is an argument to the `nls()` function, but it does not appear 
obviously in the source code, often being handled behind the scenes after 
referencing the option `na.action`. This feature also changes the data 
supplied to our nonlinear least squares solver. 
A useful, but possibly dated, description is given in:
*https://stats.idre.ucla.edu/r/faq/how-does-r-handle-missing-values/*.
The typical default action, which can be seen by using the command
`getOption("na.action")`
is `na.omit`. This option removes from computations any observations
containing missing values (i.e. any row of a data frame containing an NA). 
`na.exclude` does much of the same for solver computations, but keeps the rows with 
NA elements so that predictions are in the correct row position. We recommend that 
workers actually test output to verify the behaviour is as wanted.
See *https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r*.
As with `subset`, our concern with `na.action` is that users may be unaware of the
effects of an option they may not even be aware has been set. Should `na.fail` be the default?

## Feature: model frame

`model` is an argument to the `nls()` and related functions, which is documented as:

> **model** logical. If true, the model frame is returned as part of the object. Default is FALSE.

Indeed, the argument only gets used when `nls()` is about to return its result
object, and the
element `model` is NULL unless the calling argument `model` is TRUE. (Using the same 
name could
be confusing.) Despite this, the model frame is used within the function code in 
the form of the object
`mf`. We feel that users could benefit from more extensive documentation and 
examples of its use since it is used to implement features like `subset`.

## Weights on observations

All four main tools we consider here allow a `weights` argument that
specifies a vector of fixed weights the same length as the number of residuals. 
Each residual is multiplied by the square root of the corresponding weight. 
Where available, the values returned by the
`residuals()` function are weighted, and the `fitted()` or `predict()` function are 
used to compute raw residuals.

While fixed weights may be useful, there are many problems for which
we want weights that are determined at least partially from the model 
parameters, for
example, a measure of the standard deviation of observations.
Such dynamic weighting situations are discussed in the vignette 
"Introduction to nlsr" of package `nlsr` in section 
*Weights that are functions of the model parameters*.
`minpack.lm` offers a function `wfct()` to 
facilitate such weighting. Care is advised
in applying such ideas.

### Weights in returned functions from nls()

The function `resid()` (an alias for `residuals()`) gives WEIGHTED residuals, 
as does, for example, `result$m$resid()`. The function `nlsModel()`, which we 
have had to extract from the base R code and explicitly `source()` because it 
is not exported to the working namespace,
allows us to compute residuals for particular coefficient sets.

```{r nlswtx, echo=TRUE}
wts <- 0.5^tt # simple weights
frmlogis <- weed ~ Asym/(1 + exp((xmid - tt)/scal))
Asym<-1; xmid<-1; scal<-1
nowt<-nls(weed ~ SSlogis(tt, Asym, xmid, scal)) # UNWEIGHTED
rnowt<-nowt$m$resid() # This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.
attr(rnowt, "gradient")<-NULL; rnowt
usewt <- nls(weed ~ SSlogis(tt, Asym, xmid, scal), weights=wts)
rusewt<-usewt$m$resid() # WEIGHTED. Does NOT take coefficients.
attr(rusewt,"gradient")<-NULL; rusewt
source("nlsModel.R")
nmod0 <- nlsModel(frmlogis, data=weeddf, start=c(Asym=1, xmid=1, scal=1), wts=wts)
rn0<-nmod0$resid() # Parameters are supplied in nlsModel() `start` above.
attr(rn0,"gradient")<-NULL; rn0
nmod <- nlsModel(frmlogis, data=weeddf, start=coef(usewt), wts=wts)
rn<-nmod$resid()
attr(rn,"gradient")<-NULL; rn
```

## Minor issues with nonlinear least-squares tools

### Interim output from the "port" algorithm

As the `nls()` **man** page states, when the "port" algorithm is used with the
`trace` argument TRUE, the iterations display the objective function value which 
is 1/2 the sum of squares (or deviance). The trace display is likely
embedded in the Fortran of the `nlminb` routine that is called to execute 
the "port" algorithm, but the factor of 2 discrepancy is nonetheless
unfortunate for users.

### Failure to return the best result achieved

If `nls()` reaches a point where it cannot continue but has not found a point 
where the relative offset convergence criterion is met, it may simply exit,
especially if a "singular gradient" (singular Jacobian) is found. However, 
this may occur AFTER the function has made considerable progress in reducing 
the sum of squared residuals. 
<!-- An example is to be found in the `Tetra_1.R` example from the `nlsCompare` package. -->
<!-- ?? AB: Are we far enough along to include reference? -->
Here is an abbreviated example:

```{r tetrarun}
time <- c( 1,  2,  3,  4,  6 , 8, 10, 12, 16)
conc <- c( 0.7, 1.2, 1.4, 1.4, 1.1, 0.8, 0.6, 0.5, 0.3)
NLSdata <- data.frame(time,conc)
NLSstart <- c(lrc1 = -2, lrc2 = 0.25, A1 = 150, A2 = 50) # a starting vector (named!)
NLSformula <- conc ~ A1 * exp(-exp(lrc1) * time) + A2 * exp(-exp(lrc2) * time)
tryit <- try(nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE))
print(tryit)
```

Note that the sum of squares has been reduced from 61216 to 1.6211, but 
unless `trace` is invoked, the user will not get any information about this.
This almost trivial change to the `nls()` function could be 
useful to R users.

## Estimating models that are partially linear

The variable projection method (@Golub1973, @OlearyRust13) is 
generally much more effective than general approaches in finding good solutions 
to nonlinear least squares 
problems when some of the parameters appear linearly. In our logistic examples, 
the asymptote parameters are an illustration. However, identifying which parameters
are linear and communicating this information to estimating functions 
is not a trivial task. `nls()` has an option `algorithm="plinear"` that allows some partially linear
models to be solved. The other tools, as far as we are aware, do not offer any
such capability. The `nlstac` package uses a different algorithm for similar goals.

Within `nls()` itself we must, unfortunately, use
different specifications with different `algorithm` options.
For example, the explicit model `y ~ a * x + b` does not work with the linear
modeling function `lm()`, which requires this model to be specified as `y ~ x`.
Within `nls()`, consider the following FOUR different specifications for the same 
problem, plus an intuitive choice, labeled `fm2a`, that does NOT work. 
In this failed attempt, putting the `Asym` parameter in the model causes the
`plinear` algorithm
to try to add another term to the model. We believe this is unfortunate, and would
like to see a consistent syntax. At the time of writing, we do
not foresee a resolution for this issue. In the example, we have NOT evaluated
the commands to save space.

```{r log4ways, echo=TRUE, eval=FALSE}
DNase1 <- subset(DNase, Run == 1) # select the data
## using a selfStart model - do not specify the starting parameters
fm1 <- nls(density ~ SSlogis(log(conc), Asym, xmid, scal), DNase1)
summary(fm1)

## using conditional linearity - leave out the Asym parameter
fm2 <- nls(density ~ 1 / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(xmid = 0, scal = 1),
                 algorithm = "plinear")
summary(fm2)

## without conditional linearity
fm3 <- nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1,
                 start = list(Asym = 3, xmid = 0, scal = 1))
summary(fm3)

## using Port's nl2sol algorithm
fm4 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(Asym = 3, xmid = 0, scal = 1),
                 algorithm = "port"))
summary(fm4)

## using conditional linearity AND Asym does NOT work
fm2a <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)), 
                 data = DNase1, start = list(Asym=3, xmid = 0, scal = 1),
                 algorithm = "plinear", trace = TRUE))
summary(fm2a)
```

## Models with indexed parameters

Some models have several common parameters and 
others that are tied to particular cases. 
The **man** file for `nls()` includes an example of a situation in which
parameters are indexed but which uses the "plinear" option as an added complication.
Running this example reveals that the answers for the parameters are NOT indexed
as in a vector. 
<!-- e.g. `coef(musc.2)` is a single level vector of parameters.  -->
That is, we do not see `a[1], a[2], a[3]` but `a1, a2, a3`. 
<!-- This is because the model -->
<!-- must integrate all the parameters because `th` is a common parameter across the index `Strip`. -->
This is no doubt because programming for indexed parameters is challenging.
<!-- However, we believe the current structure could cause confusion and error, and  -->
<!-- propose an alternative approach below. -->
We note that there are capabilities in packages for mixed-effects modeling
such as `nlme` (@p-nlme), `bbmle` (@p-bbmle), and `lme4` (p-lme4) for 
estimating models of such type.

<!-- Ben Bolker mentioned a possibility of using a nonlinear model in  -->
<!-- ??I think?? nlme with some sort of linear model form of the parameters. -->
<!-- May need to mention. -->

<!-- ===```{r nlsindx1} -->
<!-- ## The muscle dataset in MASS is from an experiment on muscle -->
<!-- ## contraction on 21 animals.  The observed variables are Strip -->
<!-- ## (identifier of muscle), Conc (Cacl concentration) and Length -->
<!-- ## (resulting length of muscle section). -->
<!-- if(! requireNamespace("MASS", quietly = TRUE)) stop("Need MASS pkg") -->
<!-- mm<- MASS::muscle[1:12,] # take only 1st few values of Strip (TRUNCATION OF EXAMPLE) -->
<!-- mm<-droplevels(mm) # remove unused levels after truncation -->
<!-- nlev <- nlevels(mm) -->
<!-- withAutoprint({ -->
<!--   ## The non linear model considered is -->
<!--   ##       Length = alpha + beta*exp(-Conc/theta) + error -->
<!--   ## where theta is constant. For now alpha and beta do NOT vary with Strip. -->
<!--   with(mm, table(Strip)) # 2, 3 or 4 obs per strip -->
<!--   nl <- nlevels(mm$Strip) -->
<!--   ## We first use the plinear algorithm to fit an overall model, -->
<!--   ## ignoring that alpha and beta might vary with Strip. -->
<!--   musc.1 <- nls(Length ~ cbind(1, exp(-Conc/th)), mm, -->
<!--                 start = list(th = 1), algorithm = "plinear") -->
<!--   summary(musc.1) -->

<!--   ## Then we use nls' indexing feature for parameters in non-linear -->
<!--   ## models to use the conventional algorithm to fit a model in which -->
<!--   ## alpha and beta vary with Strip.  The starting values are provided -->
<!--   ## by the previously fitted model. -->
<!--   ## Note that with indexed parameters, the starting values must be -->
<!--   ## given in a list (with names): -->
<!--   b <- coef(musc.1) -->
<!--   musc.2 <- nls(Length ~ a[Strip] + b[Strip]*exp(-Conc/th), data=mm, -->
<!--                 start = list(a = rep(b[2], nl), b = rep(b[3], nl), th = b[1])) -->
<!--   summary(musc.2) -->
<!-- }) -->
<!-- === ``` -->

## Tests and use-case examples

Maintainers of packages need suitable tests and use-case examples in order

- to ensure packages work properly, in particular, giving
results comparable to or better than the functions they are to replace. 
- to test individual solver functions to ensure they work across the range of 
  calling mechanisms, that is, different ways of supplying inputs to the solver(s);
- to pose "silly" inputs to see if these bad inputs are caught by the programs.

Such goals align with the aims of **unit testing** 
(e.g., *https://towardsdatascience.com/unit-testing-in-r-68ab9cc8d211*, @HWtestthat11, @HWdevtools21
and the conventional R package testing tools).
In our work, one of us has developed a working prototype package at *https://github.com/ArkaB-DS/nlsCompare* .
A primary design objective of this is to allow the summarisation of multiple tests in
a compact output. The prototype has a vignette to illustrate its use.

## Documentation and resources

In our investigation, we built several resources, which are now part of 
the repository *https://github.com/nashjc/RNonlinearLS/*. Particular items
are:

- A BibTex bibliography for use with all documents in this project, but which 
  has wider application to nonlinear least squares projects in general
  (*https://github.com/nashjc/RNonlinearLS/blob/main/BibSupport/ImproveNLS.bib*).

- `MachID.R` offers a suggested concise summary function
  to identify a particular computational system used for tests. A discussion 
  of how it was built and the resources needed across platforms is given in
  at *https://github.com/nashjc/RNonlinearLS/tree/main/MachineSummary*.

- @PkgFromRbase22 is an explanation of the construction of the `nlspkg` from the
  `nls()` code in R-base.

- As the 2021 Summer of Code period was ending, one of us (JN) was invited to prepare
  a review of optimization in R. Ideas from the present work have been instrumental
  in the creation of @NashWires22.

<!-- ## Should nls() be refactored? -->

<!-- When we started *Improvements to nls()* for the 2021 Google Summer of Code we wanted: -->

<!-- - to provide a packaged version of `nls()` (call it `nlsalt`) coded entirely in R  -->
<!--   that matches the features in base R or what is packaged in `nlspkg` as described  -->
<!--   in @PkgFromRbase22. Our intent in providing a purely R version of the code was -->
<!--   either to replace `nls()` or to serve as a reference version aiding maintenance  -->
<!--   of C, Fortran, or other components of the software. -->

<!-- - to streamline the overall `nls()` infrastructure. By this we mean a  -->
<!--   refactoring of the routines so they are better suited to maintenance of both -->
<!--   the existing `nls()` methods and features as well as new features we or others -->
<!--   would like to add. -->

<!-- - to explain what we do, either in comments, examples, or separate maintainer  -->
<!--   documentation. While R and its packages are generally very well documented for -->
<!--   usage, that documentation -->
<!--   rarely extends as far as it might for code maintenance. The paucity of such -->
<!--   documentation is exacerbated by the mixed R/C/Fortran code base, -->
<!--   where seemingly trivial differences in things like representation of numbers or -->
<!--   base index for arrays lead to traps for unwary programmers. -->

<!-- These goals echo themes in the recent discussion by @Vidoni21 and accompanying  -->
<!-- commentary. -->

<!-- We found the legacy code and examples prevented us achieving a revised `nls()`  -->
<!-- with all its current capabilities. Unfortunately, `nls()` is part of the base R -->
<!-- distribution,  -->
<!-- so users are drawn to using it even when it is less than ideal. A possible way forward -->
<!-- is to  -->

<!-- - use a packaged version of `nls()` to provide legacy features, with notification -->
<!--   to users that there are other tools available -->

<!-- - if maintainers are willing, to merge and rationalize the capabilities of  -->
<!--   `minpack.lm`, `nlsr` and `gslnls` -->

<!-- - to encourage the (separate) development of capabilities such as indexed -->
<!--   parameter models if these are in demand. -->

<!-- This direction does not, of course, resolve the inconsistencies in model  -->
<!-- specification in formulas for nonlinear models, which undoubtedly is an issue -->
<!-- for other modeling tools.   -->

<!-- A possible workaround to allow consistency for the `plinear` algorithm  -->
<!-- would be to specify the linear parameters when the `plinear` option is specified.  -->
<!-- For example, we could use `algorithm="plinear(Asym)"` while keeping the general  -->
<!-- model formula from the default specification. This would allow  -->
<!-- for the output of different `algorithm` options to be consistent. However, we  -->
<!-- have not yet tried to code such a change, and welcome discussion from those working -->
<!-- with other packages using model formulas. -->

<!-- A further complication of model specification in R is that some formulas require -->
<!-- the user to surround a term with `I()` to inhibit interpretation of arithmetic  -->
<!-- operators. We would prefer that formulas be usable without this complication. -->

## Future of nonlinear model estimation in R

Given its importance to R, it is possible that `nls()` will remain more or less 
as it has been for the past several decades. If so, the focus of discussion 
should be the measures needed to secure its continued operation for legacy 
purposes and how that may be accomplished. 
We welcome an opportunity to participate in such conversations.

To advance the stability and maintainability of R, we believe the program 
objects (R functions)
that are created by tools such as `nls()` should have minimal 
cross-linkages and side-effects.
The aspects of `nls()` that concern us follow.

- R tools presume data and parameters needed are available in an 
  accessible environment. This provides a compact syntax to invoke the
  calculations, but the wrong data can be used if the 
  internal search finds a valid name that is not the object we want, 
  or if `subset` or `na.action` settings modify the selection, or 
  `weights` are applied.

- Mixing of R, C and Fortran code adds to the burden of following the 
  program logic.

- The class `nls` structure simplifies calls with its rich set of functions, but
  also adds to the task of understanding what has been done. A design that isolates
  the setup, solution, and post-solution parts of complicated calculations reduces 
  the number of objects that must be kept in alignment.

Given the existence of examples of good practices such as analytic derivatives, 
stabilized solution of Gauss-Newton equations, and bounds-constrained parameters, 
base R tools should 
be moving to incorporate them. These capabilities are available now by 
using several tools, but it would be helpful if they were unified.

# Acknowledgments

Hans Werner Borchers was helpful in developing the GSoC project motivating this article
and in comments on this and related work. Heather Turner co-mentored the project 
and helped guide the progress of the work. Exchanges with Fernando Miguez helped to
clarify aspects of `selfStart` models and instigated the "Introduction to nlsr" 
vignette. Colin Gillespie (package `benchmarkme`) has been helpful in guiding our 
attempts to succinctly summarize computing environments.

# References

<!-- ------------------------------ -->

<!-- ## Comparison dimensions -->

<!-- The tools for nonlinear estimation we consider have been built with different -->
<!-- goals and contexts. We will make our comparison considering the overlapping -->
<!-- headings summarized below. -->

<!-- ### Modeling versus minimizing a sum of squares -->

<!-- The goal of many statisticians and other workers is to estimate a model. The -->
<!-- minimization of a sum of squared residuals -- the nonlinear least squares problem -- -->
<!-- is the mechanism to estimate the model. Some tools focus on modeling,  -->
<!-- so attempt to provide measures of uncertainty and other information.  -->
<!-- The interface to modeling tools will generally try to streamline that task. By -->
<!-- contrast, nonlinear least squares tools are trying to solve the more limited -->
<!-- numerical task. -->

<!-- ### Programming language(s) -->

<!-- R tools are our interest, but they often call programs in other programming -->
<!-- languages. There are advantages in using a well-established program that has -->
<!-- been used and tested. Early versions of R were relatively sluggish in  -->
<!-- computational speed. The downside is that we must interface between two or more -->
<!-- programming languages, adding to the risk of errors. Changes to one or the  -->
<!-- other language or its library support may -->
<!-- force maintenance issues. Moreover, the pool of expertise with languages like C -->
<!-- and Fortran is diminishing. -->

<!-- ### Internal solver stabilization -->

<!-- Early nonlinear least squares programs generally used the Gauss-Newton method  -->
<!-- (@Hartley1961). This is often very effective and efficient, but for a quite large  -->
<!-- proportion of problems can fail to find the solution. Stabilization of the  -->
<!-- computations is relatively straightforward in principle, and overcomes most -->
<!-- of the failures, though sometimes at the cost of more calculations. -->

<!-- ### Jacobian calculation -->

<!-- The necessarily iterative solution of a nonlinear least squares problem  -->
<!-- to carry out nonlinear modeling requires that we can find smaller values -->
<!-- of a sum of squared residuals. This implies there is a "downhill" direction. -->
<!-- The gradient of the residual sum of squares is computed as its product -->
<!-- with the Jacobian matrix: the partial derivatives of the residual vector  -->
<!-- with respect to the modeling parameters. -->

<!-- Most researchers are challenged to program the differential calculus  -->
<!-- required to specify the Jacobian to a nonlinear estimation tool.  -->
<!-- The workaround is generally to use a numerical approximation, -->
<!-- but there is the possibility of using symbolic or automatic differentiation. -->
<!-- There are also particular models where workers have provided code for the -->
<!-- Jacobian. -->

<!-- ### Algorithm and termination choices -->

<!-- Iterative algorithms need to be stopped at some point, and how termination -->
<!-- is forced may alter results. Furthermore, even supposedly similar algorithms -->
<!-- may differ greatly in detail.  -->

<!-- ### Constraints on parameters -->

<!-- Reality imposes conditions on parameters. While general constraints can -->
<!-- be difficult to include, bounds on parameters can be accommodated by many  -->
<!-- solvers. For some reason, fixed parameters (masks) are not often considered -->
<!-- but are convenient when a parameter has a commonly accepted value that we -->
<!-- later wish to allow to be estimated from the data. -->

<!-- ### Special capabilities -->

<!-- Users often have particular problems that introduce other requirements. It is -->
<!-- common to weight observations where their reliability is thought to differ. -->
<!-- Some situations present a number of related situations where a combined model -->
<!-- with some common and some local parameters are present, and the local parameters -->
<!-- could be indexed. The common exponential decay and related **partially linear** -->
<!-- models (sometimes called **separable** models) is a difficult one that can  -->
<!-- benefit from special treatment.  -->

<!-- ### Output  -->

<!-- While nonlinear estimation aims to return a set of parameters for a -->
<!-- particular model, there is a variety of additional information which may be -->
<!-- provided. From the perspective of the nonlinear least squares problem, there -->
<!-- are measures of work and information on the quality of the solution, while -->
<!-- for statistical purposes measures of uncertainty in the parameters are helpful. -->

<!-- Some tools allow constraints to be imposed on the solution, and it is helpful -->
<!-- to know if these constraints are active in the proposed solution. -->

<!-- ### Documentation -->

<!-- We believe that the availability of good documentation with examples that users -->
<!-- can modify to deal with their problems is essential. The presence of vignettes,  -->
<!-- demos, and tests indicates a package is well-supported. -->