### 2023-05-17 JNABsub20230515/ComparingNLStools.tex
add submission may 23

% !TeX root = RJwrapper.tex
\title{A Comparison of R Tools for Nonlinear Least Squares Modeling}
\author{by John C. Nash and Arkajyoti Bhattacharjee}

\maketitle

\abstract{%
Our Google Summer of Code project ``Improvements to \texttt{nls()}'' considered the features and limitations of the R function \texttt{nls()} with the aim of improving and rationalizing R tools for nonlinear regression. The rich features of \texttt{nls()} are weakened by several deficiencies and inconsistencies such as a lack of stabilization of the Gauss-Newton solver. Further considerations are the usability and maintainability of the code base that provides the functionality \texttt{nls()} claims to offer. Various packages, including our \texttt{nlsr}, provide alternative capabilities. We consider the differences in goals, approaches, and features of different tools for nonlinear least squares modeling in R. Discussion of these matters is relevant to improving R generally as well as its nonlinear estimation tools.
}

\hypertarget{the-nls-function}{%
\subsection{\texorpdfstring{The \texttt{nls()} function}{The nls() function}}\label{the-nls-function}}

\texttt{nls()} is the tool for estimating nonlinear statistical models in base R, the
primary software distribution
from the Comprehensive R Archive Network (\emph{\url{https://cran.r-project.org}}).
The function dates to the 1980s and the work related to Bates and Watts (1988) in S (see
\emph{\url{https://en.wikipedia.org/wiki/S_\%28programming_language\%29}}).

The \texttt{nls()} function has a remarkable and comprehensive set of capabilities for
estimating nonlinear models that are expressed as \textbf{formulas}. In particular, we
note that it

\begin{itemize}
\tightlist
\item
  handles formulas that include R functions, even ones that call calculations in
  other programming languages
\item
  allows data to be weighted or subset
\item
  can estimate bound-constrained parameters
\item
  provides a mechanism for handling partially linear models
\item
  permits parameters to be indexed over a set of related data
\item
  produces measures of variability (i.e., standard error estimates) for the
  estimated parameters
\item
  has related profiling capabilities for exploring the likelihood surface as
  parameters are changed
\item
  links to many pre-coded (\texttt{selfStart}) models that do not require
  initial parameter values.
\end{itemize}

With such a range of features and long history, the code has become untidy
and overly patched, difficult to maintain, and its underlying methods could
be improved. Various workers have developed packages to overcome these
concerns, and we will address some of these here.

\hypertarget{scope-of-our-comparison}{%
\subsection{Scope of our comparison}\label{scope-of-our-comparison}}

The tools we will consider are the base-R \texttt{nls()} function
and some packages that are available on the CRAN repository.
Of these, we will pay particular attention to \textbf{nlsr} (John C Nash and Duncan Murdoch (2023)),
\textbf{minpack.lm} (Elzhov et al. (2012)), and \textbf{gslnls} (Chau (2023)) which are
general nonlinear least-squares solvers.

While we will provide capsule comments for some other CRAN packages,
we will not pursue those in the Bioconductor (Gentleman et al. (2004))
collection, nor those on repositories such as
GitHub (\url{https://github.com}) and Gitlab (\url{https://about.gitlab.com}).

\hypertarget{some-other-cran-packages-for-nonlinear-modeling}{%
\subsection{Some other CRAN packages for nonlinear modeling}\label{some-other-cran-packages-for-nonlinear-modeling}}

\texttt{onls} (Andrej-Nikolai Spiess (2022)) -- The usual optimization in estimating nonlinear models
is the vertical difference between the dependent variable and the functional
model. \texttt{onls} minimizes the sum of \textbf{orthogonal} residuals. The objective is
therefore different and involves nontrivial extra calculation. A vignette with
the package and the blog article
\url{https://www.r-bloggers.com/2015/01/introducing-orthogonal-nonlinear-least-squares-regression-in-r/}
give some description with illustrative graphs. \texttt{onls} appears to be limited to
problems with one independent and one dependent variable. The Wikipedia article
\url{https://en.wikipedia.org/wiki/Total_least_squares} presents an overview of some
ideas, with references to the literature. The approach needs a wider
discussion and tutorial examples to allow its merits to be judged than can be
included here.

\texttt{crsnls} (Tvrdík (2016)) -- This package allows nonlinear estimation by
controlled random search via two methods. There is unfortunately no vignette.
A modest trial we carried out showed \texttt{nlsr::nlxb()} gave the same results in
a small fraction of the time required by either of the methods in \texttt{crsnls}. The
method discussed in Josef Tvrdík and Ivan Křivý and Ladislav Mišík (2007) claims better reliability in finding solutions
than a Levenberg-Marquardt code (actually from Matlab), but the tests were conducted
on the extreme NIST examples mentioned next.

\texttt{NISTnls} (National Institutes for Standards and R port by Douglas Bates (2012)) -- This package provides R code and data for a set of
(numerically ill-conditioned) nonlinear least squares problems from the U.S. National
Institute for Standards and Technology. These may not represent real-world situations.

\texttt{nlshelper} (Duursma (2017)) -- This package, which unfortunately lacks a
vignette, provides a few utilities for summarizing, testing, and plotting
non-linear regression models estimated with nls(), nlsList() or nlme() that
are linked or grouped in some way.

\texttt{nlsic} (Sokol (2022)) -- This solves nonlinear least squares problems with optional
equality and/or inequality constraints. It is clearly \textbf{not} about modeling, and
the input and output are quite different from class \texttt{nls} methods. However, there do
not appear to be other R packages with these capabilities.

\texttt{nlsMicrobio} (Baty and Delignette-Muller (2014)) -- Data sets and nonlinear regression models
dedicated to predictive microbiology, including a vignette, by authors of the
\texttt{nlstools} package.

\texttt{nlstools} (Baty and Delignette-Muller (2013)) -- This package provides several tools for aiding the estimation
of nonlinear models, particularly using \texttt{nls()}. The vignette is actually a journal
article, and the authors have considerable experience in the subject.

\texttt{nlsmsn} (Prates, Lachos, and Garay (2021)) -- Fit univariate non-linear scale mixture of skew-normal(NL-SMSN)
regression, with details in Garay, Lachos, and Abanto-Valle (2011). The problem
here is to minimize an objective that is modified from the traditional sum of squared
residuals.

\texttt{nls.multstart} (Padfield and Matheson (2020)) -- Non-linear least squares regression using
AIC scores with the Levenberg-Marquardt algorithm using multiple starting values
for increasing the chance that the minimum found is the global minimum.

\texttt{nls2} (Grothendieck (2022)) -- Nonlinear least squares by brute force has similar
motivations to \texttt{nls.multstart}, but uses \texttt{nls()} within multiple trials. The author
has extensive expertise in R.

\texttt{nlstac} (Rodriguez-Arias et al. (2020)) -- A set of functions implementing the algorithm described in
Torvisco, Fernandez, and Sanchez (2018) for fitting separable nonlinear regression curves. The
special class of problem for which this package is intended is an important and
difficult one. No vignette is provided, unfortunately.

\texttt{easynls} -- Fit and plot some nonlinear models. Thirteen models are treated, but
there is minimal documentation and no vignette. Package \texttt{nlraa} is to be preferred.

\texttt{nlraa} (Miguez (2021)) -- a set of nonlinear \emph{selfStart} models, primarily from
agriculture. Most include analytic Jacobian code.

\texttt{optimx} (John C. Nash and Varadhan (2011)) -- This provides
optimizers that can be applied to minimize a nonlinear function which could be
a nonlinear sum of
squares. NOT generally recommended if nonlinear
least squares programs can be easily used, but provides a check and alternative
solvers.

\hypertarget{an-illustrative-example}{%
\subsection{An illustrative example}\label{an-illustrative-example}}

The Hobbs weed infestation problem (John C. Nash (1979, 120)) is a growth curve modeling
task that seems straightforward but is quite nasty. Its very succinct statement
provides the ``short reproducible example'' much requested on R mailing lists.
This is a real problem from a field researcher. The data and graph follow.

\begin{verbatim}
weed <- c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443,
          38.558, 50.156, 62.948, 75.995, 91.972)
tt <- 1:12
weeddf <- data.frame(tt, weed)
plot(weeddf, main="Hobbs weed infestation data")
\end{verbatim}

\includegraphics{ComparingNLStools_files/figure-latex/ex01-1.pdf}

Three suggested models for this data are (with names to allow for easy reference)

\textbf{Logistic3U}:
\[  y \approx  b_1 / (1 + b_2 * exp(- b_3 * t)) \]

\textbf{Logistic3S}:
\[ y \approx  100 * c_1 / (1 + 10 * c_2 * exp(- 0.1 * c_3 * t)) \]

\textbf{Logistic3T}:
\[ y \approx Asym / (1 + exp((xmid - t)/scal)) \]

where we will use \texttt{weed} for \(y\) and \texttt{tt} for \(t\).
The functions above are equivalent, but the first is generally more awkward
to solve numerically due to its poor scaling. The parameters of the three forms
are related as follows:

\[   Asym =  b_1 = 100 * c_1 \]
\[ exp(xmid/scal)  =  b_2 = 10 * c_2 \]
\[ 1/scal  =  b_3 \]
To allow for simpler discussion, let us say that the parameters form a (named) vector
\(p\) and the model function is called \(model(p)\). The residuals can be written
either as \(r_1 = y - model(p)\) or \(r = model(p) - y\)
since their sum of squares has an identical value. The second form
avoids a potential sign error if we need to evaluate derivatives.

We wish to minimize the sum of squared residuals, which is our loss (or objective) function.
Starting with some guess for the parameters, we aim to alter these parameters to obtain a smaller
loss function. We then iterate until we can make no further progress.

Let us consider there are \(n\) parameters and \(m\) residuals.
The loss function is
\[ S(p) = r' r = \sum_{i=1}^m { r_i^2 }\]
The gradient of \(S(p)\) is
\[ 2 * J' r\]
where the Jacobian \(J\) is given by elements
\[ J_{i,j} = \partial r_i / \partial p_j \]
and the Hessian is defined by elements
\[ H_{i,j} = \partial ^2 S(p) / {\partial p_i \partial p_j} \]
If we expand the Hessian for nonlinear least squares problems, we find
\[  H_{i,j} = \sum_{k = 1}^m{ J_{k,i} J_{k,j}}  + \sum_{k = 1}^m {r_k * \partial{r_k}/{\partial p_i \partial p_j}}\]
Let us use \(D_{i,j}\) for the elements of the second term of this expression.
What is generally called \textbf{Newton's method} for function minimization tries
to set the gradient to zero (to find a stationary point of the function \(S(p)\)).
This leads to \textbf{Newton's equation}
\[ H \delta = -g \]
Given a set of parameters \(p\), we solve this equation for \(\delta\), adjust \(p\)
to \(p + \delta\) and iterate, hopefully to converge on a solution.
Applying this to a sum of squares problem gives
\[ H \delta = (J'J + D) \delta = - J' r\]
In this expression, only the elements of \(D\) have second partial derivatives.
Gauss, attempting to model planetary orbits, had small residuals, and noted
that these multiplied the second partial derivatives of \(r\), so he approximated
\[ H \approx J' J\]
by assuming \(D \approx 0\). This results in the Gauss-Newton method where we
solve
\[ J' J \delta = - J' r\]
though we can avoid some loss of accuracy by NOT forming the inner product matrix
\(J' J\) and solving the linear least squares matrix problem
\[ J \delta \approx -r \]
by one of several matrix decomposition methods.

In reality, there are many problems were \(D\) should not be ignored, but the work
to compute it precisely is considerable. Many work-arounds have been
proposed, of which the Levenberg-Marquardt stabilization
(Levenberg (1944), Marquardt (1963)) is
the most commonly used. For convenience, we will use ``Marquardt'', as we believe he
first incorporated the ideas into a practical computer program.

The usual suggestion is that \(D\) be replaced by a multiple of the unit matrix or else
a multiple of the diagonal part of \(J' J\). In low precision, some elements of
\(J' J\) could underflow to zero (John C. Nash (1977)), and a linear combination of both
choices is an effective compromise. Various choices for
\(D\), as well as a possible line search along the
direction \(\delta\) rather than a unit step (Hartley (1961)),
give rise to several variant
algorithms. ``Marquardt's method'' is a family of methods. Fortunately, most choices
work well.

\hypertarget{problem-setup}{%
\subsubsection{Problem setup}\label{problem-setup}}

Let us specify in R the three model formulas and set some starting
values for parameters. These starting points are NOT equivalent and
are deliberately crude choices. Workers performing many calculations
of a similar nature should try to provide good starting points to reduce
computation time and avoid finding a false solution.

\begin{verbatim}
# model formulas
frmu <- weed ~ b1/(1+b2*exp(-b3*tt))
frms <- weed ~ 100*c1/(1+10*c2*exp(-0.1*c3*tt))
frmt <- weed ~ Asym /(1 + exp((xmid-tt)/scal))
#
# Starting parameter sets
stu1<-c(b1=1, b2=1, b3=1)
sts1<-c(c1=1, c2=1, c3=1)
stt1<-c(Asym=1, xmid=1, scal=1)
\end{verbatim}

One of the useful features of \texttt{nls()} is the possibility of a \texttt{selfStart} model,
where starting parameter values are not required. However, if a \texttt{selfStart} model
is not available, \texttt{nls()} sets all the starting parameters
to 1. This is tolerable, but could be improved by using a set of values
that are all slightly different, which, in the case of the example
model \(y \,\sim\, a \,*\, exp(-b \,*\, x) + c\,*\,exp(-d \,*\, x)\)
would avoid a singular Jacobian because \(b\) and \(d\) were equal in value.
Program modifications to give a sequence like 1.0, 1.1, 1.2, 1.3 for the four
parameters are fairly obvious.

It is also possible to provide R functions for the residual and Jacobian.
This is usually more work for the user if the formula setup is possible.
To illustrate, we show the functions for the unscaled 3 parameter logistic.

\begin{verbatim}
# Logistic3U
hobbs.res  <-  function(x){ # scaled Hobbs weeds problem -- residual
  # This variant uses looping
  if(length(x) != 3) stop("hobbs.res -- parameter vector n!=3")
  y  <-  c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443, 
           38.558, 50.156, 62.948, 75.995, 91.972)
  tt  <-  1:12
  res  <-  x[1]/(1+x[2]*exp(-x[3]*tt)) - y
}

hobbs.jac  <-  function(x) { # scaled Hobbs weeds problem -- Jacobian
  jj  <-  matrix(0.0, 12, 3)
  tt  <-  1:12
  yy  <-  exp(-x[3]*tt)
  zz  <-  1.0/(1+x[2]*yy)
  jj[tt,1]   <-   zz
  jj[tt,2]   <-   -x[1]*zz*zz*yy
  jj[tt,3]   <-   x[1]*zz*zz*yy*x[2]*tt
  attr(jj, "gradient") <- jj
  jj
}
\end{verbatim}

\hypertarget{estimation-of-models-specified-as-formulas}{%
\subsection{Estimation of models specified as formulas}\label{estimation-of-models-specified-as-formulas}}

Using a formula specification was a principal advantage made
with \texttt{nls()} when it became available in S sometime in the 1980s. It uses a
Gauss-Newton (i.e., unstabilized) iteration with a step reduction line
search. This works very efficiently as long as \(J\) is not ill-conditioned.
Below we see \texttt{nls()} does poorly on the example problem. To save page space,
we use 1-line result display functions from package \texttt{nlsr}, namely \texttt{pnls()}
and \texttt{pshort()}.

\begin{verbatim}
#> Error in nls(formula = frmu, start = stu1, data = weeddf) : 
#>   singular gradient
\end{verbatim}

\begin{verbatim}
#> Error in nls(formula = frms, start = sts1, data = weeddf) : 
#>   singular gradient
\end{verbatim}

\begin{verbatim}
#> Error in nls(formula = frmt, start = stt1, data = weeddf) : 
#>   singular gradient
\end{verbatim}

Here we see the infamous ``singular gradient'' termination message of \texttt{nls()}.

\hypertarget{solution-attempts-with-nlsr}{%
\subsubsection{Solution attempts with nlsr}\label{solution-attempts-with-nlsr}}

\begin{verbatim}
#> residual sumsquares =  2.5873  on  12 observations
#>     after  19    Jacobian and  25 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> b1               196.186         11.31      17.35  3.167e-08  -4.859e-09        1011  
#> b2               49.0916         1.688      29.08  3.284e-10  -3.099e-08      0.4605  
#> b3               0.31357      0.006863      45.69  5.768e-12   2.305e-06     0.04714
\end{verbatim}

\begin{verbatim}
#> snlx1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 34 res/ 23 jac
\end{verbatim}

\begin{verbatim}
#> tnlx1  -- ss= 2.5873 :  Asym = 196.19  xmid = 12.417  scal = 3.1891; 36 res/ 27 jac
\end{verbatim}

Though we have found solutions, the Jacobian is essentially singular as
shown by its singular
values. Note that these are \textbf{displayed} by package \texttt{nlsr} in a single column in
the output to provide a compact layout, but the values do \textbf{NOT} correspond to
the individual parameters in
whose row they appear; they are a property of the whole problem.

\hypertarget{solution-attempts-with-minpack.lm}{%
\subsubsection{Solution attempts with minpack.lm}\label{solution-attempts-with-minpack.lm}}

\begin{verbatim}
#> unlm1  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 17  itns
\end{verbatim}

\begin{verbatim}
#> snlm1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 7  itns
\end{verbatim}

\begin{verbatim}
#> tnlm1  -- ss= 9205.4 :  Asym = 35.532  xmid = 43376  scal = -2935.4; 39  itns
\end{verbatim}

\hypertarget{solution-attempts-with-gslnls}{%
\subsubsection{Solution attempts with gslnls}\label{solution-attempts-with-gslnls}}

\begin{verbatim}
#> ugslnls1  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\begin{verbatim}
#> sgslnls1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 9  itns
\end{verbatim}

\begin{verbatim}
#> tgslnls1  -- ss= 9205.4 :  Asym = 35.532  xmid = 20846  scal = -1741.1; 47  itns
\end{verbatim}

\hypertarget{comparison-notes-for-formula-setup-solutions}{%
\subsubsection{Comparison notes for formula-setup solutions}\label{comparison-notes-for-formula-setup-solutions}}

\texttt{nlsr::nlxb()} uses \texttt{print()} to output standard errors and
singular values of the Jacobian (for diagnostic purposes). By contrast,
\texttt{minpack.lm::nlsLM()} and \texttt{nls()}
use \texttt{summary()}, which does NOT display the sum of squares, while \texttt{print()}
gives the sum of squares, but not the standard error of the residuals.

The singular values allow us to gauge how ``nearly singular'' the
Jacobian is at the solution, and the ratio of the smallest to largest of the
singular values is a simple but effective measure. The ratios are
4.6641e-05 for Logistic3U, 0.021022 for Logistic3S, and 0.001055 for
Logistic3T, so Logistic3S is the ``least singular''.

The results from \texttt{nlsLM} and \texttt{gsl\_nls} for the transformed model Logistic3T have a very
large sum of squares, which may suggest that these programs have failed.
Since \texttt{nls()}, \texttt{nlsLM()}, and \texttt{gsl\_nls()} do not
offer singular values, we need to extract the Jacobian and compute its
singular values. The following script shows how to do this, using as Jacobian what is called
the \texttt{gradient} element in the returned solution for these solvers.

\begin{verbatim}
# for nlsLM
if (inherits(tnlm1, "try-error"))  {
   print("Cannot compute solution -- likely singular Jacobian")
 } else {  
   JtnlsLM <- tnlm1$m$gradient() # actually the Jacobian
   svd(JtnlsLM)$d # Singular values
}   
\end{verbatim}

\begin{verbatim}
#> [1] 3.4641e+00 3.2530e-10 8.9643e-12
\end{verbatim}

\begin{verbatim}
# for gsl_nls
if (inherits(tgslnls1, "try-error")) {
   cat("Cannot compute solution -- likely singular Jacobian")
} else {  
   JtnlsLM <- tgslnls1$m$gradient()
   svd(JtnlsLM)$d # Singular values
}   
\end{verbatim}

\begin{verbatim}
#> [1] 3.4641e+00 9.4495e-09 4.1324e-11
\end{verbatim}

We see that there are differences in detail, but the more important result is that
two out of three singular values are essentially 0. Our Jacobian is singular, and no
method of the Gauss-Newton type should be able to continue. Indeed, from the parameters
reported at this saddle point, \texttt{nlsr::nlxb()} cannot proceed.

\begin{verbatim}
stspecial<- c(Asym = 35.532,  xmid = 43376,  scal = -2935.4)
badstart<-nlxb(formula=frmt, start=stspecial, data=weeddf)
print(badstart)
\end{verbatim}

\begin{verbatim}
#> residual sumsquares =  9205.4  on  12 observations
#>     after  2    Jacobian and  2 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> Asym             35.5321            NA         NA         NA  -9.694e-09       3.464  
#> xmid               43376            NA         NA         NA  -1.742e-09    2.61e-10  
#> scal             -2935.4            NA         NA         NA    -2.4e-08    7.12e-16
\end{verbatim}

\hypertarget{functional-specification-of-problems}{%
\subsection{Functional specification of problems}\label{functional-specification-of-problems}}

We illustrate how to solve nonlinear least squares problems using
a function to define the residual. Note that \texttt{gsl\_nls()} requires
a vector \texttt{y} that is the expected value of what we have called the
residual, but in this case, is actually the model. \texttt{gsl\_nls} uses
a numerical approximation for the Jacobian if the argument \texttt{jac}
is missing. Note function \texttt{nlsr::pnlslm()} for a 1-line display of
the results of \texttt{minpack.lm::nls.lm()}.

\begin{verbatim}
#> hobnlfb  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25 res/ 19 jac
\end{verbatim}

\begin{verbatim}
#> hobnlm  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 17  itns
\end{verbatim}

\begin{verbatim}
#> hobgsln  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\begin{verbatim}
#> hobgsl  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\hypertarget{design-goals-termination-tests-and-output-objects}{%
\subsection{Design goals, termination tests, and output objects}\label{design-goals-termination-tests-and-output-objects}}

The output object of \texttt{nlxb()} is smaller than the class \texttt{nls} object returned
by \texttt{nls()}, \texttt{nlsLM()}, and \texttt{gsl\_nls()}. Package \texttt{nlsr} emphasizes the solution
of the nonlinear least squares problem rather than the estimation of a nonlinear
model that fits or explains the data. The object of class \texttt{nls} allows for
a number of specialized modeling and diagnostic extensions. For compatibility,
the \texttt{nlsr} package has the function \texttt{wrapnlsr()}, for which \texttt{nlsr()} is an alias.
This uses \texttt{nlxb()} to find good parameters, then calls \texttt{nls()} to return the class
\texttt{nls} object. Unless particular modeling features are needed, the use of
\texttt{wrapnlsr()} is unnecessary and wasteful of resources.

The design goals of the different tools may also be revealed in the so-called
``convergence tests'' for the iterative solvers. In the manual page for \texttt{nls()}
in R 4.0.0 there was the warning:

\begin{quote}
\emph{Do not use \texttt{nls} on artificial ``zero-residual'' data.}
\end{quote}

with suggested addition of small perturbations to the data. This admits
\texttt{nls()} could not solve well-posed problems unless data is polluted
with errors. Zero-residual problems are not always artificial,
since problems in function approximation and nonlinear equations can be
approached with nonlinear least squares. Fortunately, a small adjustment
to the ``termination test'' for the \textbf{program}, rather than for the
``convergence'' of the underlying \textbf{algorithm}, fixes the defect. The test
is the Relative Offset Convergence
Criterion (see Bates, Douglas M. and Watts, Donald G. (1981)). This scales an estimated reduction in the loss
function by its current value. If the loss function is very small, we are close
to a zero-divide. Adding a small quantity to the divisor avoids trouble.
In 2021, one of us (J. Nash) proposed that \texttt{nls.control()} have an
additional parameter \texttt{scaleOffset} with a default value of zero. Setting it to a
small number -- 1.0 is a reasonable choice -- allows small-residual problems
(i.e., near-exact fits) to be dealt with easily. We call this the
\textbf{safeguarded relative offset convergence criterion}, and it has been in \texttt{nlsr}
since it was introduced. The default value
gives the legacy behavior. This improvement has been in the R distributed code
since version 4.1.0.

Additional termination tests can be used. \texttt{nlsr} has a \textbf{small sum of
squares} test (\textbf{smallsstest}) that compares the latest evaluated sum of squared
(weighted) residuals to \texttt{e4} times the initial sum of squares, where
\texttt{e4\ \textless{}-\ (100*.Machine\$double.eps)\^{}4} is approximately 2.43e-55.
\%\% \^{} is picked up in rjtools check

Termination after what may be considered excessive computation is also important.
\texttt{nls()} stops after \texttt{maxiter} ``iterations''. The meaning of ``iteration'' may require an
examination of the code. We terminate execution when the number of residual or
Jacobian evaluations exceed set limits. Generally, we prefer larger limits than the
default \texttt{maxiter\ =\ 50} of \texttt{nls()} to avoid stopping early.

\hypertarget{returned-results-of-nls-and-other-tools}{%
\subsubsection{Returned results of nls() and other tools}\label{returned-results-of-nls-and-other-tools}}

As mentioned, the output of \texttt{nls()} is an object of class ``nls'' which has a quite rich structure
described in the manual file or revealed by applying the \texttt{str()} function to the
result of \texttt{nls()}.

The complexity of this object is a challenge to users. Let us use the
\texttt{result} as the returned object from \texttt{nls()}, \texttt{minpack.lm::nlsLM()} or,
\texttt{gslnls::gsl\_nls()}. For example, the \texttt{data} return element is an R symbol. To
actually access the data from this element, we need to use the syntax:

\begin{verbatim}
eval(parse(text=result$data))
\end{verbatim}

However, if the call is made with \texttt{model=TRUE}, then there is a returned element
\texttt{model} which contains the data, and we can list its contents using:

\begin{verbatim}
ls(result$model)
\end{verbatim}

and if there is an element called \texttt{xdata}, then it can be accessed as
\texttt{result\$model\$xdata}.

By contrast, \texttt{nlsr::nlxb()} returns a much simpler structure of 11 items
in one level. Moreover, \texttt{nlxb} explicitly returns the sum of squares, the residual vector, Jacobian, and counts of evaluations.

\hypertarget{when-to-compute-ancillary-information}{%
\subsubsection{When to compute ancillary information}\label{when-to-compute-ancillary-information}}

Tools that produce a class \texttt{nls} output object create a rich set of functions
and structures that are then used in a variety of modeling tasks, including the
least squares solution. By contrast,
\texttt{nlsr} computes quantities as they are requested or needed, with additional
features in separate functions. For example, the singular values of the Jacobian
are actually computed in the \texttt{print} and \texttt{summary} methods for the result.
These two approaches
lead to different consequences for performance and how features are
provided. \texttt{nlsr} has antecedents in the methods of John C. Nash (1979), where storage for
data and programs was at a ridiculous premium in the small computers of the era.
Thus the code in \texttt{nlsr} is likely of value for workers to copy and modify
for customized tools.

\hypertarget{jacobian-calculation}{%
\subsection{Jacobian calculation}\label{jacobian-calculation}}

Gauss-Newton/Marquardt methods all need a Jacobian matrix at each iteration.
By default, \texttt{nlsr::nlxb()} will try to evaluate this using analytic
expressions using symbolic and automatic differentiation tools.
When using a formula specification of the model, \texttt{nls()}, \texttt{minpack.lm::nlsLM()}
and \texttt{gslnls::gsl\_nls()} use a finite difference approximation to compute the
Jacobian, though \texttt{gsl\_nls()} does have an option to attempt symbolic expressions.
Package \texttt{nlsr} provides, via appropriate calling syntax, four numeric approximation
options for the Jacobian, with a further control \texttt{ndstep} for the size of the
step used in the approximation.

Using the ``gradient'' attribute of the output of the Jacobian function to hold
the Jacobian matrix lets us embed this in the \textbf{residual} function as well,
so that the call to \texttt{nlsr::nlfb()} can be made with the same name used for
both residual and Jacobian function arguments. This programming trick saves a
lot of trouble for the package developer, but it can be a nuisance for users
trying to understand the code.

As far as we can understand the logic in \texttt{nls()}, the Jacobian computation during
parameter estimation is carried out within the called C-language program
and its wrapper R code function \texttt{numericDeriv()}, part of
\texttt{./src/library/stats/R/nls.R}
in the R distribution source code. This is used to provide Jacobian information in
the \texttt{nlsModel()} and \texttt{nlsModel.plinear()} functions, which are \textbf{not} exported for
general use. \texttt{gsl\_nls()} also appears to use \texttt{numericDeriv()}.

\texttt{numericDeriv()} uses a
simple forward difference approximation of derivatives, though a central
difference approximation can be specified in control parameters.
We are unclear why \texttt{numericDeriv()} in base R calls \texttt{C\_numeric\_deriv},
as we were easily able to create a more compact version entirely in R.
See \emph{\url{https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS}}.

\texttt{minpack.lm::nlsLM()} invokes \texttt{numericDeriv()} in its local
version of \texttt{nlsModel()}, but it appears to use an internal approximate Jacobian
code from
the original Fortran \texttt{minpack} code, namely, \texttt{lmdif.f}. Such differences in approach
can lead to different behavior, usually minor, but sometimes annoying with
ill-conditioned problems.

\begin{itemize}
\item
  A pasture regrowth problem (Huet et al. (2004), page 1, based on Ratkowsky (1983)) has a
  poorly conditioned Jacobian and \texttt{nls()} fails with ``singular gradient''.
  Worse, numerical approximation to the Jacobian gives the error
  ``singular gradient matrix at initial parameter estimates'' for \texttt{minpack.lm::nlsLM}
  so that the Marquardt stabilization is unable to take effect, while the analytic
  derivatives of \texttt{nlsr::nlxb} give a solution.
\item
  Karl Schilling (private communication) provided an example where a model specified
  with the formula \texttt{y\ \textasciitilde{}\ a\ *\ (x\ ˆ\ b)} causes \texttt{nlsr::nlxb} to fail because the partial
  derivative w.r.t. \texttt{b} is \texttt{a\ *\ (x\^{}b\ *\ log(x))}. If there is data for which \texttt{x\ =\ 0},
  the derivative is undefined, but the model can be computed. In such cases,
  we observed that \texttt{nls()} and
  \texttt{minpack.lm::nlsLM} found a solution, though this seems to be a lucky accident.
\end{itemize}

\hypertarget{jacobian-code-in-selfstart-models}{%
\subsubsection{Jacobian code in selfStart models}\label{jacobian-code-in-selfstart-models}}

Analytic Jacobian code can be provided to all the solvers discussed.
Most \texttt{selfStart} models that automatically provide starting parameters
also include such code. There is documentation in R
of \texttt{selfStart} models, but their construction is non-trivial. A number
of such models are included with base R in \texttt{./src/library/stats/R/zzModels.R},
with package \texttt{nlraa} (Miguez (2021)) providing a richer set.
There are also some in the now-archived package \texttt{NRAIA}.
These provide the Jacobian in the ``gradient'' attribute of the ``one-sided'' formula
that defines each model, and these Jacobians are often the analytic forms.

The \texttt{nls()} function, after
computing the ``right-hand side'' or \texttt{rhs} of the residual, checks to see if the
``gradient'' attribute is defined, otherwise using \texttt{numericDeriv()} to compute a
Jacobian into that attribute. This code is within the \texttt{nlsModel()} or
\texttt{nlsModel.plinear()} functions. The use of analytic Jacobians
almost certainly contributes to the good performance of \texttt{nls()} on \texttt{selfStart}
models.

The use of \texttt{selfStart} models with
\texttt{nlsr} is described in the ``Introduction to nlsr'' vignette. However, since \texttt{nlsr}
generally can use very crude starting values, we have rarely needed them, though
it should be pointed out that our work is primarily diagnostic. If we were carrying
out a large number of similar estimations, such initial parameters are
critical to efficiency.

In considering \texttt{selfStart} models, we noted that the base-R function \texttt{SSlogis}
is intended to solve problem \textbf{Logistic3T} above. When this function is used
via \texttt{getInitial()} to find
starting values, it actually calls \texttt{nls()} with the `plinear' algorithm
and finds a (full) solution. It then passes the solution coefficients to the default
algorithm unnecessarily. Moreover, the implicit double call is, in our view, prone to creating
errors in code maintenance.
To provide simpler starting parameters, the function \texttt{SSlogisJN}
is now part of the package \texttt{nlsr}, but is most useful for \texttt{nls()}.

Users may also want to profit from the Jacobian code of \texttt{selfStart} models
but supply explicit starting values other than those suggested by \texttt{getInitial()}.
This does not appear to be possible with \texttt{nls()}. \texttt{nlsr::nlxb()} always requires
starting parameters, and can either use \texttt{getInitial()} to find them from the \texttt{selfStart}
model or provide explicit values, but the \texttt{selfStart} model is used in the formula
for the \texttt{nlxb()} call.

We are also surprised that the
analytic expressions for the Jacobian (``gradient'') in the \texttt{SSLogis} function
and others save quantities in ``hidden''
variables, i.e., with names preceded by ``.''.
These are then not displayed by the \texttt{ls()}
command, making them difficult to access by users who may wish to create
their own selfStart model via copy and edit.
Interactive tools, such as ``visual fitting'' (John C. Nash and Velleman (1996)) might be
worth considering as another way to find starting parameters, but we know of no
R capability of this type.

As a side note, the introduction of \texttt{scaleOffset} in R 4.1.1 to deal with the
convergence test for small residual problems now requires that the \texttt{getInitial()}
function have dot-arguments (\texttt{...}) in its argument list. This illustrates the
entanglement of many features in \texttt{nls()} that complicate its maintenance and
improvement.

\hypertarget{bounds-constraints-on-parameters}{%
\subsection{Bounds constraints on parameters}\label{bounds-constraints-on-parameters}}

For many problems, we know that parameters cannot be bigger or smaller than some
externally known limits. Such limits should be built into models, but there
are some important details for using the tools in R.

\begin{itemize}
\item
  \texttt{nls()} can only impose bounds if the \texttt{algorithm="port"} argument is
  used in the call. Unfortunately, the documentation warns us:

  \emph{The algorithm = ``port'' code appears unfinished, and does not even
  check that the starting value is within the bounds. Use with
  caution, especially where bounds are supplied.}
\item
  \texttt{gsl\_nls()} does not offer bounds.
\item
  bounds are part of the default method for package \texttt{nlsr}.
\item
  \texttt{nlsLM()} includes bounds in the standard call, but we have observed cases
  where it fails to get the correct answer. From an examination of the code,
  we believe the authors have not taken into account all possibilities, though
  all programs have some weakness regarding constrained optimization in that
  programmers have to work with assumptions on the scale of numbers,
  and some problems will be outside the scope envisaged.
\end{itemize}

\begin{verbatim}
# Start MUST be feasible i.e. on or within bounds
anlshob1b <- nls(frms, start=sts1, data=weeddf, lower=c(0,0,0),
             upper=c(2,6,3), algorithm='port')
pnls(anlshob1b) #  check the answer (short form)
\end{verbatim}

\begin{verbatim}
#> anlshob1b  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 10  itns
\end{verbatim}

\begin{verbatim}
# nlsLM seems NOT to work with bounds in this example
anlsLM1b <- nlsLM(frms, start=sts1, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pnls(anlsLM1b)
\end{verbatim}

\begin{verbatim}
#> anlsLM1b  -- ss= 881.02 :  c1 = 2  c2 = 6  c3 = 3; 2  itns
\end{verbatim}

\begin{verbatim}
# also no warning if starting out of bounds, but gets a good answer!!
st4<-c(c1=4, c2=4, c3=4)
anlsLMob <- nlsLM(frms, start=st4, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pnls(anlsLMob)
\end{verbatim}

\begin{verbatim}
#> anlsLMob  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 4  itns
\end{verbatim}

\begin{verbatim}
# Try nlsr::nlxb()
anlx1b <- nlxb(frms, start=sts1, data=weeddf, lower=c(0,0,0), upper=c(2,6,3))
pshort(anlx1b)
\end{verbatim}

\begin{verbatim}
#> anlx1b  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 12 res/ 12 jac
\end{verbatim}

\hypertarget{philosophical-considerations}{%
\subsubsection{Philosophical considerations}\label{philosophical-considerations}}

Bounds on parameters raise some interesting questions about how
uncertainty in parameter estimates should be computed or reported. That is, the
traditional ``standard errors'' are generally taken to imply symmetric intervals about
the point estimate in which the parameter may be expected to be found with some
probability under certain assumptions.
Bounds change those assumptions. Hence, \texttt{nlsr::nlxb()} does not compute standard
errors nor their derived statistics when bounds are active.

\hypertarget{fixed-parameters-masks}{%
\subsubsection{Fixed parameters (masks)}\label{fixed-parameters-masks}}

Let us try to fix (mask) the first parameter in the first two example problems.

\begin{verbatim}
# Hobbsmaskx.R -- masks with formula specification of the problem
require(nlsr); require(minpack.lm); traceval<-FALSE
stu <- c(b1=200, b2=50, b3=0.3) # a default starting vector (named!)
sts <- c(c1=2, c2=5, c3=3) # a default scaled starting vector (named!)
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
            upper=c(200, 60, 3), trace=traceval))
print(anxbmsk1)
\end{verbatim}

\begin{verbatim}
#> residual sumsquares =  2.6182  on  12 observations
#>     after  4    Jacobian and  4 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> b1                   200U M         NA         NA         NA           0          NA  
#> b2               49.5108          1.12      44.21  8.421e-13  -2.887e-07        1022  
#> b3              0.311461      0.002278      136.8  1.073e-17   0.0001635      0.4569
\end{verbatim}

\begin{verbatim}
anlM1 <- try(nlsLM(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
            upper=c(200, 60, 3), trace=traceval))
pnls(anlM1)
\end{verbatim}

\begin{verbatim}
#> anlM1  -- ss= 2.6182 :  b1 = 200  b2 = 49.511  b3 = 0.31146; 4  itns
\end{verbatim}

\begin{verbatim}
anlsmsk1 <- try(nls(frmu, start=stu, data=weeddf, lower=c(200,0,0), 
        upper=c(200, 60, 3),  algorithm="port", trace=traceval))
pnls(anlsmsk1)
\end{verbatim}

\begin{verbatim}
#> anlsmsk1  -- ss= 2.6182 :  b1 = 200  b2 = 49.511  b3 = 0.31146; 5  itns
\end{verbatim}

\begin{verbatim}
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- nlxb(frms, start=sts, data=weeddf, lower=c(2,0,0),
                  upper=c(2,6,30))
print(anlxmsks1)
\end{verbatim}

\begin{verbatim}
#> residual sumsquares =  2.6182  on  12 observations
#>     after  4    Jacobian and  4 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> c1                     2U M         NA         NA         NA           0          NA  
#> c2               4.95108         0.112      44.21  8.421e-13  -2.981e-06       104.2  
#> c3               3.11461       0.02278      136.8  1.073e-17   1.583e-05       4.482
\end{verbatim}

\begin{verbatim}
anlshmsk1 <- nls(frms, start=sts, trace=traceval, data=weeddf, lower=c(2,0,0),
             upper=c(2,6,30), algorithm='port')
pnls(anlshmsk1)
\end{verbatim}

\begin{verbatim}
#> anlshmsk1  -- ss= 2.6182 :  c1 = 2  c2 = 4.9511  c3 = 3.1146; 5  itns
\end{verbatim}

\begin{verbatim}
anlsLMmsks1 <- nlsLM(frms, start=sts, data=weeddf, lower=c(2,0,0),
                 upper=c(2,6,30))
pnls(anlsLMmsks1)
\end{verbatim}

\begin{verbatim}
#> anlsLMmsks1  -- ss= 2.6182 :  c1 = 2  c2 = 4.9511  c3 = 3.1146; 4  itns
\end{verbatim}

\begin{verbatim}
# Test with all parameters masked
anlxmskall<- try(nlxb(frms, start=sts, data=weeddf, lower=sts, upper=sts))
print(anlxmskall)
\end{verbatim}

\begin{verbatim}
#> residual sumsquares =  158.23  on  12 observations
#>     after  0    Jacobian and  1 function evaluations
#>   name            coeff          SE       tstat      pval      gradient    JSingval   
#> c1                     2U M         NA         NA         NA          NA          NA  
#> c2                     5U M         NA         NA         NA          NA          NA  
#> c3                     3U M         NA         NA         NA          NA          NA
\end{verbatim}

\texttt{nlsr} has an output format that indicates the constraint status of the parameter
estimates. For \texttt{nlsr}, we have \textbf{chosen} to suppress the calculation of
approximate standard errors in the parameters when constraints are active because
their meaning under constraints is unclear, though we believe this
policy worthy of discussion and further investigation.

\hypertarget{stabilization-of-gauss-newton-computations}{%
\subsection{Stabilization of Gauss-Newton computations}\label{stabilization-of-gauss-newton-computations}}

All four major tools illustrated solve some variant of the Gauss-Newton
equations. \texttt{nls()} uses a modification of an approach suggested by Hartley (1961),
while \texttt{nlsr}, \texttt{gslnls}, and \texttt{minpack.lm} use flavors of Marquardt (1963). \texttt{gslnls}
offers an accelerated Marquardt method and three alternative methods which we
have not investigated.
Control settings for \texttt{nlxb()} or \texttt{nlfb()} allow exploration of Hartley and Marquardt
algorithm variants. In general, the Levenberg-Marquardt stabilization
is important in obtaining
solutions in methods of the Gauss-Newton family, as \texttt{nls()} terminates
too frequently and unnecessarily with \texttt{singular\ gradient} errors.

\hypertarget{programming-language}{%
\subsubsection{Programming language}\label{programming-language}}

An important choice made in developing \texttt{nlsr} was to code entirely within the R
programming language. \texttt{nls()} uses a mix of R, C, and Fortran, as does \texttt{minpack.lm}.
\texttt{gslnls} is an R wrapper to various C-language routines in the Gnu Scientific
Library (Galassi et al. (2009)).
Generally, keeping to a single programming language can allow for
easier maintenance and upgrades. It also avoids some work when there are changes
or upgrades to libraries for the non-R languages.
R is usually considered slower
than most computing environments because it keeps track of objects and
because it is usually interpreted. In recent years, the
performance penalty for using code entirely in R has been much reduced
with the just-in-time compiler and other improvements. All-R computation
may now offer acceptable performance. In \texttt{nlsr}, the use of R may be a
smaller performance cost than the aggressive approach to a solution,
which can cause more iterations to be used.

\hypertarget{data-sources-for-problems}{%
\subsection{Data sources for problems}\label{data-sources-for-problems}}

\texttt{nls()} can be called without specifying the \texttt{data} argument. In this case, it will
search in the available environments (i.e., workspaces) for suitable data objects.
We do NOT like this approach, but it is ``the R way''. R allows users to leave many
objects in the default (\texttt{.GlobalEnv}) workspace. Moreover, users have to actively
suppress saving this workspace (\texttt{.RData}) on exit; otherwise, any such file in
the path will be loaded on startup. R users in our
acquaintance avoid saving the workspace because of lurking data and
functions that may cause unwanted results.

\hypertarget{feature-subsetting}{%
\subsubsection{Feature: Subsetting}\label{feature-subsetting}}

\texttt{nls()} and other class \texttt{nls} tools accept an argument \texttt{subset}. This acts through
the mediation of \texttt{model.frame}, which not obvious in the
source code files \texttt{/src/library/stats/R/nls.R} and \texttt{/src/library/stats/src/nls.C}.
Having \texttt{subset} at the level of the call to a function like
\texttt{nls()} saves effort, but it does mean that the programmer of the
solver needs to be aware of the
origin (and value) of objects such as the data, residuals and Jacobian.
By preference,
we would implement subsetting by zero-value weights, with observation counts
(and degrees of freedom) computed via the numbers of non-zero weights. Alternatively,
we would extract a working dataframe from the relevant elements in the original.

\hypertarget{feature-na.action-missing-value-treatment}{%
\subsubsection{Feature: na.action (missing value treatment)}\label{feature-na.action-missing-value-treatment}}

\texttt{na.action} is an argument to the \texttt{nls()} function, but it does not appear
obviously in the source code, often being handled behind the scenes after
referencing the option \texttt{na.action}. This feature also changes the data
supplied to our nonlinear least squares solver.
A useful, but possibly dated, description is given in:
\emph{\url{https://stats.idre.ucla.edu/r/faq/how-does-r-handle-missing-values/}}.
The typical default action, which can be seen by using the command
\texttt{getOption("na.action")}
is \texttt{na.omit}. This option removes from computations any observations
containing missing values (i.e.~any row of a data frame containing an NA).
\texttt{na.exclude} does much of the same for solver computations, but keeps the rows with
NA elements so that predictions are in the correct row position. We recommend that
workers actually test output to verify the behavior is as wanted.
See \emph{\url{https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r}}.
As with \texttt{subset}, our concern with \texttt{na.action} is that users may be unaware of the
effects of an option they may not even be aware has been set. Should \texttt{na.fail} be the default?

\hypertarget{feature-model-frame}{%
\subsection{Feature: model frame}\label{feature-model-frame}}

\texttt{model} is an argument to the \texttt{nls()} and related functions, which is documented as:

\begin{quote}
\textbf{model} logical. If true, the model frame is returned as part of the object. Default is FALSE.
\end{quote}

Indeed, the argument only gets used when \texttt{nls()} is about to return its result
object, and the
element \texttt{model} is NULL unless the calling argument \texttt{model} is TRUE. (Using the same
name could
be confusing.) Despite this, the model frame is used within the function code in
the form of the object
\texttt{mf}. We feel that users could benefit from more extensive documentation and
examples of its use since it is used to implement features like \texttt{subset}.

\hypertarget{weights-on-observations}{%
\subsection{Weights on observations}\label{weights-on-observations}}

All four main tools we consider here allow a \texttt{weights} argument that
specifies a vector of fixed weights the same length as the number of residuals.
Each residual is multiplied by the square root of the corresponding weight.
Where available, the values returned by the
\texttt{residuals()} function are weighted, and the \texttt{fitted()} or \texttt{predict()} function are
used to compute raw residuals.

While fixed weights may be useful, there are many problems for which
we want weights that are determined at least partially from the model
parameters, for
example, a measure of the standard deviation of observations.
Such dynamic weighting situations are discussed in the vignette
``Introduction to nlsr'' of package \texttt{nlsr} in section
\emph{Weights that are functions of the model parameters}.
\texttt{minpack.lm} offers a function \texttt{wfct()} to
facilitate such weighting. Care is advised
in applying such ideas.

\hypertarget{weights-in-returned-functions-from-nls}{%
\subsubsection{Weights in returned functions from nls()}\label{weights-in-returned-functions-from-nls}}

The function \texttt{resid()} (an alias for \texttt{residuals()}) gives WEIGHTED residuals,
as does, for example, \texttt{result\$m\$resid()}. The function \texttt{nlsModel()}, which we
have had to extract from the base R code and explicitly \texttt{source()} because it
is not exported to the working namespace,
allows us to compute residuals for particular coefficient sets.

\begin{verbatim}
wts <- 0.5^tt # simple weights
frmlogis <- weed ~ Asym/(1 + exp((xmid - tt)/scal))
Asym<-1; xmid<-1; scal<-1
nowt<-nls(weed ~ SSlogis(tt, Asym, xmid, scal)) # UNWEIGHTED
rnowt<-nowt$m$resid() # This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.
attr(rnowt, "gradient")<-NULL; rnowt
\end{verbatim}

\begin{verbatim}
#>  [1] -0.011900  0.032755 -0.092030 -0.208782 -0.392634  0.057594  1.105728
#>  [8] -0.715786  0.107647  0.348396 -0.652592  0.287569
\end{verbatim}

\begin{verbatim}
usewt <- nls(weed ~ SSlogis(tt, Asym, xmid, scal), weights=wts)
rusewt<-usewt$m$resid() # WEIGHTED. Does NOT take coefficients.
attr(rusewt,"gradient")<-NULL; rusewt
\end{verbatim}

\begin{verbatim}
#>  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
#>  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}

\begin{verbatim}
source("nlsModel.R")
nmod0 <- nlsModel(frmlogis, data=weeddf, start=c(Asym=1, xmid=1, scal=1), wts=wts)
rn0<-nmod0$resid() # Parameters are supplied in nlsModel() `start` above.
attr(rn0,"gradient")<-NULL; rn0
\end{verbatim}

\begin{verbatim}
#>  [1] 3.3998 3.2545 3.0961 2.9784 2.8438 2.7748 2.6910 2.3474 2.1724 1.9359
#> [11] 1.6572 1.4214
\end{verbatim}

\begin{verbatim}
nmod <- nlsModel(frmlogis, data=weeddf, start=coef(usewt), wts=wts)
rn<-nmod$resid()
attr(rn,"gradient")<-NULL; rn
\end{verbatim}

\begin{verbatim}
#>  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
#>  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}

\hypertarget{minor-issues-with-nonlinear-least-squares-tools}{%
\subsection{Minor issues with nonlinear least-squares tools}\label{minor-issues-with-nonlinear-least-squares-tools}}

\hypertarget{interim-output-from-the-port-algorithm}{%
\subsubsection{Interim output from the ``port'' algorithm}\label{interim-output-from-the-port-algorithm}}

As the \texttt{nls()} \textbf{man} page states, when the ``port'' algorithm is used with the
\texttt{trace} argument TRUE, the iterations display the objective function value which
is 1/2 the sum of squares (or deviance). The trace display is likely
embedded in the Fortran of the \texttt{nlminb} routine that is called to execute
the ``port'' algorithm, but the factor of 2 discrepancy is nonetheless
unfortunate for users.

\hypertarget{failure-to-return-the-best-result-achieved}{%
\subsubsection{Failure to return the best result achieved}\label{failure-to-return-the-best-result-achieved}}

If \texttt{nls()} reaches a point where it cannot continue but has not found a point
where the relative offset convergence criterion is met, it may simply exit,
especially if a ``singular gradient'' (singular Jacobian) is found. However,
this may occur AFTER the function has made considerable progress in reducing
the sum of squared residuals.
Here is an abbreviated example:

\begin{verbatim}
#> 61216.    (3.56e+03): par = (-2 0.25 150 50)
#> 2.1757    (2.23e+01): par = (-1.9991 0.31711 2.6182 -1.3668)
#> 1.6211    (7.14e+00): par = (-1.9605 -2.6203 2.5753 -0.55599)
#> Error in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE) : 
#>   singular gradient
\end{verbatim}

\begin{verbatim}
#> [1] "Error in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE) : \n  singular gradient\n"
#> attr(,"class")
#> [1] "try-error"
#> attr(,"condition")
#> <simpleError in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE): singular gradient>
\end{verbatim}

Note that the sum of squares has been reduced from 61216 to 1.6211, but
unless \texttt{trace} is invoked, the user will not get any information about this.
This almost trivial change to the \texttt{nls()} function could be
useful to R users.

\hypertarget{estimating-models-that-are-partially-linear}{%
\subsection{Estimating models that are partially linear}\label{estimating-models-that-are-partially-linear}}

The variable projection method (Golub and Pereyra (1973), O'Leary and Rust (2013)) is
generally much more effective than general approaches in finding good solutions
to nonlinear least squares
problems when some of the parameters appear linearly. In our logistic examples,
the asymptote parameters are an illustration. However, identifying which parameters
are linear and communicating this information to estimating functions
is not a trivial task. \texttt{nls()} has an option \texttt{algorithm="plinear"} that allows some partially linear
models to be solved. The other tools, as far as we are aware, do not offer any
such capability. The \texttt{nlstac} package uses a different algorithm for similar goals.

Within \texttt{nls()} itself we must, unfortunately, use
different specifications with different \texttt{algorithm} options.
For example, the explicit model \texttt{y\ \textasciitilde{}\ a\ *\ x\ +\ b} does not work with the linear
modeling function \texttt{lm()}, which requires this model to be specified as \texttt{y\ \textasciitilde{}\ x}.
Within \texttt{nls()}, consider the following FOUR different specifications for the same
problem, plus an intuitive choice, labeled \texttt{fm2a}, that does NOT work.
In this failed attempt, putting the \texttt{Asym} parameter in the model causes the
\texttt{plinear} algorithm
to try to add another term to the model. We believe this is unfortunate, and would
like to see a consistent syntax. At the time of writing, we do
not foresee a resolution for this issue. In the example, we have NOT evaluated
the commands to save space.

\begin{verbatim}
DNase1 <- subset(DNase, Run == 1) # select the data
## using a selfStart model - do not specify the starting parameters
fm1 <- nls(density ~ SSlogis(log(conc), Asym, xmid, scal), DNase1)
summary(fm1)

## using conditional linearity - leave out the Asym parameter
fm2 <- nls(density ~ 1 / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(xmid = 0, scal = 1),
                 algorithm = "plinear")
summary(fm2)

## without conditional linearity
fm3 <- nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1,
                 start = list(Asym = 3, xmid = 0, scal = 1))
summary(fm3)

## using Port's nl2sol algorithm
fm4 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(Asym = 3, xmid = 0, scal = 1),
                 algorithm = "port"))
summary(fm4)

## using conditional linearity AND Asym does NOT work
fm2a <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)), 
                 data = DNase1, start = list(Asym=3, xmid = 0, scal = 1),
                 algorithm = "plinear", trace = TRUE))
summary(fm2a)
\end{verbatim}

\hypertarget{models-with-indexed-parameters}{%
\subsection{Models with indexed parameters}\label{models-with-indexed-parameters}}

Some models have several common parameters and
others that are tied to particular cases.
The \textbf{man} file for \texttt{nls()} includes an example of a situation in which
parameters are indexed but which uses the ``plinear'' option as an added complication.
Running this example reveals that the answers for the parameters are NOT indexed
as in a vector.
That is, we do not see \texttt{a{[}1{]},\ a{[}2{]},\ a{[}3{]}} but \texttt{a1,\ a2,\ a3}.
This is no doubt because programming for indexed parameters is challenging.
We note that there are capabilities in packages for mixed-effects modeling
such as \texttt{nlme} (Pinheiro et al. (2013)), \texttt{bbmle} (Bolker and Team (2013)), and \texttt{lme4} (p-lme4) for
estimating models of such type.

\hypertarget{tests-and-use-case-examples}{%
\subsection{Tests and use-case examples}\label{tests-and-use-case-examples}}

Maintainers of packages need suitable tests and use-case examples in order

\begin{itemize}
\tightlist
\item
  to ensure packages work properly, in particular, giving
  results comparable to or better than the functions they are to replace.
\item
  to test individual solver functions to ensure they work across the range of
  calling mechanisms, that is, different ways of supplying inputs to the solver(s);
\item
  to pose ``silly'' inputs to see if these bad inputs are caught by the programs.
\end{itemize}

Such goals align with the aims of \textbf{unit testing}
(e.g., \emph{\url{https://towardsdatascience.com/unit-testing-in-r-68ab9cc8d211}}, Wickham (2011), Wickham et al. (2021)
and the conventional R package testing tools).
In our work, one of us has developed a working prototype package at \emph{\url{https://github.com/ArkaB-DS/nlsCompare}} .
A primary design objective of this is to allow the summarization of multiple tests in
a compact output. The prototype has a vignette to illustrate its use.

\hypertarget{documentation-and-resources}{%
\subsection{Documentation and resources}\label{documentation-and-resources}}

In our investigation, we built several resources, which are now part of
the repository \emph{\url{https://github.com/nashjc/RNonlinearLS/}}. Particular items
are:

\begin{itemize}
\item
  A BibTex bibliography for use with all documents in this project, but which
  has wider application to nonlinear least squares projects in general
  (\emph{\url{https://github.com/nashjc/RNonlinearLS/blob/main/BibSupport/ImproveNLS.bib}}).
\item
  \texttt{MachID.R} offers a suggested concise summary function
  to identify a particular computational system used for tests. A discussion
  of how it was built and the resources needed across platforms is given in
  at \emph{\url{https://github.com/nashjc/RNonlinearLS/tree/main/MachineSummary}}.
\item
  John C. Nash and Bhattacharjee (2022) is an explanation of the construction of the \texttt{nlspkg} from the
  \texttt{nls()} code in R-base.
\item
  As the 2021 Summer of Code period was ending, one of us (JN) was invited to prepare
  a review of optimization in R. Ideas from the present work have been instrumental
  in the creation of John C. Nash (2022).
\end{itemize}

\hypertarget{future-of-nonlinear-model-estimation-in-r}{%
\subsection{Future of nonlinear model estimation in R}\label{future-of-nonlinear-model-estimation-in-r}}

Given its importance to R, it is possible that \texttt{nls()} will remain more or less
as it has been for the past several decades. If so, the focus of discussion
should be the measures needed to secure its continued operation for legacy
purposes and how that may be accomplished.
We welcome an opportunity to participate in such conversations.

To advance the stability and maintainability of R, we believe the program
objects (R functions)
that are created by tools such as \texttt{nls()} should have minimal
cross-linkages and side-effects.
The aspects of \texttt{nls()} that concern us follow.

\begin{itemize}
\item
  R tools presume data and parameters needed are available in an
  accessible environment. This provides a compact syntax to invoke the
  calculations, but the wrong data can be used if the
  internal search finds a valid name that is not the object we want,
  or if \texttt{subset} or \texttt{na.action} settings modify the selection, or
  \texttt{weights} are applied.
\item
  Mixing of R, C and Fortran code adds to the burden of following the
  program logic.
\item
  The class \texttt{nls} structure simplifies calls with its rich set of functions, but
  also adds to the task of understanding what has been done. A design that isolates
  the setup, solution, and post-solution parts of complicated calculations reduces
  the number of objects that must be kept in alignment.
\end{itemize}

Given the existence of examples of good practices such as analytic derivatives,
stabilized solution of Gauss-Newton equations, and bounds-constrained parameters,
base R tools should
be moving to incorporate them. These capabilities are available now by
using several tools, but it would be helpful if they were unified.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

Hans Werner Borchers was helpful in developing the GSoC project motivating this article
and in comments on this and related work. Heather Turner co-mentored the project
and helped guide the progress of the work. Exchanges with Fernando Miguez helped to
clarify aspects of \texttt{selfStart} models and instigated the ``Introduction to nlsr''
vignette. Colin Gillespie (package \texttt{benchmarkme}) has been helpful in guiding our
attempts to succinctly summarize computing environments.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{references-1}{%
\section*{References}\label{references-1}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-R-onls-manual}{}}%
Andrej-Nikolai Spiess. 2022. \emph{Onls: Orthogonal Nonlinear Least-Squares Regression}. \url{https://CRAN.R-project.org/package=onls}.

\leavevmode\vadjust pre{\hypertarget{ref-bateswatts}{}}%
Bates, D. M., and D. G. Watts. 1988. \emph{Nonlinear Regression Analysis and Its Applications}. Wiley.

\leavevmode\vadjust pre{\hypertarget{ref-BatesWatts81}{}}%
Bates, Douglas M., and Watts, Donald G. 1981. {``A Relative Offset Orthogonality Convergence Criterion for Nonlinear Least Squares.''} \emph{Technometrics} 23 (2): 179--83.

\leavevmode\vadjust pre{\hypertarget{ref-p-nlstools}{}}%
Baty, Florent, and Marie-Laure Delignette-Muller. 2013. \emph{Nlstools: Tools for Nonlinear Regression Diagnostics}.

\leavevmode\vadjust pre{\hypertarget{ref-nlsMicrobio-manual}{}}%
---------. 2014. \emph{nlsMicrobio: Data Sets and Nonlinear Regression Models Dedicated to Predictive Microbiology}.

\leavevmode\vadjust pre{\hypertarget{ref-p-bbmle}{}}%
Bolker, Ben, and R Development Core Team. 2013. \emph{Bbmle: Tools for General Maximum Likelihood Estimation}. \url{http://CRAN.R-project.org/package=bbmle}.

\leavevmode\vadjust pre{\hypertarget{ref-gslnls23}{}}%
Chau, Joris. 2023. \emph{Gslnls: GSL Nonlinear Least-Squares Fitting}. \url{https://CRAN.R-project.org/package=gslnls}.

\leavevmode\vadjust pre{\hypertarget{ref-nlshelper-manual}{}}%
Duursma, Remko. 2017. \emph{Nlshelper: Convenient Functions for Non-Linear Regression}. \url{https://CRAN.R-project.org/package=nlshelper}.

\leavevmode\vadjust pre{\hypertarget{ref-minpacklm12}{}}%
Elzhov, Timur V., Katharine M. Mullen, Andrej-Nikolai Spiess, and Ben Bolker. 2012. \emph{Minpack.lm: R Interface to the Levenberg-Marquardt Nonlinear Least-Squares Algorithm Found in MINPACK, Plus Support for Bounds}. R Project for Statistical Computing. \url{http://CRAN.R-project.org/package=minpack.lm}.

\leavevmode\vadjust pre{\hypertarget{ref-GSL-manual}{}}%
Galassi, Mark, Jim Davies, James Theiler, Brian Gough, and Gerard Jungman. 2009. \emph{GNU Scientific Library - Reference Manual, Third Edition, for GSL Version 1.12 (3. Ed.).}

\leavevmode\vadjust pre{\hypertarget{ref-Bioconductor}{}}%
Gentleman, Robert C., Vincent J. Carey, Douglas M. Bates, Ben Bolstad, Marcel Dettling, Sandrine Dudoit, Byron Ellis, et al. 2004. {``Bioconductor: Open Software Development for Computational Biology and Bioinformatics.''} \emph{{Genome Biology}} 5 (R80). \url{https://doi.org/10.1186/gb-2004-5-10-r80}.

\leavevmode\vadjust pre{\hypertarget{ref-Golub1973}{}}%
Golub, G. H., and V. Pereyra. 1973. {``The Differentiation of Pseudo-Inverses and Nonlinear Least Squares Problems Whose Variables Separate.''} \emph{SIAM Journal of Numerical Analysis} 10 (2): 413--32.

\leavevmode\vadjust pre{\hypertarget{ref-nls2-22}{}}%
Grothendieck, G. 2022. \emph{Nls2: Non-Linear Regression with Brute Force}. \url{https://CRAN.R-project.org/package=nls2}.

\leavevmode\vadjust pre{\hypertarget{ref-Hartley1961}{}}%
Hartley, H. O. 1961. {``The Modified Gauss-Newton Method for the Fitting of Non-Linear Regression Functions by Least Squares.''} \emph{Technometrics} 3: 269-\/-280.

\leavevmode\vadjust pre{\hypertarget{ref-Huet2004}{}}%
Huet, S., A. Bouvier, M.-A. Poursat, and E. Jolivet. 2004. \emph{Statistical Tools for Nonlinear Regression: A Practical Guide with {S-PLUS} Examples, 2nd Edition}. Berlin \& New York: Springer-Verlag.

\leavevmode\vadjust pre{\hypertarget{ref-nlsr2023manual}{}}%
John C Nash, and Duncan Murdoch. 2023. \emph{{nlsr: Functions for Nonlinear Least Squares Solutions}}.

\leavevmode\vadjust pre{\hypertarget{ref-tvrdik2007}{}}%
Josef Tvrdík and Ivan Křivý and Ladislav Mišík. 2007. {``Adaptive Population-Based Search: Application to Estimation of Nonlinear Regression Parameters.''} \emph{{Computational Statistics \& Data Analysis}} 52 (2): 713--24. \url{https://doi.org/10.1016/j.csda.2006.10.014}.

\leavevmode\vadjust pre{\hypertarget{ref-Levenberg1944}{}}%
Levenberg, Kenneth. 1944. {``A Method for the Solution of Certain Non-Linear Problems in Least Squares.''} \emph{Quarterly of Applied Mathematics} 2: 164-\/-168.

\leavevmode\vadjust pre{\hypertarget{ref-Marquardt1963}{}}%
Marquardt, Donald W. 1963. {``{An Algorithm for Least-Squares Estimation of Nonlinear Parameters}.''} \emph{SIAM Journal on Applied Mathematics} 11 (2): 431--41.

\leavevmode\vadjust pre{\hypertarget{ref-MiguezNLRAA2021}{}}%
Miguez, Fernando. 2021. \emph{{nlraa: Nonlinear Regression for Agricultural Applications}}. \url{https://CRAN.R-project.org/package=nlraa}.

\leavevmode\vadjust pre{\hypertarget{ref-jn77ima}{}}%
Nash, John C. 1977. {``Minimizing a Nonlinear Sum of Squares Function on a Small Computer.''} \emph{Journal of the Institute for Mathematics and Its Applications} 19: 231--37.

\leavevmode\vadjust pre{\hypertarget{ref-cnm79}{}}%
---------. 1979. \emph{Compact Numerical Methods for Computers: Linear Algebra and Function Minimisation}. Bristol: Adam Hilger.

\leavevmode\vadjust pre{\hypertarget{ref-NashWires22}{}}%
---------. 2022. {``{Function minimization and nonlinear least squares in R}.''} \emph{WIREs Computational Statistics}, no. e1580. https://doi.org/\url{https://doi.org/10.1002/wics.1580}.

\leavevmode\vadjust pre{\hypertarget{ref-PkgFromRbase22}{}}%
Nash, John C., and Arkajyoti Bhattacharjee. 2022. {``Making a Package from Base {R} Files.''} \emph{R-Bloggers}, January. \url{https://www.r-bloggers.com/2022/01/making-a-package-from-base-r-files/}.

\leavevmode\vadjust pre{\hypertarget{ref-p-optimx}{}}%
Nash, John C, and Ravi Varadhan. 2011. \emph{\emph{Optimx}: A Replacement and Extension of the {optim()} Function}. Nash Information Services Inc.; Johns Hopkins University.

\leavevmode\vadjust pre{\hypertarget{ref-nash1996nonlinear}{}}%
Nash, John C., and Paul Velleman. 1996. {``Nonlinear Estimation Combining Visual Fitting with Optimization Methods.''} In \emph{Proceedings of the Section on Physical and Engineering Sciences of the American Statistical Association}, 256--61. American Statistical Association.

\leavevmode\vadjust pre{\hypertarget{ref-NISTnls-manual}{}}%
National Institutes for Standards, original from, and Technology http://www.itl.nist.gov/div898/strd/nls/nls\_main.shtml R port by Douglas Bates. 2012. \emph{NISTnls: Nonlinear Least Squares Examples from NIST}. \url{https://CRAN.R-project.org/package=NISTnls}.

\leavevmode\vadjust pre{\hypertarget{ref-OlearyRust13}{}}%
O'Leary, Dianne P., and Bert W. Rust. 2013. {``Variable Projection for Nonlinear Least Squares Problems.''} \emph{Computational Optimization and Applications} 54 (3): 579--93.

\leavevmode\vadjust pre{\hypertarget{ref-nlsmultstart}{}}%
Padfield, Daniel, and Granville Matheson. 2020. \emph{Nls.multstart: Robust Non-Linear Regression Using AIC Scores}. \url{https://CRAN.R-project.org/package=nls.multstart}.

\leavevmode\vadjust pre{\hypertarget{ref-p-nlme}{}}%
Pinheiro, Jose, Douglas Bates, Saikat DebRoy, Deepayan Sarkar, and R Core Team. 2013. \emph{Nlme: Linear and Nonlinear Mixed Effects Models}.

\leavevmode\vadjust pre{\hypertarget{ref-m-nlsmsn}{}}%
Prates, Marcos, Victor Lachos, and Aldo Garay. 2021. \emph{Nlsmsn: Fitting Nonlinear Models with Scale Mixture of Skew-Normal Distributions}. \url{https://CRAN.R-project.org/package=nlsmsn}.

\leavevmode\vadjust pre{\hypertarget{ref-Ratkowsky1983}{}}%
Ratkowsky, David A. 1983. \emph{Nonlinear Regression Modeling: A Unified Practical Approach}. New York; Basel: Marcel Dekker Inc.

\leavevmode\vadjust pre{\hypertarget{ref-m-nlstac}{}}%
Rodriguez-Arias, Mariano, Juan Antonio Fernandez, Javier Cabello, and Rafael Benitez. 2020. \emph{Nlstac: An r Package for Fitting Separable Nonlinear Models}. \url{https://CRAN.R-project.org/package=nlstac}.

\leavevmode\vadjust pre{\hypertarget{ref-nlsic-manual}{}}%
Sokol, Serguei. 2022. \emph{Nlsic: Non Linear Least Squares with Inequality Constraints}. \url{https://CRAN.R-project.org/package=nlsic}.

\leavevmode\vadjust pre{\hypertarget{ref-Torvisco2018}{}}%
Torvisco, Juan Antonio Fernandez, Mariano Rodrıguez-Arias Fernandez, and Javier Cabello Sanchez. 2018. {``A New Algorithm to Fit Exponential Decays Without Initial Guess.''} {Faculty of Sciences and Mathematics, University of Niˇs, Serbia}. \url{https://doi.org/doi.org/10.2298/FIL1812233T}.

\leavevmode\vadjust pre{\hypertarget{ref-R-crsnls-manual}{}}%
Tvrdík, Josef. 2016. \emph{Crsnls: Nonlinear Regression Parameters Estimation by 'CRS4HC' and 'CRS4HCe'}. \url{https://CRAN.R-project.org/package=crsnls}.

\leavevmode\vadjust pre{\hypertarget{ref-HWtestthat11}{}}%
Wickham, Hadley. 2011. {``{testthat: Get Started with Testing}.''} \emph{The R Journal} 3: 5--10. \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-HWdevtools21}{}}%
Wickham, Hadley, Jim Hester, Winston Chang, and Jennifer Bryan. 2021. \emph{{devtools: Tools to Make Developing R Packages Easier}}. \url{https://CRAN.R-project.org/package=devtools}.

\end{CSLReferences}

\bibliography{./ImproveNLS.bib}

\address{%
John C. Nash\\
retired professor, University of Ottawa\\%
Telfer School of Management\\ Ottawa ON Canada K1N 6N5\\
%
%
\textit{ORCiD: \href{https://orcid.org/0000-0002-2762-8039}{0000-0002-2762-8039}}\\%
\href{mailto:profjcnash@gmail.com}{\nolinkurl{profjcnash@gmail.com}}%
}

\address{%
Arkajyoti Bhattacharjee\\
Indian Institute of Technology\\%
Department of Mathematics and Statistics\\ Kanpur\\
%
%
%
\href{mailto:arkastat98@gmail.com}{\nolinkurl{arkastat98@gmail.com}}%
}