### 2023-07-03 RefactoringNLS/revision202307/ReviewResponses.txt
Responses -- first draft

Review 1

Referee report of “A Comparison of R Tools for Nonlinear Least
Squares Modeling”.
ID: 2023-15
General
This is a long paper that covers many topics related to nonlinear least 
squares fitting with R. This mixture of many topics can be confusing and 
I think the paper would benefit from a clearer definition of its purpose 
and focus. Once a clearer focus is established it will be easier to decide 
which topics are relevant to that purpose.

The paper’s title indicates that the purpose is to compare and contrast 
facilities in base R, meaning the nls function and related facilities 
in the stats package, and several external packages for nonlinear
least-squares modeling that are available on CRAN. There is a listing 
of packages related to nonlinear least squares on page 2 but the majority 
of the discussion is focussed on nls, in the stats package that is part 
of base R, and the nlsr package (Nash and Murdoch 2023), written by this 
paper’s authors and others (on CRAN this package is said to be the work 
of four authors but only two of them are cited in the paper).

Response 1:01

Fernando Miguez and Arkajyoti Bhattacharjee, who are listed as contributors,
provided comments that were very helpful in revising package nlsr, but Nash 
and Murdoch provided all the code to our knowledge.

There are 
also indications that the paper is about replacing nls in base R with 
functions from nlsr. The abstract begins by citing the Google Summer 
of Code project “Improvements to nls()” and the first sentence in the 
“Principal messages”” section indicates that the purpose is to create 
a refactored nls() based on the work described here. Toward the end of 
the paper, the section on “Future of nonlinear model estimation in R” 
also indicates that the goal is to enhance or replace nls() and related parts
of base R.

So is the purpose of the paper:
• To introduce the nlsr package and describe its capabilities?
• To compare and contrast several available packages for nonlinear least-squares?
• To advocate that the authors’ code replace nls in base R?

I would recommend concentrating on the first objective — introducing 
the nlsr package and demonstrating the advantages that it provides 
relative to nls — and I will address my comments to that purpose. I
hope this approach will be of use to the authors and the editor.
It is certainly appropriate when describing the nlsr package to 
contrast it with facilities like nls() in base R. To help with this 
I will provide some background on the development of nls() and the stats
package in general.

Response 1:02
     This overlaps with comments by the second referee. See Response 2:01.

Detailed comments

nls is S and in R

Although nls() has been part of R for over 25 years, it originated 
much earlier in the “Statistical Models in S” project, documented 
in Chambers and Hastie (1992), and based on work starting in the mid-1980’s.
The purpose was to create a unified set of tools for fitting statistical 
models, including nonlinear models with a scalar objective, such as a 
log-likelihood, or as a nonlinear regression model. Many capabilities for
R, including S3 classes and methods, the formula language for linear 
predictors, and model frames and model matrices, were based directly on 
the results of this project in S. (As described by Ross Ihaka, one of the 
two original developers of R, their goal was to produce a language that 
was “not unlike S” so the resemblance was not accidental.)

What was called QPE (quantitative programming environment), then “The New 
S Language”, then “S3” represented a huge step forward for nonlinear 
modeling because it could represent model functions as expressions 
(i.e. abstract syntax trees) that could be manipulated in the system 
itself. This allowed for symbolic differentiation.

The original implementation of the deriv function for symbolic 
differentiation was written in S. When porting deriv to R in the 
mid-1990’s Ross Ihaka moved quite a bit of the functionality to C code 
because it was comparatively easy to do so and, on the hardware of the 
day, the C implementation provided a considerable speed advantage. The 
speed of the differentiation itself isn’t usually a problem - it’s the
common-sub-expression elimination that can be the bottleneck.

This legacy is also the reason that the Jacobian matrix is the "gradient" 
attribute of the returned value of the model function. For a nonlinear 
least squares problem, where the model function returns a vector, this 
deriv attribute should be called "Jacobian" but the version for the 
scalar objective, for which the vector of partial derivatives is the 
gradient, was written first. Because deriv only has access to the 
expression and not to the model frame it is not possible to determine 
if the value will be a scalar or a vector and the name of the attribute 
was left as "gradient". A further complication is that numerical values 
in S and R are always stored as vectors, with the understanding that a 
scalar is represented as a vector of length 1. Even if it were desirable 
to distinguish between scalars and vectors, internally it is impossible 
to determine if a vector of length 1 is intended to be a scalar or to 
be a vector that just happens to have length 1.

The legacy is also the reason that initial data manipulation, such as 
evaluating a subset or na.action, is performed by non-standard evaluation 
of a model.frame. The non-standard evaluation means that, within the 
nls function evaluation, the code accesses its own call (with match.call) 
and re-writes the function name to be stats::model.frame then evaluates 
the rewritten call (around line 580 of nls.R). Handling all the special 
cases is complicated but the purpose is so that arguments like subset 
or na.action and conversion of formulas to model frames to model matrices 
are handled consistently across the modeling functions in the stats 
package and without code duplication.

Thomas Lumley has called this “replace the function name in my own call” 
approach the “standard, non-standard evaluation idiom” in R.

Another aspect of this legacy is the expected behavior of print, summary, 
plot, etc. applied to an object representing a fitted model. Peculiarities 
like summary producing longer descriptions than print or the choice of 
printing the sum of squared residuals or the estimated standard deviation 
of the residuals are for consistency with the other model output.

Response 1:06 We now mention the background to summary/print.

It is also why the extractor for the parameter values is coef, even though 
the parameters in a nonlinear regression model are not necessarily 
coefficients, and why they are shown in the summary using print.coefmat.

These choices, which allow users to transfer experience from one modelling 
situation to another, are neither arbitrary nor trivial.

Function closure with a shared environment

One of the major differences between the S and R implementations of nls is 
the use of function closures in R to allow several functions in the m field 
of an nls model (see below for an example) shared access to parameter values, 
covariate values, the response vector and various pieces of auxillary 
information.

In R if a function is defined during the evaluation of another function its 
“closure” includes the evaluation environment of its parent. If several 
functions are defined and returned, say in a list, then they have access 
to this common environment. This shared environment is where parameter 
updates and subsequent evaluation of the model function, the Jacobian and 
various decompositions occur.

Self-starting models in S and R

Self-starting models were developed for nlme package for S, Pinheiro and 
Bates (2000), to automate the fitting of nonlinear regression models to 
repeated measures data collected on several observational units, e.g. 
several different subjects. Rather than requiring the analyst to provide 
starting values for each subject’s data, the steps of formulating and 
refining such starting values were encapsulated in functions. These 
functions also used symbolic derivatives to provide Jacobian evaluations 
and frequently fit a reduced form of the model using the plinear algorithm 
to refine the starting values, resulting in the “starting values” being 
the actual parameter estimates. These are fed into the default algorithm 
to produce a summary but with the understanding that the default algorithm 
will converge at the first evaluation (i.e. 0 iterative steps) if the 
convergence criterion is the relative offset. The point here is that the 
relative offset criterion depends only on the current parameter values 
and not on changes in the parameter values so there is only one additional 
evaluation required after the plinear algorithm converges.

For the illustrative example starting at the bottom of page 2 of this 
paper the SSlogis self-starting model can be used.

weeddf <- data.frame(
tt = 1:12,
weed = c(
5.308, 7.24, 9.638, 12.866, 17.069, 23.192,
31.443, 38.558, 50.156, 62.948, 75.995, 91.972
)
)
summary(hobbsm1 <- nls(weed ~ SSlogis(tt, Asym, xmid, scal), weeddf))
Formula: weed ~ SSlogis(tt, Asym, xmid, scal)
Parameters:
Estimate Std. Error t value Pr(>|t|)
Asym 196.1862
11.3069
17.35 3.17e-08 ***
xmid 12.4173
0.3346
37.11 3.72e-11 ***
scal
3.1891
0.0698
45.69 5.77e-12 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 0.5362 on 9 degrees of freedom
Number of iterations to convergence: 0
Achieved convergence tolerance: 1.118e-06
str(hobbsm1)
List of 6
$ m
:List of 16
..$ resid
:function ()
..$ fitted
:function ()
..$ formula
:function ()
..$ deviance :function ()
..$ lhs
:function ()
..$ gradient :function ()
..$ conv
:function ()
..$ incr
:function ()
..$ setVarying:function (vary = rep_len(TRUE, np))
..$ setPars
:function (newPars)
..$ getPars
:function ()
..$ getAllPars:function ()
..$ getEnv
:function ()
3..$ trace
:function ()
..$ Rmat
:function ()
..$ predict
:function (newdata = list(), qr = FALSE)
..- attr(*, "class")= chr "nlsModel"
$ convInfo
:List of 5
..$ isConv
: logi TRUE
..$ finIter
: int 0
..$ finTol
: num 1.12e-06
..$ stopCode
: int 0
..$ stopMessage: chr "converged"
$ data
: symbol weeddf
$ call
: language nls(formula = weed ~ SSlogis(tt, Asym, xmid, scal), data = weeddf, algorithm
$ dataClasses: Named chr "numeric"
..- attr(*, "names")= chr "tt"
$ control
:List of 7
..$ maxiter
: num 50
..$ tol
: num 1e-05
..$ minFactor : num 0.000977
..$ printEval : logi FALSE
..$ warnOnly
: logi FALSE
..$ scaleOffset: num 0
..$ nDcentral : logi FALSE
- attr(*, "class")= chr "nls"
This only takes a few milliseconds on a laptop computer
microbenchmark::microbenchmark(nls(weed ~ SSlogis(tt, Asym, xmid, scal), weeddf), times=100L)
Unit: milliseconds
expr
min
lq
mean
nls(weed ~ SSlogis(tt, Asym, xmid, scal), weeddf) 3.483984 3.624459 4.599596
median
uq
max neval
3.92275 5.035838 13.76178
100
although it turns out that the SSlogisJN formulation is slightly faster in this case
summary(hobbsm2 <- nls(weed ~ nlsr::SSlogisJN(tt, Asym, xmid, scal), weeddf))
Formula: weed ~ nlsr::SSlogisJN(tt, Asym, xmid, scal)
Parameters:
Estimate Std. Error t value Pr(>|t|)
Asym 196.1862
11.3069
17.35 3.17e-08 ***
xmid 12.4173
0.3346
37.11 3.72e-11 ***
scal
3.1891
0.0698
45.69 5.77e-12 ***
---
Signif. codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
Residual standard error: 0.5362 on 9 degrees of freedom
Number of iterations to convergence: 3
Achieved convergence tolerance: 2.419e-06
microbenchmark::microbenchmark(nls(weed ~ nlsr::SSlogisJN(tt, Asym, xmid, scal), weeddf), times=10
4Unit: milliseconds
expr
min
lq
nls(weed ~ nlsr::SSlogisJN(tt, Asym, xmid, scal), weeddf) 1.960326 2.069342
mean
median
uq
max neval
2.492946 2.184441 2.554391 11.43154
100

Some of the discussion in the “Jacobian code in selfStart models” section 
(p. 9) of this paper claims that the plinear fit followed by a fit using 
the default approach would be difficult to maintain and could be very slow
but that has not been the case. It is known (and, in fact, stated in the 
documentation for SSlogis) that the call to the default nls algorithm will 
always declare convergence immediately.

Response 1:03 Can we have a pointer to the documentation of the immediate 
      convergence? We have checked the documentation ?SSlogis, as well as 
      selfStart and getInitial, but do not see a mention in R 4.3.1.
      ??Interim response --

Also in that section it is claimed that deriv creates “hidden variables” 
because the default for the tag argument to deriv is .expr. This tag is 
only used in bindings within the evaluation environment of the model 
function, which is never visible to the user unless they are inspecting 
the function evaluation with a symbolic debugger. The only reason for 
the default tag starting with "." is to avoid conflicts with parameter 
or covariate names in the model expression. But it is only the default.

Response 1:04 Thanks. Reworded to explain the "why".

Recommendations

As stated in the “General” section, I think the purpose of the paper 
should be clarified. If the purpose is to introduce the nlsr package 
and contrast it with the existing implementation of nls in the stats
package, then it might be helpful to understand why nls was implemented 
the way it was.

Response 1:05 The historical context could be very useful to the R community.
    Would the editors consider asking that Referee 1 consider recasting
    the material as a companion note?

    See also 2:01 for clarification of purpose.

I hope that this unconventional referee’s report is of use to the authors 
and the editor.

References

Chambers, John M., and Trevor J. Hastie, eds. 1992. Statistical Models 
in S. Routledge. https://doi.org/10.1201/9780203738535.

Nash, John C, and Duncan Murdoch. 2023. Nlsr: Functions for Nonlinear 
Least Squares Solutions - Updated 2022. https://CRAN.R-project.org/package=nlsr.

Pinheiro, José, and Douglas Bates. 2000. Mixed-Effects Models in S and 
S-PLUS. Springer New York.

Review 2

Referee report: A Comparison of R Tools for Nonlinear Least
Squares Modeling
Id: 2023-15
2023-06-28

General

This article provides a vignette-style overview of the nlsr-package and 
highlights its design choices and behavior in comparison to base R’s nls() 
function as well as the nonlinear least squares solvers provided by the 
R-packages minpack.lm and gslnls. The article’s main contribution is to 
make R users (more) aware of the historic shortcomings and pitfalls in 
using R’s nls() function and how the nlsr-package is able to serve as a 
more reliable alternative. Overall, the article is well-written and 
structured into coherent subthemes and sections. Given that nls() is 
the de facto standard for solving nonlinear least squares problems in R, I
believe this article provides valuable insights and is well-suited for 
publication in the R Journal after some revisions and minor clarifications.

My main advice is to revise the title and introduction of the article. As 
it currently stands, it is suggested that the article provides an 
(objective) comparison of nonlinear least squares solvers available in R. 
The body of the article reads more like a vignette for the nlsr-package, 
explaining its design choices and comparing its behavior to nls() and other 
NLS solvers based on a few selected examples. As such, a more appropriate title
could be, for instance, The current state of R tools for nonlinear least 
squares modeling or similar.

Response 2:01

We appreciate the positive comments and agree with the suggested title change.

Response 2:02

Both referees have requested revisions to the introduction, and we hope that 
we have addressed the comments, essentially by clarifying the goals and how
they arose.

Detailed comments

1. The Hobbs weed example as an illustration of the performance of different 
NLS solvers is lacking a thorough comparison. I understand that the aim of 
the authors is to provide a data example that demonstrates the reliability 
of nlsr in favor of nls(), but nls() converges to the same solution as nlsr 
for all model setups when using the Port algorithm. This goes unnoticed in 
the text.

Response 2:03 "port" cases added. The "appears unfinished" comment in the 
       documentation led us to avoid its use. This is the sort of issue we 
       would very much like to see resolved, since "port" is the only tool
       for bounds constraints.

Also, the model setup (Logistic3T) for which gsl_nls() fails can easily be 
made to converge correctly by changing either the algorithm or the scaling 
method, see the Illustrating code section. My advice is to provide a more 
comprehensive comparison of the available NLS solvers, or to pick a different
data example that better highlights the benefits of nlsr. For instance, trying 
to solve the BoxBOD regression problem 
(https://www.itl.nist.gov/div898/strd/nls/data/boxbod.shtml), 
nls() fails with starting values θ1 = 1, θ2 = 1 using both the Gauss-Newton 
and Port algorithm, whereas nlsr::nlxb() converges correctly.

Response 2:04 We have added a paragraph

	We consider the code inside package `nlsr` to be competently coded, 
	with the advantage over `nls()` of the Levenberg-Marquardt stabilization 
	in common with `minpack.lm` and `glsnls`. The success rate is high,
        but is not perfect. However, our principal message is that we obtain 
        diagnostic information in the form of the singular values of the Jacobian 
	that allow us to try alternative starts, algorithms or programs in cases 
	where there are indications of early termination.

   We appreciate the pointers on other examples. However, a substantive comparison
   would make for a very long and likely boring paper. The BoxBOD example is, in
   many ways, similar to the Hobbs example in its nasty characteristics. 
   Some years ago, one of us created a package NISTopt to use the NIST examples 
   on function minimization tools (as per package NISTnls). See, for example, 
   https://rdrr.io/rforge/NISTopt/ 
   Experiences with these problems have led to an opinion (i.e., open to debate)
   as to the usefulness of such extreme cases. They underline a point of view that
   good diagnostics are more helpful than a "usually succeeds" level of reliability.

The section Returned results of nls() and other tools discusses the complexity 
of S3-objects of class nls, illustrated by the retrieval of the input data 
from an nls-object. It seems ill-advised to run eval(parse(text=result$data)), 
since it evaluates the name of the data object in the user’s environment, 
which may not exist. Instead, one should use model = TRUE to store the input 
data in the returned nls-object. This behavior is consistent with lm(), so 
perhaps less confusing than currently presented in the manuscript. The 
authors mention that nlxb() returns a much simpler structure of 11
items in one level, but I did not manage to find the input data object in 
the nlsr-object returned by nlxb().

Response 2:05 ??

The section Functional specification of problems mentions the use of 
nlsr::nlfb() to solve nonlinear least squares problems using a function 
to define the residuals. Since nlsr::nlfb() only has a notion of the 
residuals, how is it able to disentangle model and response, for the 
purpose of e.g. generating model predictions? The authors mention that 
the function wrapnlsr() can be used to return an object of class nls, 
but this does not seem to work for problems defined through a residual 
function as in nlsr::nlfb()? If indeed the case, how does one evaluate 
e.g. model predictions, parameter confidence intervals, etc. when using 
nlsr::nlfb()? This remains unclear to me.

Response 2:06 The referee's comments point out that we have not explained
    clearly enough the important differences that arise between modeling
    and minimizing a nonlinear sum of squares. We have added a short 
    exposition using the Rosenbrock problem. The difference in context
    is important and we thank the referee, as other readers would surely
    have had similar concerns.

4. The section Philosophical considerations hints at the challenge of 
performing inference when parameter bound-constraints are involved in 
the optimization. In case of physical or model constraints on the
parameters, a common approach is to reparametrize the model to be 
unconstrained in the parameters. Perhaps the authors can highlight an 
example scenario where imposing parameter constraints is preferable to 
model reparametrization (the included data examples with parameter 
constraints appear somewhat artificial), or elaborate further on 
the possible downsides of reparametrizing the model, if any.

Response 2:07 Reparametrization is likely the appropriate approach if
    traditional measures of dispersion are wanted.

    Opinion: One of us (JN) has been interested for > 25 years in 
    comparing the approaches of numerical analysts and statisticians.
    The former are insterested in the nonlinear LS problem, for which
    constraints are a natural characteristic of some problems, while the
    latter want nonlinear models, where bounds constraints can be a
    nuisance to the dispersion ideas. I'll not pretend my knowledge is
    sufficiently broad to be able to capture all the nuances, but I
    believe a dialog on these matters could be very fruitful in long
    term better understanding. As far as I am aware, the literature on
    this area from a statistical viewpoint is quite thin.

    For the paper: We agree the examples are essentially artificial. They
    are intended to show how to use the features. A discussion of an upper 
    limit on the Asym parameter (i.e., weed density in some units) or the 
    scal parameter (measuring how fast this is reached) might be possible, 
    but the people who provided the problem are no longer available.

    We have added two small comments on recommended usage of bounds. One
    is to control for wild excursions in the optimization trajectory, as 
    we believe this is sensible when bounds constraints are available, 
    as in nlsr::nlxb. We also note that in cases where a simple and 
    obvious transformation is available, such as squaring a parameter to 
    ensure positivity, this is likely a better approach than imposing bounds.

5. The section Programming language makes the argument that keeping to a 
single programming language can allow for easier maintenance and upgrades, 
and that the performance penalty for using code entirely in R has been 
much reduced in recent years. I agree that there is not much of a 
performance penalty when solving small problems, such as the data examples 
included in the article. However, when solving problems using a large amount 
of data or a large number of parameters, I have experienced that the run-time 
(as well as memory usage) can become an actual bottleneck. As an example, 
consider the Penalty function I problem (Moré, Garbow, and Hillstrom (1981)) 
included in the Illustrating code section.

Response 2:08  In that we are in general agreement with this comment, we have added
    some clarification to advise that it is worthwhile being aware of how long
    calculations are taking.

    The referee's commends ant Illustrating Code led us to try the Pen1 problem
    (it is part of the funconstrain package https://github.com/jlmelville/funconstrain
    to which one of us recently added Hessian code). This showed that the R code
    of nlsr::nlfb() was faster than the compiled Fortran of minpack.lm::nls.lm.
    Moreover, the all-R code optimx::Rcgmin was 2 orders of magnitude faster --
    it does not build a matrix!

Minor comments

• Avoid emphasizing words by writing in all capitals. Examples: NOT (pg. 14, 15), 
WEIGHTED (pg. 13) and more.

Response 2:09  Fair enought. Done.

• The link https://towardsdatascience.com/unit-testing-in-r-68ab9cc8d211 (as well 
as the links to github) may be subject to change or become unavailable in the future.

Response 2:10  We are not sure how to answer this, though we agree with the sentiment. Does
     the editor have any advice?

Illustrating code

## Hobbs weed example
weeddf <- data.frame(
tt = 1:12,
weed = c(5.308, 7.24, 9.638, 12.866, 17.069, 23.192, 31.443, 38.558, 50.156, 62.948, 75.995,
 91.972)
)
### nls-port all setups
nls(formula = weed ~ b1 / (1 + b2 * exp(-b3 * tt)), data = weeddf,
     start = c(b1 = 1, b2 = 1, b3 = 1), algorithm = "port")
nls(formula = weed ~ 100 * c1 / (1 + 10 * c2 * exp(-0.1 * c3 * tt)),
     data = weeddf, start = c(c1 = 1, c2 = 1, c3 = 1), algorithm = "port")
nls(formula = weed ~ Asym /(1 + exp((xmid - tt) / scal)),data = weeddf,
     start = c(Asym = 1, xmid = 1, scal = 1), algorithm = "port")
### gsl_nls Logistic3T setup
gslnls::gsl_nls(fn = weed ~ Asym / (1 + exp((xmid - tt) / scal)),data = weeddf,
    start = c(Asym = 1, xmid = 1, scal = 1), algorithm = "lmaccel")
gslnls::gsl_nls(fn = weed ~ Asym / (1 + exp((xmid - tt) / scal)),data = weeddf,
    start = c(Asym = 1, xmid = 1, scal = 1),algorithm = "dogleg")
gslnls::gsl_nls(fn = weed ~ Asym / (1 + exp((xmid - tt) / scal)),data = weeddf,
    start = c(Asym = 1, xmid = 1, scal = 1), algorithm = "ddogleg")
gslnls::gsl_nls(fn = weed ~ Asym / (1 + exp((xmid - tt) / scal)),data = weeddf,
    start = c(Asym = 1, xmid = 1, scal = 1),algorithm = "subspace2D")
gslnls::gsl_nls(fn = weed ~ Asym / (1 + exp((xmid - tt) / scal)),data = weeddf,
    start = c(Asym = 1, xmid = 1, scal = 1),control = list(scale = "levenberg")

Response 2:11  We had missed the "scale" setting i.e., had looked at the documentation
    of `gsl_nls()` but not `gsl_nls_control()`, and in particular the original
    Gnu Scientific Library documentation. It turns out the default scaling for the
    `gsl_nls()` code is the value of the maximum diagonal element of J'J. (J. Mor'e).
    Marquardt scaling is the diagonal of J'J, while Levenberg is a unit matrix.
    nlsr uses a linear combination of Levenberg and Marquardt as described in 
    Nash, John C., Minimizing a nonlinear sum of squares function on a small computer,
    Journal of the Institute for Mathematics and its Applications, 1977,
    volume 19, pages 231-237. Details are unfortunately important.

## Penalty function I [More, Garbow, Hillstrom 1981]
alpha <- 1e-5
n <- 1000
### residuals
fres <- function(x) {
c(sqrt(alpha) * (x - 1), sum(xˆ2) - 0.25)
}
### jacobian
fjac <- function(x) {
jj <- rbind(diag(sqrt(alpha), nrow = length(x)), 2 * t(x))
attr(jj, "gradient") <- jj
jj
}
bench::mark(
"large problem" = nlsr::nlfb(start = 1:n, resfn = fres, jacfn = fjac),
iterations = 1
)

References

Moré, Jorge J., Burton S. Garbow, and Kenneth E. Hillstrom. 1981. “Testing Unconstrained Optimization Software.” ACM Transactions on Mathematical Software 7: 17–41.