### 2023-07-16 RefactoringNLS/revision202307/RJournal2023-15Rev1Resubmission/StateOfRToolsForNLS.tex
Prep for resubmission

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The Current State of R Tools for Nonlinear Least Squares Modeling},
  pdfauthor={John C. Nash ; Arkajyoti Bhattacharjee },
  colorlinks=true,
  linkcolor={red},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

\title{The Current State of R Tools for Nonlinear Least Squares
Modeling}
\author{John C. Nash
\thanks{ retired professor, Telfer School of Management, University of Ottawa} \and Arkajyoti
Bhattacharjee
\thanks{Department of Mathematics and Statistics, Indian Institute of Technology, Kanpur}}
\date{2023-7-16}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

Our Google Summer of Code project ``Improvements to \texttt{nls()}''
considered the features and limitations of the R function \texttt{nls()}
with the aim of improving and rationalizing R tools for nonlinear
regression. The rich features of \texttt{nls()} are weakened by several
deficiencies and inconsistencies such as a lack of stabilization of the
Gauss-Newton solver. Further considerations are the usability and
maintainability of the code base that provides the functionality
\texttt{nls()} offers. Various packages, in particular \texttt{nlsr}
(John C Nash and Duncan Murdoch
(\protect\hyperlink{ref-nlsr2023manual}{2023})), provide alternative
capabilities. We consider the differences in goals, approaches, and
features of different tools for nonlinear least squares modeling in R.
Discussion of these matters is relevant to improving R generally as well
as its nonlinear estimation tools.

\hypertarget{the-nls-function}{%
\subsection{\texorpdfstring{The \texttt{nls()}
function}{The nls() function}}\label{the-nls-function}}

\texttt{nls()} is the tool for estimating nonlinear statistical models
in base R, the primary software distribution from the Comprehensive R
Archive Network (CRAN). Users do not have to request that \texttt{nls()}
be loaded. The function dates to the 1980s and the work related to D. M.
Bates and Watts (\protect\hyperlink{ref-bateswatts}{1988}) and others in
the S programming language.

The \texttt{nls()} function has a remarkable and comprehensive set of
capabilities for estimating nonlinear models that are expressed as
\textbf{formulas}. In particular, we note that it

\begin{itemize}
\tightlist
\item
  handles formulas that include R functions, even ones that call
  calculations in other programming languages;
\item
  allows data to be weighted or subset;
\item
  can estimate bound-constrained parameters;
\item
  provides a mechanism for handling partially linear models;
\item
  permits parameters to be indexed over a set of related data;
\item
  produces measures of variability (i.e., standard error estimates) for
  the estimated parameters;
\item
  has related profiling capabilities for exploring the likelihood
  surface as parameters are changed;
\item
  links to many pre-coded (\texttt{selfStart}) models that do not
  require initial parameter values.
\end{itemize}

With such a range of features and long history, the code has become
untidy and overly patched, difficult to maintain, and its underlying
methods could be improved. Various workers have developed packages to
overcome these concerns, and we will address some of these here.

\hypertarget{scope-of-our-comparison}{%
\subsection{Scope of our comparison}\label{scope-of-our-comparison}}

Our Google Summer of Code project for 2021, ``Improvements to
\texttt{nls()}'' had the goal of providing a version of the code that
would address some of the deficiencies discovered over the long period
that \texttt{nls()} has been part of R. However, it was also critical to
preserve legacy functionality, and we did not achieve a result that
provided desired improvements and preserved legacy results
simultaneously. On the other hand, working with \texttt{nls()} suggested
ways package \texttt{nlsr} could be improved in features, while also
revealing how it has a very different internal structure and design.

The primary goal of this paper is to present some of these issues, with
a limited consideration of some other R tools that exist for nonlinear
least squares calculations. We hope to encourage others to engage,
possibly with us, to find ways to bring improvements and to consolidate
tools where reasonable. We want to ease the task of users in choosing
which tools to use. Unfortunately, it is simpler to create new packages
than to bring together and simplify existing ones.

\hypertarget{tools-considered}{%
\subsubsection{Tools considered}\label{tools-considered}}

Besides the base-R \texttt{nls()} function, we consider some packages
that are available in the CRAN repository. Of these, we will pay
particular attention to \textbf{nlsr} (John C Nash and Duncan Murdoch
(\protect\hyperlink{ref-nlsr2023manual}{2023})), \textbf{minpack.lm}
(Elzhov et al. (\protect\hyperlink{ref-minpacklm12}{2012})), and
\textbf{gslnls} (Chau (\protect\hyperlink{ref-gslnls23}{2023})) which
are general nonlinear least-squares solvers. We will not pursue tools in
the Bioconductor (Gentleman et al.
(\protect\hyperlink{ref-Bioconductor}{2004})) collection, nor those on
repositories such as GitHub and Gitlab.

\hypertarget{an-illustrative-example}{%
\subsection{An illustrative example}\label{an-illustrative-example}}

The Hobbs weed infestation problem (John C. Nash
(\protect\hyperlink{ref-cnm79}{1979, 120})) is a growth curve modeling
task that seems straightforward, but is quite nasty. Its very succinct
statement provides the ``short reproducible example'' much requested on
R mailing lists. This is a real problem from a field researcher, though
the units of measurement and background information are not available.
The data and graph follow.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{weed }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{5.308}\NormalTok{, }\FloatTok{7.24}\NormalTok{, }\FloatTok{9.638}\NormalTok{, }\FloatTok{12.866}\NormalTok{, }\FloatTok{17.069}\NormalTok{, }\FloatTok{23.192}\NormalTok{, }\FloatTok{31.443}\NormalTok{,}
          \FloatTok{38.558}\NormalTok{, }\FloatTok{50.156}\NormalTok{, }\FloatTok{62.948}\NormalTok{, }\FloatTok{75.995}\NormalTok{, }\FloatTok{91.972}\NormalTok{)}
\NormalTok{tt }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{12}
\NormalTok{weeddf }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(tt, weed)}
\FunctionTok{plot}\NormalTok{(weeddf, }\AttributeTok{main=}\StringTok{"Hobbs weed infestation data"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{StateOfRToolsForNLS_files/figure-latex/ex01-1.pdf}

Three suggested models for this data are (with names to allow for easy
reference)

\textbf{Logistic3U}: \[  y \approx  b_1 / (1 + b_2 * exp(- b_3 * t)) \]

\textbf{Logistic3S}:
\[ y \approx  100 * c_1 / (1 + 10 * c_2 * exp(- 0.1 * c_3 * t)) \]

\textbf{Logistic3T}: \[ y \approx Asym / (1 + exp((xmid - t)/scal)) \]

where we will use \texttt{weed} for \(y\) and \texttt{tt} for \(t\). The
functions above are equivalent, but the first is generally more awkward
to solve numerically due to its poor scaling. The parameters of the
three forms are related as follows:

\[   Asym =  b_1 = 100 * c_1 \] \[ exp(xmid/scal)  =  b_2 = 10 * c_2 \]
\[ 1/scal  =  b_3 \] To allow for simpler discussion, let us say that
the parameters form a (named) vector \(p\) and the model function is
called \(model(p)\). The residuals can be written either as
\(r_1 = y - model(p)\) or \(r = model(p) - y\) since their sum of
squares has an identical value. The second form avoids a potential sign
error if we need to evaluate derivatives.

We wish to minimize the sum of squared residuals, which is our loss (or
objective) function. Starting with some guess for the parameters, we aim
to alter these parameters to obtain a smaller loss function. We then
iterate until we can make no further progress.

Let us consider there are \(n\) parameters and \(m\) residuals. The loss
function is \[ S(p) = r' r = \sum_{i=1}^m { r_i^2 }\] The gradient of
\(S(p)\) is \[ g = 2 * J' r\] where the Jacobian \(J\) is given by
elements \[ J_{i,j} = \partial r_i / \partial p_j \] and the Hessian is
defined by elements
\[ H_{i,j} = \partial ^2 S(p) / {\partial p_i \partial p_j} \] If we
expand the Hessian for nonlinear least squares problems, we find
\[  H_{i,j}/2 = \sum_{k = 1}^m{ J_{k,i} J_{k,j}}  + \sum_{k = 1}^m {r_k * \partial{r_k}/{\partial p_i \partial p_j}}\]
Let us use \(D_{i,j}\) for the elements of the second term of this
expression. What is generally called \textbf{Newton's method} for
function minimization tries to set the gradient to zero (to find a
stationary point of the function \(S(p)\)). This leads to
\textbf{Newton's equation} \[ H \delta  = -g \] Given a set of
parameters \(p\), which provide \(H\) and \(g\), we solve this equation
for \(\delta\), adjust \(p\) to \(p + \delta\) and iterate, hopefully to
converge on a solution. Applying this to a sum of squares problem gives
\[ H \delta / 2 = (J'J + D) \delta = - J' r\] In this expression, only
the elements of \(D\) have second partial derivatives. Gauss, attempting
to model planetary orbits, had small residuals, and noted that these
multiplied the second partial derivatives of \(r\), so he approximated
\[ H / 2 \approx J' J\] by assuming \(D \approx 0\). This results in the
Gauss-Newton method where we solve \[ J' J \delta = - J' r\] though we
can avoid some loss of accuracy by not forming the inner product matrix
\(J' J\) and solving the linear least squares matrix problem
\[ J \delta \approx -r \] by one of several matrix decomposition
methods. On the other hand, when \(m >> n\), the matrix \(J' J\) uses
much less storage than \(J\).

In reality, there are many problems where \(D\) should not be ignored,
but the work to compute it precisely is considerable. Many work-arounds
have been proposed, of which the Levenberg-Marquardt stabilization
(Levenberg (\protect\hyperlink{ref-Levenberg1944}{1944}), Marquardt
(\protect\hyperlink{ref-Marquardt1963}{1963})) is the most commonly
used. For convenience, we will use ``Marquardt'', as we believe he first
incorporated the ideas into a practical computer program.

The usual suggestion is that \(D\) be replaced by a multiple of the unit
matrix or else a multiple of the diagonal part of \(J' J\). In low
precision, some elements of \(J' J\) could underflow to zero (John C.
Nash (\protect\hyperlink{ref-jn77ima}{1977})), and a linear combination
of both choices is an effective compromise. Various choices for \(D\),
as well as a possible line search along the direction \(\delta\) rather
than a unit step (Hartley (\protect\hyperlink{ref-Hartley1961}{1961})),
give rise to several variant algorithms. ``Marquardt's method'' is a
family of methods. Fortunately, most choices work well.

\hypertarget{problem-setup}{%
\subsubsection{Problem setup}\label{problem-setup}}

Let us specify in R the three model formulas and set some starting
values for parameters. These starting points are not equivalent and are
deliberately crude choices. Workers performing many calculations of a
similar nature should try to provide good starting points to reduce
computation time and avoid finding a false solution.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# model formulas}
\NormalTok{frmu }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ b1 }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ b2 }\SpecialCharTok{*} \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{b3 }\SpecialCharTok{*}\NormalTok{ tt))}
\NormalTok{frms }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}} \DecValTok{100}\SpecialCharTok{*}\NormalTok{ c1 }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \DecValTok{10} \SpecialCharTok{*}\NormalTok{ c2 }\SpecialCharTok{*} \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\FloatTok{0.1} \SpecialCharTok{*}\NormalTok{c3 }\SpecialCharTok{*}\NormalTok{ tt))}
\NormalTok{frmt }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}}\NormalTok{ tt) }\SpecialCharTok{/}\NormalTok{ scal))}
\CommentTok{\#}
\CommentTok{\# Starting parameter sets}
\NormalTok{stu1}\OtherTok{\textless{}{-}}\FunctionTok{c}\NormalTok{(}\AttributeTok{b1 =}\DecValTok{1}\NormalTok{, }\AttributeTok{b2 =} \DecValTok{1}\NormalTok{, }\AttributeTok{b3 =} \DecValTok{1}\NormalTok{)}
\NormalTok{sts1}\OtherTok{\textless{}{-}}\FunctionTok{c}\NormalTok{(}\AttributeTok{c1 =} \DecValTok{1}\NormalTok{, }\AttributeTok{c2 =} \DecValTok{1}\NormalTok{, }\AttributeTok{c3 =} \DecValTok{1}\NormalTok{)}
\NormalTok{stt1}\OtherTok{\textless{}{-}}\FunctionTok{c}\NormalTok{(}\AttributeTok{Asym =} \DecValTok{1}\NormalTok{, }\AttributeTok{xmid =} \DecValTok{1}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

One of the useful features of \texttt{nls()} is the possibility of a
\texttt{selfStart} model, where starting parameter values are not
required. However, if a \texttt{selfStart} model is not available,
\texttt{nls()} sets all the starting parameters to 1. This is tolerable,
but could be improved by using a set of values that are all slightly
different, which, in the case of the example model
\(y \,\sim\, a \,*\, exp(-b \,*\, x) + c\,*\,exp(-d \,*\, x)\) would
avoid a singular Jacobian because \(b\) and \(d\) were equal in value.
Program modifications to give a sequence like 1.0, 1.1, 1.2, 1.3 for the
four parameters are fairly obvious.

It is also possible to provide R functions for the residual and
Jacobian. \texttt{nls()} is not well-suited to using such a formulation,
but \texttt{nlsr::nlfb()} and \texttt{minpack.lm::nls.lm()} work
specifically with such a formulation. This is usually more work for the
user if the formula setup is possible. To illustrate, we show the
functions for the unscaled 3 parameter logistic.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Logistic3U}
\NormalTok{hobbs.res  }\OtherTok{\textless{}{-}}  \ControlFlowTok{function}\NormalTok{(x)\{ }\CommentTok{\# scaled Hobbs weeds problem {-}{-} residual}
  \CommentTok{\# This variant uses looping}
  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{length}\NormalTok{(x) }\SpecialCharTok{!=} \DecValTok{3}\NormalTok{) }\FunctionTok{stop}\NormalTok{(}\StringTok{"hobbs.res {-}{-} parameter vector n!=3"}\NormalTok{)}
\NormalTok{  y  }\OtherTok{\textless{}{-}}  \FunctionTok{c}\NormalTok{(}\FloatTok{5.308}\NormalTok{, }\FloatTok{7.24}\NormalTok{, }\FloatTok{9.638}\NormalTok{, }\FloatTok{12.866}\NormalTok{, }\FloatTok{17.069}\NormalTok{, }\FloatTok{23.192}\NormalTok{, }\FloatTok{31.443}\NormalTok{, }
           \FloatTok{38.558}\NormalTok{, }\FloatTok{50.156}\NormalTok{, }\FloatTok{62.948}\NormalTok{, }\FloatTok{75.995}\NormalTok{, }\FloatTok{91.972}\NormalTok{)}
\NormalTok{  tt  }\OtherTok{\textless{}{-}}  \DecValTok{1}\SpecialCharTok{:}\DecValTok{12}
\NormalTok{  res  }\OtherTok{\textless{}{-}}\NormalTok{  x[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ x[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{*} \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{x[}\DecValTok{3}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ tt)) }\SpecialCharTok{{-}}\NormalTok{ y}
\NormalTok{\}}

\NormalTok{hobbs.jac  }\OtherTok{\textless{}{-}}  \ControlFlowTok{function}\NormalTok{(x) \{ }\CommentTok{\# scaled Hobbs weeds problem {-}{-} Jacobian}
\NormalTok{  jj  }\OtherTok{\textless{}{-}}  \FunctionTok{matrix}\NormalTok{(}\FloatTok{0.0}\NormalTok{, }\DecValTok{12}\NormalTok{, }\DecValTok{3}\NormalTok{)}
\NormalTok{  tt  }\OtherTok{\textless{}{-}}  \DecValTok{1}\SpecialCharTok{:}\DecValTok{12}
\NormalTok{  yy  }\OtherTok{\textless{}{-}}  \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{x[}\DecValTok{3}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ tt)}
\NormalTok{  zz  }\OtherTok{\textless{}{-}}  \FloatTok{1.0} \SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ x[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ yy)}
\NormalTok{  jj[tt,}\DecValTok{1}\NormalTok{]   }\OtherTok{\textless{}{-}}\NormalTok{   zz}
\NormalTok{  jj[tt,}\DecValTok{2}\NormalTok{]   }\OtherTok{\textless{}{-}}   \SpecialCharTok{{-}}\NormalTok{x[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ zz }\SpecialCharTok{*}\NormalTok{ zz }\SpecialCharTok{*}\NormalTok{ yy}
\NormalTok{  jj[tt,}\DecValTok{3}\NormalTok{]   }\OtherTok{\textless{}{-}}\NormalTok{   x[}\DecValTok{1}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ zz }\SpecialCharTok{*}\NormalTok{ zz }\SpecialCharTok{*}\NormalTok{ yy }\SpecialCharTok{*}\NormalTok{ x[}\DecValTok{2}\NormalTok{] }\SpecialCharTok{*}\NormalTok{ tt}
  \FunctionTok{attr}\NormalTok{(jj, }\StringTok{"gradient"}\NormalTok{) }\OtherTok{\textless{}{-}}\NormalTok{ jj}
\NormalTok{  jj}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{estimation-of-models-specified-as-formulas}{%
\subsection{Estimation of models specified as
formulas}\label{estimation-of-models-specified-as-formulas}}

Using a formula specification was a principal advantage made with
\texttt{nls()} when it became available in S sometime in the 1980s. It
uses a Gauss-Newton (i.e., unstabilized) iteration with a step reduction
line search (Hartley (\protect\hyperlink{ref-Hartley1961}{1961})). This
works very efficiently as long as \(J\) is not ill-conditioned. Below we
see the default \texttt{nls()} algorithm does poorly on the example
problem. To save page space, we use 1-line result display functions from
package \texttt{nlsr}, namely \texttt{pnls()} and \texttt{pshort()}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unls1}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmu, }\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Error in nls(formula = frmu, start = stu1, data = weeddf) : 
##   singular gradient
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{snls1}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Error in nls(formula = frms, start = sts1, data = weeddf) : 
##   singular gradient
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnls1}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmt, }\AttributeTok{start =}\NormalTok{ stt1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Error in nls(formula = frmt, start = stt1, data = weeddf) : 
##   singular gradient
\end{verbatim}

Here we see the infamous ``singular gradient'' termination message of
\texttt{nls()}. Technically, it is the Jacobian that is singular, but R
does not distinguish scalar and vector functions, which yield partial
derivatives that are a gradient or Jacobian respectively.

\texttt{nls()} also offers the option of using the ``port'' library
algorithm choice (Fox, Hall, and Schryer
(\protect\hyperlink{ref-FoxHallSchryer1978}{1978})). There is a warning
in the \texttt{nls()} documentation that this ``appears unfinished'',
but it is successful in solving all three cases.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unls1p}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmu, }\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{algorithm =} \StringTok{"port"}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(unls1p)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## unls1p  -- ss= 2.587277 :  b1 = 196.1863  b2 = 49.09164  b3 = 0.3135697; 17  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{snls1p}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{algorithm =} \StringTok{"port"}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(snls1p)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## snls1p  -- ss= 2.587277 :  c1 = 1.961863  c2 = 4.909164  c3 = 3.135697; 15  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnls1p}\OtherTok{\textless{}{-}}\FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmt, }\AttributeTok{start =}\NormalTok{ stt1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{algorithm =} \StringTok{"port"}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(tnls1p)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tnls1p  -- ss= 2.587277 :  Asym = 196.1863  xmid = 12.4173  scal = 3.189083; 12  itns
\end{verbatim}

\hypertarget{solution-attempts-with-nlsr}{%
\subsubsection{Solution attempts with
nlsr}\label{solution-attempts-with-nlsr}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unlx1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmu, }\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{print}\NormalTok{(unlx1) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## residual sumsquares =  2.5873  on  12 observations
##     after  19    Jacobian and  25 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## b1               196.186         11.31      17.35  3.167e-08  -4.859e-09        1011  
## b2               49.0916         1.688      29.08  3.284e-10  -3.099e-08      0.4605  
## b3               0.31357      0.006863      45.69  5.768e-12   2.305e-06     0.04714
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{snlx1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{pshort}\NormalTok{(snlx1) }\CommentTok{\# a short{-}form output}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## snlx1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 34 res/ 23 jac
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnlx1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmt, }\AttributeTok{start =}\NormalTok{ stt1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{pshort}\NormalTok{(tnlx1) }\CommentTok{\# alternatively print(tnlx1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tnlx1  -- ss= 2.5873 :  Asym = 196.19  xmid = 12.417  scal = 3.1891; 36 res/ 27 jac
\end{verbatim}

Though we have found solutions, the Jacobian is essentially singular as
shown by its singular values. Note that these are \textbf{displayed} by
package \texttt{nlsr} in a single column in the output to provide a
compact layout, but the values do \textbf{not} correspond to the
individual parameters in whose row they appear; they are a property of
the whole problem.

\hypertarget{solution-attempts-with-minpack.lm}{%
\subsubsection{Solution attempts with
minpack.lm}\label{solution-attempts-with-minpack.lm}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{unlm1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlsLM}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmu, }\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{pnls}\NormalTok{(unlm1)  }\CommentTok{\# Short form of output}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## unlm1  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 17  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{snlm1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlsLM}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{pnls}\NormalTok{(snlm1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## snlm1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 7  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tnlm1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlsLM}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmt, }\AttributeTok{start =}\NormalTok{ stt1, }\AttributeTok{data =}\NormalTok{ weeddf))}
\FunctionTok{pnls}\NormalTok{(tnlm1) }\CommentTok{\# short form to give sum of squares, else use summary(tnlm1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tnlm1  -- ss= 9205.4 :  Asym = 35.532  xmid = 43376  scal = -2935.4; 39  itns
\end{verbatim}

\hypertarget{solution-attempts-with-gslnls}{%
\subsubsection{Solution attempts with
gslnls}\label{solution-attempts-with-gslnls}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(gslnls)}
\NormalTok{ugslnls1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{gsl\_nls}\NormalTok{(}\AttributeTok{fn =}\NormalTok{ frmu, }\AttributeTok{data =}\NormalTok{ weeddf,  }\AttributeTok{start =}\NormalTok{ stu1))}
\FunctionTok{pnls}\NormalTok{(ugslnls1) }\CommentTok{\# to get sum of squares}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## ugslnls1  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{sgslnls1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{gsl\_nls}\NormalTok{(}\AttributeTok{fn =}\NormalTok{ frms, }\AttributeTok{data =}\NormalTok{ weeddf,  }\AttributeTok{start =}\NormalTok{ sts1))}
\FunctionTok{pnls}\NormalTok{(sgslnls1) }\CommentTok{\# Use summary() to get display}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## sgslnls1  -- ss= 2.5873 :  c1 = 1.9619  c2 = 4.9092  c3 = 3.1357; 9  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tgslnls1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{gsl\_nls}\NormalTok{(}\AttributeTok{fn =}\NormalTok{ frmt, }\AttributeTok{data =}\NormalTok{ weeddf,  }\AttributeTok{start =}\NormalTok{ stt1))}
\FunctionTok{pnls}\NormalTok{(tgslnls1) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## tgslnls1  -- ss= 9205.4 :  Asym = 35.532  xmid = 20846  scal = -1741.1; 47  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# Other algorithms}
\DocumentationTok{\#\# tgslnls1a\textless{}{-}try(gsl\_nls(fn = frmt, data = weeddf,  start = stt1, algorithm="lmaccel"))}
\DocumentationTok{\#\# pnls(tgslnls1a) }
\DocumentationTok{\#\# tgslnls1d\textless{}{-}try(gsl\_nls(fn = frmt, data = weeddf,  start = stt1, algorithm="dogleg"))}
\DocumentationTok{\#\# pnls(tgslnls1d) }
\DocumentationTok{\#\# tgslnls1dd\textless{}{-}try(gsl\_nls(fn = frmt, data = weeddf,  start = stt1, algorithm="ddogleg"))}
\DocumentationTok{\#\# pnls(tgslnls1dd) }
\DocumentationTok{\#\# tgslnls1s2\textless{}{-}try(gsl\_nls(fn = frmt, data = weeddf,  start = stt1, algorithm="subspace2D"))}
\DocumentationTok{\#\# pnls(tgslnls1s2) }
\end{Highlighting}
\end{Shaded}

While default \texttt{gsl\_nls()} gives much of the same results as
\texttt{nlsLM()}, uncommenting the calls to the alternative algorithms
shows it can get the global minimum with any of these choices.

\hypertarget{comparison-notes-for-formula-setup-solutions}{%
\subsubsection{Comparison notes for formula-setup
solutions}\label{comparison-notes-for-formula-setup-solutions}}

\texttt{nlsr::nlxb()} uses \texttt{print()} to output standard errors
and singular values of the Jacobian (for diagnostic purposes). By
contrast, \texttt{minpack.lm::nlsLM()} and \texttt{nls()} use
\texttt{summary()}, which does not display the sum of squares, while
\texttt{print()} gives the sum of squares, but not the standard error of
the residuals. Such behaviour is consistent with other modeling
functions such as \texttt{lm()}.

The importance of the singular values is to allow us to gauge how
``nearly singular'' the Jacobian is at the solution, and the ratio of
the smallest to largest of the singular values is a simple but effective
measure. The ratios are 4.6641e-05 for Logistic3U, 0.021022 for
Logistic3S, and 0.001055 for Logistic3T; so Logistic3S is the ``least
singular''.

The results from \texttt{nlsLM} and \texttt{gsl\_nls} for the
transformed model Logistic3T have a very large sum of squares, which may
suggest that these programs have failed. Since \texttt{nls()},
\texttt{nlsLM()}, and \texttt{gsl\_nls()} do not offer singular values,
we need to extract the Jacobian and compute its singular values. The
following script shows how to do this, using as Jacobian the
\texttt{gradient} element in the returned solution for these solvers.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# for nlsLM}
\ControlFlowTok{if}\NormalTok{ (}\FunctionTok{inherits}\NormalTok{(tnlm1, }\StringTok{"try{-}error"}\NormalTok{))  \{}
   \FunctionTok{print}\NormalTok{(}\StringTok{"Cannot compute solution {-}{-} likely singular Jacobian"}\NormalTok{)}
\NormalTok{ \} }\ControlFlowTok{else}\NormalTok{ \{  }
\NormalTok{   JtnlsLM }\OtherTok{\textless{}{-}}\NormalTok{ tnlm1}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{gradient}\NormalTok{() }\CommentTok{\# actually the Jacobian}
   \FunctionTok{svd}\NormalTok{(JtnlsLM)}\SpecialCharTok{$}\NormalTok{d }\CommentTok{\# Singular values}
\NormalTok{\}   }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 3.4641e+00 3.2530e-10 8.9643e-12
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# for gsl\_nls}
\ControlFlowTok{if}\NormalTok{ (}\FunctionTok{inherits}\NormalTok{(tgslnls1, }\StringTok{"try{-}error"}\NormalTok{)) \{}
   \FunctionTok{cat}\NormalTok{(}\StringTok{"Cannot compute solution {-}{-} likely singular Jacobian"}\NormalTok{)}
\NormalTok{\} }\ControlFlowTok{else}\NormalTok{ \{  }
\NormalTok{   JtnlsLM }\OtherTok{\textless{}{-}}\NormalTok{ tgslnls1}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{gradient}\NormalTok{()}
   \FunctionTok{svd}\NormalTok{(JtnlsLM)}\SpecialCharTok{$}\NormalTok{d }\CommentTok{\# Singular values}
\NormalTok{\}   }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] 3.4641e+00 9.4495e-09 4.1324e-11
\end{verbatim}

We see that there are differences in detail, but the more important
result is that two out of three singular values are essentially 0. Our
Jacobian is singular, and no method of the Gauss-Newton type should be
able to continue. Indeed, from the parameters reported at this saddle
point, \texttt{nlsr::nlxb()} cannot proceed.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{stspecial }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{Asym =} \FloatTok{35.532}\NormalTok{,  }\AttributeTok{xmid =} \DecValTok{43376}\NormalTok{,  }\AttributeTok{scal =} \SpecialCharTok{{-}}\FloatTok{2935.4}\NormalTok{)}
\NormalTok{badstart }\OtherTok{\textless{}{-}} \FunctionTok{nlxb}\NormalTok{(}\AttributeTok{formula =}\NormalTok{ frmt, }\AttributeTok{start =}\NormalTok{ stspecial, }\AttributeTok{data =}\NormalTok{ weeddf)}
\FunctionTok{print}\NormalTok{(badstart)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## residual sumsquares =  9205.4  on  12 observations
##     after  2    Jacobian and  2 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## Asym             35.5321            NA         NA         NA  -9.694e-09       3.464  
## xmid               43376            NA         NA         NA  -1.742e-09    2.61e-10  
## scal             -2935.4            NA         NA         NA    -2.4e-08    7.12e-16
\end{verbatim}

We also note that changing the start to

\begin{verbatim}
stt2 <- c(Asym=1, xmid=1, scal=2)
\end{verbatim}

allows both \texttt{nlsLM} and \texttt{gsl\_nls} to find the minimum.
\texttt{gsl\_nls} also succeeds when the particular form of the
Marquardt stabilization is changed from the default by the control
setting \texttt{scale\ =\ "levenberg"}.

We consider the code inside package \texttt{nlsr} to be competently
coded, with the advantage over \texttt{nls()} of the Levenberg-Marquardt
stabilization in common with \texttt{minpack.lm} and \texttt{glsnls}.
The success rate is high, but \texttt{nlsr} has its premature
terminations too. However, it provides diagnostic information in the
form of the singular values of the Jacobian that allows us to try
alternative starts, algorithms or programs in cases where there are
indications of early termination.

\hypertarget{functional-specification-of-problems}{%
\subsection{Functional specification of
problems}\label{functional-specification-of-problems}}

Let us write a slightly modified form of the nonlinear least squares
problem:

\[ S(p) = \sum_{i=1}^m { r_i(p)^2 }\]

R users are likely to think almost always that the functions \(r_i\) are
all of the same form, differing in the details of the data they use.
However, the famous Rosenbrock banana-shaped valley (Rosenbrock
(\protect\hyperlink{ref-Rosenbrock1960}{1960})) has
\[ r_1(p) = (1 - p_1)\] \[ r_2(p) = 10 (p_2 - {p_1}^2)\]

Thus, there can be problems where the idea of a ``model'' or a
``prediction'' makes no sense, though the nonlinear least squares
problem is still well-posed. In our code, we still refer to the
functions as ``residuals'', though this is purely a convenience.

We illustrate how to solve nonlinear least squares problems using a
function to define the residual. Note that \texttt{gsl\_nls()} requires
a vector \texttt{y} that is the expected value of what we have called
the residual, but in this case is actually the model. \texttt{gsl\_nls}
uses a numerical approximation for the Jacobian if the argument
\texttt{jac} is missing. Note functions \texttt{nlsr::pnls()} and
\texttt{nslr::pnlslm()} for a 1-line display of the results.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hobnlfb }\OtherTok{\textless{}{-}} \FunctionTok{nlfb}\NormalTok{(}\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{resfn =}\NormalTok{ hobbs.res, }\AttributeTok{jacfn =}\NormalTok{ hobbs.jac)}
\FunctionTok{pshort}\NormalTok{(hobnlfb) }\CommentTok{\# use print(hobnlfb) for more detail}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## hobnlfb  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25 res/ 19 jac
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hobnlm }\OtherTok{\textless{}{-}} \FunctionTok{nls.lm}\NormalTok{(}\AttributeTok{par =}\NormalTok{ stu1, }\AttributeTok{fn =}\NormalTok{ hobbs.res, }\AttributeTok{jac =}\NormalTok{ hobbs.jac)}
\FunctionTok{pnlslm}\NormalTok{(hobnlm)  }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## hobnlm  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 17  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hobgsln }\OtherTok{\textless{}{-}} \FunctionTok{gsl\_nls}\NormalTok{(}\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{fn =}\NormalTok{ hobbs.res, }\AttributeTok{y =} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{12}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(hobgsln)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## hobgsln  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hobgsl }\OtherTok{\textless{}{-}} \FunctionTok{gsl\_nls}\NormalTok{(}\AttributeTok{start =}\NormalTok{ stu1, }\AttributeTok{fn =}\NormalTok{ hobbs.res, }\AttributeTok{y =} \FunctionTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{12}\NormalTok{), }\AttributeTok{jac =}\NormalTok{ hobbs.jac)}
\FunctionTok{pnls}\NormalTok{(hobgsl) }\CommentTok{\# using analytic Jacobian}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## hobgsl  -- ss= 2.5873 :  b1 = 196.19  b2 = 49.092  b3 = 0.31357; 25  itns
\end{verbatim}

\hypertarget{design-goals-termination-tests-and-output-objects}{%
\subsection{Design goals, termination tests, and output
objects}\label{design-goals-termination-tests-and-output-objects}}

The output object of \texttt{nlxb()} is smaller than the class
\texttt{nls} object returned by \texttt{nls()}, \texttt{nlsLM()}, and
\texttt{gsl\_nls()}. Package \texttt{nlsr} emphasizes the solution of
the nonlinear least squares problem rather than the estimation of a
nonlinear model that fits or explains the data. The object of class
\texttt{nls} allows for a number of specialized modeling and diagnostic
extensions. For compatibility, the \texttt{nlsr} package has the
function \texttt{wrapnlsr()}, for which \texttt{nlsr()} is an alias,
that is intended for problems specified as formulas. This uses
\texttt{nlxb()} to find good parameters, then calls \texttt{nls()} to
return the class \texttt{nls} object. Unless particular modeling
features are needed, the use of \texttt{wrapnlsr()} is unnecessary and
wasteful of resources.

The design goals of the different tools may also be revealed in the
so-called ``convergence tests'' for the iterative solvers. In the manual
page for \texttt{nls()} in R 4.0.0 there was the warning:

\begin{quote}
\emph{Do not use \texttt{nls} on artificial ``zero-residual'' data.}
\end{quote}

with suggested addition of small perturbations to the data. This admits
\texttt{nls()} could not solve well-posed problems unless data is
polluted with errors. Zero-residual problems are not always artificial,
since problems in function approximation and nonlinear equations can be
approached with nonlinear least squares. Fortunately, a small adjustment
to the ``termination test'' for the \textbf{program}, rather than for
the ``convergence'' of the underlying \textbf{algorithm}, fixes the
defect. The test is the Relative Offset Convergence Criterion (see
Bates, Douglas M. and Watts, Donald G.
(\protect\hyperlink{ref-BatesWatts81}{1981})). This scales an estimated
reduction in the loss function by its current value. If the loss
function is very small, we are close to a zero-divide. Adding a small
quantity to the divisor avoids trouble. In 2021, one of us (J. Nash)
proposed that \texttt{nls.control()} have an additional parameter
\texttt{scaleOffset} with a default value of zero. Setting it to a small
number -- 1.0 is a reasonable choice -- allows small-residual problems
(i.e., near-exact fits) to be dealt with easily. We call this the
\textbf{safeguarded relative offset convergence criterion}, and it has
been in \texttt{nlsr} since it was introduced. The default value gives
\texttt{nls()} its legacy behaviour. This improvement has been in the R
distributed code since version 4.1.0.

Additional termination tests can be used. \texttt{nlsr} has a
\textbf{small sum of squares} test (\texttt{smallsstest}) that compares
the latest evaluated sum of squared (weighted) residuals to \texttt{e4}
times the initial sum of squares, where
\texttt{e4\ \textless{}-\ (100\ *\ .Machine\$double.eps)\ \^{}\ 4} is
approximately 2.43e-55.

Termination after what may be considered excessive computation is also
important. \texttt{nls()} stops after \texttt{maxiter} ``iterations''.
The meaning of ``iteration'' may require an examination of the code.
\texttt{nlsr} terminates when the number of residual or Jacobian
evaluations (\texttt{feval} and \texttt{jeval}) exceed set limits, and
we prefer larger limits (\texttt{jemax\ =\ 5000} is the default) than
the \texttt{maxiter\ =\ 50} of \texttt{nls()} to avoid stopping early.
This may cause overly long run-times for \texttt{nlsr}, but our work
often involves problems that are sufficiently difficult that we want to
be sure we do not terminate early.

\hypertarget{returned-results-of-nls-and-other-tools}{%
\subsubsection{Returned results of nls() and other
tools}\label{returned-results-of-nls-and-other-tools}}

As mentioned, the output of \texttt{nls()} is an object of class ``nls''
which has a quite rich structure described in the manual file or
revealed by applying the \texttt{str()} function to the result of
\texttt{nls()}.

The complexity of this object is a challenge to users. Let us use
\texttt{result} as the name of the returned object from \texttt{nls()},
\texttt{minpack.lm::nlsLM()} or, \texttt{gslnls::gsl\_nls()}. For
example, the \texttt{data} return element is an R symbol. If the user's
environment exists, which it may not, we could see this data via the
syntax:

\begin{verbatim}
eval(parse(text = result$data))
\end{verbatim}

However, since the evaluating environment might not exist, a better
approach is to make the call to \texttt{nls()} with
\texttt{model\ =\ TRUE}, which returns an element \texttt{model} which
contains the data, and we can list its contents using

\begin{verbatim}
ls(result$model)
\end{verbatim}

and if there is an element called \texttt{xdata}, then it can be
accessed as \texttt{result\$model\$xdata}.

By contrast, \texttt{nlsr::nlxb()} returns a much simpler structure of
11 items in one level. Moreover, \texttt{nlxb} explicitly returns the
sum of squares, the residual vector, Jacobian, and counts of
evaluations, but does not include the input data.

\hypertarget{when-to-compute-ancillary-information}{%
\subsubsection{When to compute ancillary
information}\label{when-to-compute-ancillary-information}}

Tools that produce a class \texttt{nls} output object create a rich set
of functions and structures that are then used in a variety of modeling
tasks, including the least squares solution. By contrast, \texttt{nlsr}
computes quantities as they are requested or needed, with additional
features in separate functions. For example, the singular values of the
Jacobian are actually computed in the \texttt{print} and
\texttt{summary} methods for the result. These two approaches lead to
different consequences for performance and how features are provided.
\texttt{nlsr} has antecedents in the methods of John C. Nash
(\protect\hyperlink{ref-cnm79}{1979}), where storage for data and
programs was at a ridiculous premium in the small computers of the era.
Thus the code in \texttt{nlsr} is likely of value for workers to copy
and modify for customized tools, with code in R to support such uses.

\hypertarget{jacobian-calculation}{%
\subsection{Jacobian calculation}\label{jacobian-calculation}}

Gauss-Newton/Marquardt methods all need a Jacobian matrix at each
iteration. By default, \texttt{nlsr::nlxb()} will try to evaluate this
using analytic expressions using symbolic and automatic differentiation
tools. When using a formula specification of the model, \texttt{nls()},
\texttt{minpack.lm::nlsLM()} and \texttt{gslnls::gsl\_nls()} use a
finite difference approximation to compute the Jacobian, though
\texttt{gsl\_nls()} does have an option to attempt symbolic expressions.
Package \texttt{nlsr} provides, via appropriate calling syntax, four
numeric approximation options for the Jacobian, with a further control
\texttt{ndstep} for the size of the step used in the approximation.
These features are mainly to allow for investigation of approximate
derivatives. Regular users will rarely need them.

Using the \texttt{gradient} attribute of the output of the Jacobian
function to hold the Jacobian matrix lets us embed this in the
\textbf{residual} function as well, so that the call to
\texttt{nlsr::nlfb()} can be made with the same name used for both
residual and Jacobian function arguments. This programming trick saves a
lot of trouble for the package developer, but it can be a nuisance for
users trying to understand the code.

As far as we can understand the logic in \texttt{nls()}, the Jacobian
computation during parameter estimation is carried out within the called
C-language program and its wrapper R code function
\texttt{numericDeriv()}, part of \texttt{./src/library/stats/R/nls.R} in
the R distribution source code. This is used to provide Jacobian
information in the \texttt{nlsModel()} and \texttt{nlsModel.plinear()}
functions, which are \textbf{not} exported for general use.
\texttt{gsl\_nls()} also appears to use \texttt{numericDeriv()}.

\texttt{numericDeriv()} uses a simple forward difference approximation
of derivatives, though a central difference approximation can be
specified in control parameters. We are unclear why
\texttt{numericDeriv()} in base R calls \texttt{C\_numeric\_deriv}, as
we were easily able to create a more compact version entirely in R. See
\emph{\url{https://github.com/nashjc/RNonlinearLS/tree/main/DerivsNLS}}.

\texttt{minpack.lm::nlsLM()} invokes \texttt{numericDeriv()} in its
local version of \texttt{nlsModel()}, but it appears to use an internal
approximate Jacobian code from the original Fortran \texttt{minpack}
code, namely, \texttt{lmdif.f}. Such differences in approach can lead to
different behaviour, usually minor, but sometimes annoying with
ill-conditioned problems.

\begin{itemize}
\item
  A pasture regrowth problem (Huet et al.
  (\protect\hyperlink{ref-Huet2004}{2004}), page 1, based on Ratkowsky
  (\protect\hyperlink{ref-Ratkowsky1983}{1983})) has a poorly
  conditioned Jacobian and \texttt{nls()} fails with ``singular
  gradient''. Worse, numerical approximation to the Jacobian gives the
  error ``singular gradient matrix at initial parameter estimates'' for
  \texttt{minpack.lm::nlsLM} so that the Marquardt stabilization is
  unable to take effect, while the analytic derivatives of
  \texttt{nlsr::nlxb} give a solution.
\item
  Karl Schilling (private communication) provided an example where a
  model specified with the formula
  \texttt{y\ \textasciitilde{}\ a\ *\ (x\ ˆ\ b)} causes
  \texttt{nlsr::nlxb} to fail because the partial derivative w.r.t.
  \texttt{b} is \texttt{a\ *\ (x\^{}b\ *\ log(x))}. If there is data for
  which \texttt{x\ =\ 0}, the derivative is undefined, but the model can
  be computed. In such cases, we observed that \texttt{nls()} and
  \texttt{minpack.lm::nlsLM} found a solution through the lucky accident
  that the approximate derivative can be evaluated.
\end{itemize}

\hypertarget{jacobian-code-in-selfstart-models}{%
\subsubsection{Jacobian code in selfStart
models}\label{jacobian-code-in-selfstart-models}}

Analytic Jacobian code can be provided to all the solvers discussed.
Most \texttt{selfStart} models that automatically provide starting
parameters also include such code. There is documentation in R of
\texttt{selfStart} models, but their construction is non-trivial. A
number of such models are included with base R in
\texttt{./src/library/stats/R/zzModels.R}, with package \texttt{nlraa}
(Miguez (\protect\hyperlink{ref-MiguezNLRAA2021}{2021})) providing a
richer set. There are also some in the now-archived package
\texttt{NRAIA}. The Jacobian is provided via the \texttt{gradient}
attribute of the ``one-sided'' formula that defines each model, and
these Jacobians are often the analytic forms.

The \texttt{nls()} function, after computing the ``right-hand side'' or
\texttt{rhs} of the residual, checks to see if the \texttt{gradient}
attribute is defined, otherwise using \texttt{numericDeriv()} to compute
a Jacobian into that attribute. This code is within the
\texttt{nlsModel()} or \texttt{nlsModel.plinear()} functions. The use of
analytic Jacobians almost certainly contributes to the good performance
of \texttt{nls()} on \texttt{selfStart} models.

The use of \texttt{selfStart} models with \texttt{nlsr} is described in
the ``Introduction to nlsr'' vignette. However, since \texttt{nlsr}
generally can use very crude starting values, we have rarely needed
them, though it should be pointed out that our work is primarily
diagnostic. If we were carrying out a large number of similar
estimations, such initial parameters are critical to efficiency.

In considering \texttt{selfStart} models, we noted that the base-R
function \texttt{SSlogis} is intended to solve problem
\textbf{Logistic3T} above. When this function is used via
\texttt{getInitial()} to find starting values, it actually calls
\texttt{nls()} with the `plinear' algorithm and finds a (full) solution.
It then passes the solution coefficients to the default algorithm which
converges in one iteration and builds the standard output. We found the
implicit double-call a challenge to unravel. A more conventional
approach to starting parameters, following the lead of Ratkowsky
(\protect\hyperlink{ref-Ratkowsky1983}{1983}), is in the function
\texttt{SSlogisJN}, now part of the package \texttt{nlsr}.

Users may also want to profit from the Jacobian code of
\texttt{selfStart} models but supply explicit starting values other than
those suggested by \texttt{getInitial()}. This does not appear to be
possible with \texttt{nls()}. \texttt{nlsr::nlxb()} always requires
starting parameters. We could get these by separately calling
\texttt{getInitial()} to find them from the \texttt{selfStart} model,
but could use other values. The \texttt{selfStart} model is also used in
the formula for the \texttt{nlxb()} call.

We note that the analytic expressions for the Jacobian
(\texttt{gradient}) in the \texttt{SSLogis} function and others save
quantities by default in ``hidden'' variables, i.e., with names preceded
by ``.''. This behaviour is to avoid name conflicts with user
quantities. Hidden variables are not displayed by the \texttt{ls()}
command, so users need to remember to use
\texttt{ls(all.names\ =\ TRUE)} to see which quantities are defined, for
example, in developing their own selfStart model via copy and edit.

Interactive tools, such as ``visual fitting'' (John C. Nash and Velleman
(\protect\hyperlink{ref-nash1996nonlinear}{1996})) might be worth
considering as another way to find starting parameters, but we know of
no R capability of this type.

As a side note, the introduction of \texttt{scaleOffset} in R 4.1.1 to
deal with the convergence test for small residual problems now requires
that the \texttt{getInitial()} function have dot-arguments
(\texttt{...}) in its argument list. This illustrates the entanglement
of many features in \texttt{nls()} that complicate its maintenance and
improvement.

\hypertarget{bounds-constraints-on-parameters}{%
\subsection{Bounds constraints on
parameters}\label{bounds-constraints-on-parameters}}

For many problems, we know that parameters cannot be bigger or smaller
than some externally known limits. Such limits should be built into
models, but there are some important details for using the tools in R.

\begin{itemize}
\item
  \texttt{nls()} can only impose bounds if the
  \texttt{algorithm\ =\ "port"} argument is used in the call.
  Unfortunately, the documentation warns us:

  \emph{The algorithm = ``port'' code appears unfinished, and does not
  even check that the starting value is within the bounds. Use with
  caution, especially where bounds are supplied.}
\item
  \texttt{gsl\_nls()} does not offer bounds.
\item
  bounds are part of the default method for package \texttt{nlsr}.
\item
  \texttt{nlsLM()} includes bounds in the standard call, but we have
  observed cases where it fails to get the correct answer. From an
  examination of the code, we believe the authors have not taken into
  account all possibilities, though all programs have some weakness
  regarding constrained optimization in that programmers have to work
  with assumptions on the scale of numbers, and some problems will be
  outside the scope envisaged.
\end{itemize}

Examples of the use of bounds constraints for nonlinear regression
modeling are quite scarce. Traditionally, workers transform variables to
ensure bounds are satisfied, such as squaring a parameter, or using a
parameter which is the \(\log()\) of the actual parameter, so that its
exponential is positive. These transformations can create scaling
issues, but seem to generally work.

Bounds can also be helpful to avoid unanticipated excursions of the
minimization trajectory into inadmissible parameter regions. Such bounds
should not be too tight, as this may hinder progress to an optimum by
causing the methods to bump along the constraint in many cycles. Our aim
is to avoid rubbish, not specify where the solution will be, and we hope
the solution is not complicated by having a parameter on a bound.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Start MUST be feasible i.e. on or within bounds}
\NormalTok{anlshob1b }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}
             \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{3}\NormalTok{), }\AttributeTok{algorithm =} \StringTok{\textquotesingle{}port\textquotesingle{}}\NormalTok{)}
\FunctionTok{pnls}\NormalTok{(anlshob1b) }\CommentTok{\#  check the answer (short form)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlshob1b  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 10  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# nlsLM seems not to work with bounds in this example}
\NormalTok{anlsLM1b }\OtherTok{\textless{}{-}} \FunctionTok{nlsLM}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }\AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{3}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(anlsLM1b)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlsLM1b  -- ss= 881.02 :  c1 = 2  c2 = 6  c3 = 3; 2  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# also no warning if starting out of bounds, but gets a good answer!!}
\NormalTok{st4}\OtherTok{\textless{}{-}}\FunctionTok{c}\NormalTok{(}\AttributeTok{c1 =} \DecValTok{4}\NormalTok{, }\AttributeTok{c2 =} \DecValTok{4}\NormalTok{, }\AttributeTok{c3 =} \DecValTok{4}\NormalTok{)}
\NormalTok{anlsLMob }\OtherTok{\textless{}{-}} \FunctionTok{nlsLM}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ st4, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }\AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{3}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(anlsLMob)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlsLMob  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 4  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Try nlsr::nlxb()}
\NormalTok{anlx1b }\OtherTok{\textless{}{-}} \FunctionTok{nlxb}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts1, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }\AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{3}\NormalTok{))}
\FunctionTok{pshort}\NormalTok{(anlx1b)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlx1b  -- ss= 9.4726 :  c1 = 2  c2 = 4.4332  c3 = 3; 12 res/ 12 jac
\end{verbatim}

\hypertarget{philosophical-considerations}{%
\subsubsection{Philosophical
considerations}\label{philosophical-considerations}}

Bounds on parameters raise some interesting questions about how
uncertainty in parameter estimates should be computed or reported. That
is, the traditional ``standard errors'' are generally taken to imply
symmetric intervals about the point estimate in which the parameter may
be expected to be found with some probability under certain assumptions.
Bounds change those assumptions. Hence, \texttt{nlsr::nlxb()} does not
compute standard errors nor their derived statistics when bounds are
active. Transformations to implicitly enforce bounds avoid this issue,
but at the cost of changing the form of the model.

\hypertarget{fixed-parameters-masks}{%
\subsubsection{Fixed parameters (masks)}\label{fixed-parameters-masks}}

Let us try to fix (mask) the first parameter in the first two example
problems. We might want to do this if it is considered that the
asymptote is known from external sources, but we wish later to have it
available to be estimated.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Hobbsmaskx.R {-}{-} masks with formula specification of the problem}
\FunctionTok{require}\NormalTok{(nlsr); }\FunctionTok{require}\NormalTok{(minpack.lm); traceval }\OtherTok{\textless{}{-}} \ConstantTok{FALSE}
\NormalTok{stu }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{b1 =} \DecValTok{200}\NormalTok{, }\AttributeTok{b2 =} \DecValTok{50}\NormalTok{, }\AttributeTok{b3 =} \FloatTok{0.3}\NormalTok{) }\CommentTok{\# a default starting vector (named!)}
\NormalTok{sts }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{c1 =} \DecValTok{2}\NormalTok{, }\AttributeTok{c2 =} \DecValTok{5}\NormalTok{, }\AttributeTok{c3 =} \DecValTok{3}\NormalTok{) }\CommentTok{\# a default scaled starting vector (named!)}
\CommentTok{\# fix first parameter}
\NormalTok{anxbmsk1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(frmu, }\AttributeTok{start =}\NormalTok{ stu, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }
            \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{trace=}\NormalTok{traceval))}
\FunctionTok{print}\NormalTok{(anxbmsk1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## residual sumsquares =  2.6182  on  12 observations
##     after  4    Jacobian and  4 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## b1                   200U M         NA         NA         NA           0          NA  
## b2               49.5108          1.12      44.21  8.421e-13  -2.887e-07        1022  
## b3              0.311461      0.002278      136.8  1.073e-17   0.0001635      0.4569
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{anlM1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlsLM}\NormalTok{(frmu, }\AttributeTok{start =}\NormalTok{ stu, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }
            \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{3}\NormalTok{), }\AttributeTok{trace=}\NormalTok{traceval))}
\FunctionTok{pnls}\NormalTok{(anlM1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlM1  -- ss= 2.6182 :  b1 = 200  b2 = 49.511  b3 = 0.31146; 4  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{anlsmsk1 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(frmu, }\AttributeTok{start =}\NormalTok{ stu, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }
        \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{200}\NormalTok{, }\DecValTok{60}\NormalTok{, }\DecValTok{3}\NormalTok{),  }\AttributeTok{algorithm=}\StringTok{"port"}\NormalTok{, }\AttributeTok{trace =}\NormalTok{ traceval))}
\FunctionTok{pnls}\NormalTok{(anlsmsk1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlsmsk1  -- ss= 2.6182 :  b1 = 200  b2 = 49.511  b3 = 0.31146; 5  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Hobbs scaled problem with bounds, formula specification}
\NormalTok{anlxmsks1 }\OtherTok{\textless{}{-}} \FunctionTok{nlxb}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}
                  \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{30}\NormalTok{))}
\FunctionTok{print}\NormalTok{(anlxmsks1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## residual sumsquares =  2.6182  on  12 observations
##     after  4    Jacobian and  4 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## c1                     2U M         NA         NA         NA           0          NA  
## c2               4.95108         0.112      44.21  8.421e-13  -2.981e-06       104.2  
## c3               3.11461       0.02278      136.8  1.073e-17   1.583e-05       4.482
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{anlshmsk1 }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts, }\AttributeTok{trace =}\NormalTok{ traceval, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}
             \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{30}\NormalTok{), }\AttributeTok{algorithm =} \StringTok{\textquotesingle{}port\textquotesingle{}}\NormalTok{)}
\FunctionTok{pnls}\NormalTok{(anlshmsk1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlshmsk1  -- ss= 2.6182 :  c1 = 2  c2 = 4.9511  c3 = 3.1146; 5  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{anlsLMmsks1 }\OtherTok{\textless{}{-}} \FunctionTok{nlsLM}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{),}
                 \AttributeTok{upper =} \FunctionTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{,}\DecValTok{6}\NormalTok{,}\DecValTok{30}\NormalTok{))}
\FunctionTok{pnls}\NormalTok{(anlsLMmsks1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## anlsLMmsks1  -- ss= 2.6182 :  c1 = 2  c2 = 4.9511  c3 = 3.1146; 4  itns
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Test with all parameters masked}
\NormalTok{anlxmskall}\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nlxb}\NormalTok{(frms, }\AttributeTok{start =}\NormalTok{ sts, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{lower =}\NormalTok{ sts, }\AttributeTok{upper =}\NormalTok{ sts))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in nlfb(start = pnum, resfn = trjfn, jacfn = trjfn, trace = trace, :
## All parameters are masked
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{print}\NormalTok{(anlxmskall)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in summary.nlsr(x): No unmasked parameters
\end{verbatim}

\begin{verbatim}
## residual sumsquares =  158.23  on  12 observations
##     after  0    Jacobian and  1 function evaluations
##   name            coeff          SE       tstat      pval      gradient    JSingval   
## c1                     2U M         NA         NA         NA          NA          NA  
## c2                     5U M         NA         NA         NA          NA          NA  
## c3                     3U M         NA         NA         NA          NA          NA
\end{verbatim}

\texttt{nlsr} has an output format that indicates the constraint status
of the parameter estimates. For \texttt{nlsr}, we have \textbf{chosen}
to suppress the calculation of approximate standard errors in the
parameters when constraints are active because their meaning under
constraints is unclear, though we believe this policy worthy of
discussion and further investigation.

\hypertarget{stabilization-of-gauss-newton-computations}{%
\subsection{Stabilization of Gauss-Newton
computations}\label{stabilization-of-gauss-newton-computations}}

All four major tools illustrated solve some variant of the Gauss-Newton
equations. \texttt{nls()} uses a modification of an approach suggested
by Hartley (\protect\hyperlink{ref-Hartley1961}{1961}), while
\texttt{nlsr}, \texttt{gslnls}, and \texttt{minpack.lm} use flavours of
Marquardt (\protect\hyperlink{ref-Marquardt1963}{1963}). \texttt{gslnls}
offers an accelerated Marquardt method and three alternative methods
which we have not investigated beyond one example. Control settings for
\texttt{nlxb()} or \texttt{nlfb()} allow exploration of Hartley and
Marquardt algorithm variants, principally for the benefit of program
developers. In general, the Levenberg-Marquardt stabilization is
important in obtaining solutions in methods of the Gauss-Newton family,
as \texttt{nls()} terminates too frequently and unnecessarily with
\texttt{singular\ gradient} errors. A tidy incorporation of Marquardt
stabilization into \texttt{nls()} would be a major improvement.

\hypertarget{programming-language}{%
\subsection{Programming language}\label{programming-language}}

An important choice made in developing \texttt{nlsr} was to code
entirely within the R programming language. \texttt{nls()} uses a mix of
R, C, and Fortran, as does \texttt{minpack.lm}. \texttt{gslnls} is an R
wrapper to various C-language routines in the GNU Scientific Library
(Galassi et al. (\protect\hyperlink{ref-GSL-manual}{2009})). Generally,
keeping to a single programming language can allow for easier
maintenance and upgrades. It also avoids some work when there are
changes or upgrades to libraries for the non-R languages. R is usually
considered slower than most computing environments because it keeps
track of objects and because it is usually interpreted. In recent years,
the performance penalty for using code entirely in R has been much
reduced with the just-in-time compiler and other improvements. All-R
computation may now offer acceptable performance. In \texttt{nlsr}, the
use of R may be a smaller performance cost than the aggressive approach
to a solution, such as large computational limits for residuals and
Jacobians.

There are, of course, problems where use of R code will be slower. For
example, a ``large'' case of the Penalty 1 test problem of Moré, Garbow,
and Hillstrom (\protect\hyperlink{ref-More1981TUO}{1981}) will take
noticeable time to solve. Problems in large numbers of parameters create
large computational matrices for the Jacobian or its inner product,
which Gauss-Newton methods create and process. This is confirmed by
comparing \texttt{nlsr::nlfb()} and \texttt{minpack.lm::nls.lm()} on the
Penalty 1 problem of various sizes, where the latter, which uses
compiled Fortran internally, was moderately faster. However, the method
``Rcgmin'' from package \texttt{optimx} (John C. Nash and Varadhan
(\protect\hyperlink{ref-p-optimx}{2011})) proved to be much faster than
either, despite being entirely written in R, and is a much more reliable
conjugate gradient minimizer than method ``CG'' of \texttt{optim()}. One
of us created both, but now deprecates the latter. Timings of CG methods
are more sensitive to small differences in setup and arithmetic than
Gauss-Newton methods, and the pattern of such timings varies much less
regularly with problem size for CG methods.

\hypertarget{data-sources-for-problems}{%
\subsection{Data sources for problems}\label{data-sources-for-problems}}

\texttt{nls()} can be called without specifying the \texttt{data}
argument. In this case, it will search in the available environments
(i.e., workspaces) for suitable data objects. We do not like this
approach, but it is ``the R way''. R allows users to leave many objects
in the default (\texttt{.GlobalEnv}) workspace. Moreover, users have to
actively suppress saving this workspace (\texttt{.RData}) on exit;
otherwise, any such file in the path will be loaded on startup. R users
in our acquaintance avoid saving the workspace because of lurking data
and functions that may cause unwanted results.

\hypertarget{feature-subsetting}{%
\subsubsection{Feature: Subsetting}\label{feature-subsetting}}

\texttt{nls()} and other class \texttt{nls} tools accept an argument
\texttt{subset}. This acts through the mediation of
\texttt{model.frame}, which is not obvious in the source code files
\texttt{/src/library/stats/R/nls.R} and
\texttt{/src/library/stats/src/nls.C}. Having \texttt{subset} at the
level of the call to a function like \texttt{nls()} saves effort, but it
does mean that the programmer of the solver needs to be aware of the
origin (and value) of objects such as the data, residuals and Jacobian.
By preference, we would implement subsetting by zero-value weights, with
observation counts (and degrees of freedom) computed via the numbers of
non-zero weights. Alternatively, we would extract a working dataframe
from the relevant elements in the original.

\hypertarget{feature-na.action-missing-value-treatment}{%
\subsubsection{Feature: na.action (missing value
treatment)}\label{feature-na.action-missing-value-treatment}}

\texttt{na.action} is an argument to the \texttt{nls()} function, but it
does not appear obviously in the source code, often being handled behind
the scenes after referencing the option \texttt{na.action}. This feature
also changes the data supplied to our nonlinear least squares solver. A
useful, but possibly dated, description is given in:
\emph{\url{https://stats.idre.ucla.edu/r/faq/how-does-r-handle-missing-values/}}.
The typical default action, which can be seen by using the command
\texttt{getOption("na.action")} is \texttt{na.omit}. This option removes
from computations any observations containing missing values (i.e.~any
row of a data frame containing an NA). \texttt{na.exclude} does much of
the same for solver computations, but keeps the rows with NA elements so
that predictions are in the correct row position. We recommend that
workers actually test output to verify the behaviour is as wanted. See
\emph{\url{https://stats.stackexchange.com/questions/492955/should-i-use-na-omit-or-na-exclude-in-a-linear-model-in-r}}.
As with \texttt{subset}, our concern with \texttt{na.action} is that
users may be unaware of the effects of an option they may not even know
has been set. Should \texttt{na.fail} be the default?

\hypertarget{model-frame}{%
\subsection{Model frame}\label{model-frame}}

\texttt{model} is an argument to the \texttt{nls()} and related
functions, which is documented as:

\emph{\textbf{model} logical. If true, the model frame is returned as
part of the object. Default is FALSE.}

Indeed, the argument only gets used when \texttt{nls()} is about to
return its result object, and the element \texttt{model} is NULL unless
the calling argument \texttt{model} is TRUE. (Using the same name could
be confusing.) Despite this, the model frame is used within the function
code in the form of the object \texttt{mf}. We feel that users could
benefit from more extensive documentation and examples of its use since
it is used to implement features like \texttt{subset}. Moreover, it
provides access to resources that may be important in further
calculations on models.

\hypertarget{weights-on-observations}{%
\subsection{Weights on observations}\label{weights-on-observations}}

All four main tools we consider here allow a \texttt{weights} argument
that specifies a vector of fixed weights the same length as the number
of residuals. Each residual is multiplied by the square root of the
corresponding weight. Where available, the values returned by the
\texttt{residuals()} function are weighted, and the \texttt{fitted()} or
\texttt{predict()} functions are used to compute raw residuals.

While fixed weights may be useful, there are many problems for which we
want weights that are determined at least partially from the model
parameters; for example, a measure of the standard deviation of
observations. Such dynamic weighting situations are discussed in the
vignette ``Introduction to nlsr'' of package \texttt{nlsr} in section
\emph{Weights that are functions of the model parameters}.
\texttt{minpack.lm} offers a function \texttt{wfct()} to facilitate such
weighting. Care is advised in applying such ideas.

\hypertarget{weights-in-returned-functions-from-nls}{%
\subsubsection{Weights in returned functions from
nls()}\label{weights-in-returned-functions-from-nls}}

The function \texttt{resid()} (an alias for \texttt{residuals()}) gives
weighted residuals, as does, for example, \texttt{result\$m\$resid()}.
The function \texttt{nlsModel()}, which we have had to extract from the
base R code and explicitly \texttt{source()} because it is not exported
to the working namespace, allows us to compute residuals for particular
coefficient sets.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{wts }\OtherTok{\textless{}{-}} \FloatTok{0.5} \SpecialCharTok{\^{}}\NormalTok{ tt }\CommentTok{\# simple weights}
\NormalTok{frmlogis }\OtherTok{\textless{}{-}}\NormalTok{ weed }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym}\SpecialCharTok{/}\NormalTok{(}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}}\NormalTok{ tt) }\SpecialCharTok{/}\NormalTok{ scal))}
\NormalTok{Asym }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{; xmid }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{; scal }\OtherTok{\textless{}{-}} \DecValTok{1}
\NormalTok{nowt }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(weed }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(tt, Asym, xmid, scal)) }\CommentTok{\# unweighted}
\NormalTok{rnowt }\OtherTok{\textless{}{-}}\NormalTok{ nowt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# This has unweighted residual and Jacobian. Does not take coefficients.}
\FunctionTok{attr}\NormalTok{(rnowt, }\StringTok{"gradient"}\NormalTok{) }\OtherTok{\textless{}{-}} \ConstantTok{NULL}\NormalTok{; rnowt}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] -0.011900  0.032755 -0.092030 -0.208782 -0.392634  0.057594  1.105728
##  [8] -0.715786  0.107647  0.348396 -0.652592  0.287569
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{usewt }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(weed }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(tt, Asym, xmid, scal), }\AttributeTok{weights=}\NormalTok{wts)}
\NormalTok{rusewt }\OtherTok{\textless{}{-}}\NormalTok{ usewt}\SpecialCharTok{$}\NormalTok{m}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# weighted. Does not take coefficients.}
\FunctionTok{attr}\NormalTok{(rusewt,}\StringTok{"gradient"}\NormalTok{) }\OtherTok{\textless{}{-}} \ConstantTok{NULL}\NormalTok{; rusewt}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
##  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{source}\NormalTok{(}\StringTok{"nlsModel.R"}\NormalTok{)}
\NormalTok{nmod0 }\OtherTok{\textless{}{-}} \FunctionTok{nlsModel}\NormalTok{(frmlogis, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{start =} \FunctionTok{c}\NormalTok{(}\AttributeTok{Asym =} \DecValTok{1}\NormalTok{, }\AttributeTok{xmid =} \DecValTok{1}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{), }\AttributeTok{wts =}\NormalTok{ wts)}
\NormalTok{rn0}\OtherTok{\textless{}{-}}\NormalTok{nmod0}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{() }\CommentTok{\# Parameters are supplied in nlsModel() \textasciigrave{}start\textasciigrave{} above.}
\FunctionTok{attr}\NormalTok{(rn0,}\StringTok{"gradient"}\NormalTok{) }\OtherTok{\textless{}{-}} \ConstantTok{NULL}\NormalTok{; rn0}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] 3.3998 3.2545 3.0961 2.9784 2.8438 2.7748 2.6910 2.3474 2.1724 1.9359
## [11] 1.6572 1.4214
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nmod }\OtherTok{\textless{}{-}} \FunctionTok{nlsModel}\NormalTok{(frmlogis, }\AttributeTok{data =}\NormalTok{ weeddf, }\AttributeTok{start =} \FunctionTok{coef}\NormalTok{(usewt), }\AttributeTok{wts =}\NormalTok{ wts)}
\NormalTok{rn }\OtherTok{\textless{}{-}}\NormalTok{ nmod}\SpecialCharTok{$}\FunctionTok{resid}\NormalTok{()}
\FunctionTok{attr}\NormalTok{(rn,}\StringTok{"gradient"}\NormalTok{) }\OtherTok{\textless{}{-}} \ConstantTok{NULL}\NormalTok{; rn}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1]  0.0085640  0.0324442 -0.0176652 -0.0388479 -0.0579575  0.0163623
##  [7]  0.1042380 -0.0411766  0.0052509  0.0084324 -0.0194246 -0.0024053
\end{verbatim}

\hypertarget{minor-issues-with-nonlinear-least-squares-tools}{%
\subsection{Minor issues with nonlinear least-squares
tools}\label{minor-issues-with-nonlinear-least-squares-tools}}

\hypertarget{interim-output-from-the-port-algorithm}{%
\subsubsection{Interim output from the ``port''
algorithm}\label{interim-output-from-the-port-algorithm}}

As the \texttt{nls()} \textbf{man} page states, when the ``port''
algorithm is used with the \texttt{trace} argument TRUE, the iterations
display the objective function value which is 1/2 the sum of squares (or
deviance). The trace display is likely embedded in the Fortran of the
\texttt{nlminb} routine that is called to execute the ``port''
algorithm, but the factor of 2 discrepancy is nonetheless unfortunate
for users.

\hypertarget{failure-to-return-the-best-result-achieved}{%
\subsubsection{Failure to return the best result
achieved}\label{failure-to-return-the-best-result-achieved}}

If \texttt{nls()} reaches a point where it cannot continue but has not
found a point where the relative offset convergence criterion is met, it
may simply exit, especially if a ``singular gradient'' is found.
However, this may occur \textbf{after} considerable progress has been
made in reducing the sum of squared residuals. Here is an abbreviated
example:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{time }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,  }\DecValTok{2}\NormalTok{,  }\DecValTok{3}\NormalTok{,  }\DecValTok{4}\NormalTok{,  }\DecValTok{6}\NormalTok{ , }\DecValTok{8}\NormalTok{, }\DecValTok{10}\NormalTok{, }\DecValTok{12}\NormalTok{, }\DecValTok{16}\NormalTok{)}
\NormalTok{conc }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FloatTok{0.7}\NormalTok{, }\FloatTok{1.2}\NormalTok{, }\FloatTok{1.4}\NormalTok{, }\FloatTok{1.4}\NormalTok{, }\FloatTok{1.1}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.6}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.3}\NormalTok{)}
\NormalTok{NLSdata }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(time,conc)}
\NormalTok{NLSstart }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{lrc1 =} \SpecialCharTok{{-}}\DecValTok{2}\NormalTok{, }\AttributeTok{lrc2 =} \FloatTok{0.25}\NormalTok{, }\AttributeTok{A1 =} \DecValTok{150}\NormalTok{, }\AttributeTok{A2 =} \DecValTok{50}\NormalTok{) }\CommentTok{\# a starting vector (named!)}
\NormalTok{NLSformula }\OtherTok{\textless{}{-}}\NormalTok{ conc }\SpecialCharTok{\textasciitilde{}}\NormalTok{ A1 }\SpecialCharTok{*} \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{exp}\NormalTok{(lrc1) }\SpecialCharTok{*}\NormalTok{ time) }\SpecialCharTok{+}\NormalTok{ A2 }\SpecialCharTok{*} \FunctionTok{exp}\NormalTok{(}\SpecialCharTok{{-}}\FunctionTok{exp}\NormalTok{(lrc2) }\SpecialCharTok{*}\NormalTok{ time)}
\NormalTok{tryit }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(NLSformula, }\AttributeTok{data =}\NormalTok{ NLSdata, }\AttributeTok{start =}\NormalTok{ NLSstart, }\AttributeTok{trace =} \ConstantTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 61216.    (3.56e+03): par = (-2 0.25 150 50)
## 2.1757    (2.23e+01): par = (-1.9991 0.31711 2.6182 -1.3668)
## 1.6211    (7.14e+00): par = (-1.9605 -2.6203 2.5753 -0.55599)
## Error in nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE) : 
##   singular gradient
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# print(tryit) \# gives no information on best sumsquares or model parameters}
\end{Highlighting}
\end{Shaded}

Note that the sum of squares has been reduced from 61216 to 1.6211, but
unless \texttt{trace} is invoked, the user will not get any information
about this. This almost trivial change to the \texttt{nls()} function
could be useful to R users.

\hypertarget{estimating-models-that-are-partially-linear}{%
\subsection{Estimating models that are partially
linear}\label{estimating-models-that-are-partially-linear}}

The variable projection method (Golub and Pereyra
(\protect\hyperlink{ref-Golub1973}{1973}), O'Leary and Rust
(\protect\hyperlink{ref-OlearyRust13}{2013})) is usually much more
effective than general approaches in finding good solutions to nonlinear
least squares problems when some of the parameters appear linearly. In
our logistic examples, the asymptote parameters are an illustration.
However, identifying which parameters are linear and communicating this
information to estimating functions is not a trivial task.
\texttt{nls()} has an option \texttt{algorithm\ =\ "plinear"} that
allows some partially linear models to be solved. The other tools, as
far as we are aware, do not offer any such capability. The
\texttt{nlstac} package uses a different algorithm for similar goals.

Within \texttt{nls()} itself we must, unfortunately, use different
specifications with different \texttt{algorithm} options. For example,
the explicit linear model \texttt{y\ \textasciitilde{}\ a\ *\ x\ +\ b}
does not work. As with the linear modeling function \texttt{lm()}, this
must be specified as \texttt{y\ \textasciitilde{}\ x}. Within
\texttt{nls()}, consider the following four different specifications for
the same problem, plus an intuitive choice, labeled \texttt{fm2a}, that
does not work. In this failed attempt, putting the \texttt{Asym}
parameter in the model causes the \texttt{plinear} algorithm to try to
add another term to the model. We believe this is unfortunate, and would
like to see a consistent syntax. We do not foresee a resolution for this
issue, which arises from different approaches to model specification. In
the example, we have not evaluated the commands to save space.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DNase1 }\OtherTok{\textless{}{-}} \FunctionTok{subset}\NormalTok{(DNase, Run }\SpecialCharTok{==} \DecValTok{1}\NormalTok{) }\CommentTok{\# select the data}
\DocumentationTok{\#\# using a selfStart model {-} do not specify the starting parameters}
\NormalTok{fm1 }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(density }\SpecialCharTok{\textasciitilde{}} \FunctionTok{SSlogis}\NormalTok{(}\FunctionTok{log}\NormalTok{(conc), Asym, xmid, scal), DNase1)}
\FunctionTok{summary}\NormalTok{(fm1)}

\DocumentationTok{\#\# using conditional linearity {-} leave out the Asym parameter}
\NormalTok{fm2 }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(density }\SpecialCharTok{\textasciitilde{}} \DecValTok{1} \SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}} \FunctionTok{log}\NormalTok{(conc)) }\SpecialCharTok{/}\NormalTok{ scal)),}
                 \AttributeTok{data =}\NormalTok{ DNase1, }\AttributeTok{start =} \FunctionTok{list}\NormalTok{(}\AttributeTok{xmid =} \DecValTok{0}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{),}
                 \AttributeTok{algorithm =} \StringTok{"plinear"}\NormalTok{)}
\FunctionTok{summary}\NormalTok{(fm2)}

\DocumentationTok{\#\# without conditional linearity}
\NormalTok{fm3 }\OtherTok{\textless{}{-}} \FunctionTok{nls}\NormalTok{(density }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}} \FunctionTok{log}\NormalTok{(conc)) }\SpecialCharTok{/}\NormalTok{ scal)),}
                 \AttributeTok{data =}\NormalTok{ DNase1,}
                 \AttributeTok{start =} \FunctionTok{list}\NormalTok{(}\AttributeTok{Asym =} \DecValTok{3}\NormalTok{, }\AttributeTok{xmid =} \DecValTok{0}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{))}
\FunctionTok{summary}\NormalTok{(fm3)}

\DocumentationTok{\#\# using Port\textquotesingle{}s nl2sol algorithm}
\NormalTok{fm4 }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(density }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}} \FunctionTok{log}\NormalTok{(conc)) }\SpecialCharTok{/}\NormalTok{ scal)),}
                 \AttributeTok{data =}\NormalTok{ DNase1, }\AttributeTok{start =} \FunctionTok{list}\NormalTok{(}\AttributeTok{Asym =} \DecValTok{3}\NormalTok{, }\AttributeTok{xmid =} \DecValTok{0}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{),}
                 \AttributeTok{algorithm =} \StringTok{"port"}\NormalTok{))}
\FunctionTok{summary}\NormalTok{(fm4)}

\DocumentationTok{\#\# using conditional linearity AND Asym does not work}
\NormalTok{fm2a }\OtherTok{\textless{}{-}} \FunctionTok{try}\NormalTok{(}\FunctionTok{nls}\NormalTok{(density }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Asym }\SpecialCharTok{/}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{exp}\NormalTok{((xmid }\SpecialCharTok{{-}} \FunctionTok{log}\NormalTok{(conc)) }\SpecialCharTok{/}\NormalTok{ scal)), }
                 \AttributeTok{data =}\NormalTok{ DNase1, }\AttributeTok{start =} \FunctionTok{list}\NormalTok{(}\AttributeTok{Asym=}\DecValTok{3}\NormalTok{, }\AttributeTok{xmid =} \DecValTok{0}\NormalTok{, }\AttributeTok{scal =} \DecValTok{1}\NormalTok{),}
                 \AttributeTok{algorithm =} \StringTok{"plinear"}\NormalTok{, }\AttributeTok{trace =} \ConstantTok{TRUE}\NormalTok{))}
\FunctionTok{summary}\NormalTok{(fm2a)}
\end{Highlighting}
\end{Shaded}

\hypertarget{models-with-indexed-parameters}{%
\subsection{Models with indexed
parameters}\label{models-with-indexed-parameters}}

Some models have several common parameters and others that are tied to
particular cases. The \textbf{man} file for \texttt{nls()} includes an
example of a situation in which parameters are indexed but which uses
the ``plinear'' option as an added complication. Running this example
reveals that the answers for the parameters are not indexed as in a
vector. That is, we do not see \texttt{a{[}1{]},\ a{[}2{]},\ a{[}3{]}}
but \texttt{a1,\ a2,\ a3}. This is no doubt because programming for
indexed parameters is challenging. We note that there are capabilities
in packages for mixed-effects modeling such as \texttt{nlme} (Pinheiro
et al. (\protect\hyperlink{ref-p-nlme}{2013})), \texttt{bbmle} (Bolker
and Team (\protect\hyperlink{ref-p-bbmle}{2013})), and \texttt{lme4} (D.
Bates et al. (\protect\hyperlink{ref-p-lme4}{2015})) for estimating
models of such type.

\hypertarget{some-other-cran-packages-for-nonlinear-modeling}{%
\subsection{Some other CRAN packages for nonlinear
modeling}\label{some-other-cran-packages-for-nonlinear-modeling}}

\texttt{onls} (Andrej-Nikolai Spiess
(\protect\hyperlink{ref-R-onls-manual}{2022})) -- The usual optimization
in estimating nonlinear models is the vertical difference between the
dependent variable and the functional model. \texttt{onls} minimizes the
sum of \textbf{orthogonal} residuals. The objective is therefore
different and involves nontrivial extra calculation. A vignette with the
package and the blog article
\url{https://www.r-bloggers.com/2015/01/introducing-orthogonal-nonlinear-least-squares-regression-in-r/}
give some description with illustrative graphs. \texttt{onls} appears to
be limited to problems with one independent and one dependent variable.
The Wikipedia article
\url{https://en.wikipedia.org/wiki/Total_least_squares} presents an
overview of some ideas, with references to the literature. The approach
needs a wider discussion and tutorial examples to allow its merits to be
judged than can be included here.

\texttt{crsnls} (Tvrdík (\protect\hyperlink{ref-R-crsnls-manual}{2016}))
-- This package allows nonlinear estimation by controlled random search
via two methods. There is unfortunately no vignette. A modest trial we
carried out showed \texttt{nlsr::nlxb()} gave the same results in a
small fraction of the time required by either of the methods in
\texttt{crsnls}. The method discussed in Josef Tvrdík and Ivan Křivý and
Ladislav Mišík (\protect\hyperlink{ref-tvrdik2007}{2007}) claims better
reliability in finding solutions than a Levenberg-Marquardt code
(actually from Matlab), but the tests were conducted on the extreme NIST
examples mentioned next.

\texttt{NISTnls} (National Institutes for Standards and R port by
Douglas Bates (\protect\hyperlink{ref-NISTnls-manual}{2012})) -- This
package provides R code and data for a set of (numerically
ill-conditioned) nonlinear least squares problems from the U.S. National
Institute for Standards and Technology. These may not represent
real-world situations.

\texttt{nlshelper} (Duursma
(\protect\hyperlink{ref-nlshelper-manual}{2017})) -- This package, which
unfortunately lacks a vignette, provides a few utilities for
summarizing, testing, and plotting non-linear regression models
estimated with nls(), nlsList() or nlme() that are linked or grouped in
some way.

\texttt{nlsic} (Sokol (\protect\hyperlink{ref-nlsic-manual}{2022})) --
This solves nonlinear least squares problems with optional equality
and/or inequality constraints. It is clearly \textbf{not} about
modeling, and the input and output are quite different from class
\texttt{nls} methods. However, there do not appear to be other R
packages with these capabilities.

\texttt{nlsMicrobio} (Baty and Delignette-Muller
(\protect\hyperlink{ref-nlsMicrobio-manual}{2014})) -- Data sets and
nonlinear regression models dedicated to predictive microbiology,
including a vignette, by authors of the \texttt{nlstools} package.

\texttt{nlstools} (Baty and Delignette-Muller
(\protect\hyperlink{ref-p-nlstools}{2013})) -- This package provides
several tools for aiding the estimation of nonlinear models,
particularly using \texttt{nls()}. The vignette is actually a journal
article, and the authors have considerable experience in the subject.

\texttt{nlsmsn} (Prates, Lachos, and Garay
(\protect\hyperlink{ref-m-nlsmsn}{2021})) -- Fit univariate non-linear
scale mixture of skew-normal(NL-SMSN) regression, with details in Garay,
Lachos, and Abanto-Valle (2011). The problem here is to minimize an
objective that is modified from the traditional sum of squared
residuals.

\texttt{nls.multstart} (Padfield and Matheson
(\protect\hyperlink{ref-nlsmultstart}{2020})) -- Non-linear least
squares regression using AIC scores with the Levenberg-Marquardt
algorithm using multiple starting values for increasing the chance that
the minimum found is the global minimum.

\texttt{nls2} (Grothendieck (\protect\hyperlink{ref-nls2-22}{2022})) --
Nonlinear least squares by brute force has similar motivations to
\texttt{nls.multstart}, but uses \texttt{nls()} within multiple trials.
The author has extensive expertise in R.

\texttt{nlstac} (Rodriguez-Arias et al.
(\protect\hyperlink{ref-m-nlstac}{2020})) -- A set of functions
implementing the algorithm described in Fernandez Torvisco et al.~(2018)
for fitting separable nonlinear regression curves. The special class of
problem for which this package is intended is an important and difficult
one. No vignette is provided, unfortunately.

\texttt{easynls} -- Fit and plot some nonlinear models. Thirteen models
are treated, but there is minimal documentation and no vignette. Package
\texttt{nlraa} is to be preferred.

\texttt{nlraa} (Miguez (\protect\hyperlink{ref-MiguezNLRAA2021}{2021}))
-- a set of nonlinear \emph{selfStart} models, primarily from
agriculture. Most include analytic Jacobian code.

\texttt{optimx} (John C. Nash and Varadhan
(\protect\hyperlink{ref-p-optimx}{2011})) -- This provides optimizers
that can be applied to minimize a nonlinear function which could be a
nonlinear sum of squares. This is not generally recommended if nonlinear
least squares programs can be easily used, but provides a check and
alternative solvers.

\hypertarget{tests-and-use-case-examples}{%
\subsection{Tests and use-case
examples}\label{tests-and-use-case-examples}}

Maintainers of packages need suitable tests and use-case examples in
order

\begin{itemize}
\tightlist
\item
  to ensure packages work properly, in particular, giving results
  comparable to or better than the functions they are to replace.
\item
  to test individual solver functions to ensure they work across the
  range of calling mechanisms, that is, different ways of supplying
  inputs to the solver(s);
\item
  to pose ``silly'' inputs to see if these bad inputs are caught by the
  programs.
\end{itemize}

Such goals align with the aims of \textbf{unit testing} (Wickham
(\protect\hyperlink{ref-HWtestthat11}{2011}), Wickham et al.
(\protect\hyperlink{ref-HWdevtools21}{2021}), the conventional R package
testing tools, and a number of blog posts found by searching ``unit
testing in R''). In our work, one of us has developed a working
prototype package at
\emph{\url{https://github.com/ArkaB-DS/nlsCompare}}. A primary design
objective of this is to allow the summarisation of multiple tests in a
compact output. The prototype has a vignette to illustrate its use.

\hypertarget{future-of-nonlinear-model-estimation-in-r}{%
\subsection{Future of nonlinear model estimation in
R}\label{future-of-nonlinear-model-estimation-in-r}}

Given its importance to R, it is possible that \texttt{nls()} will
remain more or less as it has been for the past several decades. If so,
the focus of discussion should be the measures needed to secure its
continued operation for legacy purposes and how that may be
accomplished. We welcome an opportunity to participate in such
conversations.

To advance the stability and maintainability of R, we believe the
program objects (R functions) that are created by tools such as
\texttt{nls()} should have minimal cross-linkages and side-effects. The
aspects of \texttt{nls()} that concern us follow.

\begin{itemize}
\item
  R tools presume data and parameters needed are available in an
  accessible environment. This provides a compact syntax to invoke the
  calculations, but the wrong data can be used if the internal search
  finds a valid name that is not the object we want, or if
  \texttt{subset} or \texttt{na.action} settings modify the selection,
  or \texttt{weights} are applied.
\item
  Mixing of R, C and Fortran code adds to the burden of following the
  program logic.
\item
  The class \texttt{nls} structure simplifies calls with its rich set of
  functions, but also adds to the task of understanding what has been
  done. A design that isolates the setup, solution, and post-solution
  parts of complicated calculations reduces the number of objects that
  must be kept in alignment.
\end{itemize}

Given the existence of examples of good practices such as analytic
derivatives, stabilized solution of Gauss-Newton equations, and
bounds-constrained parameters, base R tools should be moving to
incorporate them. These capabilities are available now by using several
tools, but it would be helpful if they were unified.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

Hans Werner Borchers was helpful in developing the GSoC project
motivating this article and in comments on this and related work.
Heather Turner co-mentored the project and helped guide the progress of
the work. Exchanges with Fernando Miguez helped to clarify aspects of
\texttt{selfStart} models and instigated the ``Introduction to nlsr''
vignette. Colin Gillespie (package \texttt{benchmarkme}) has been
helpful in guiding our attempts to succinctly summarize computing
environments. Particularly thoughtful reviews by the referees, including
details of inner workings of R, have led to important revisions.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-R-onls-manual}{}}%
Andrej-Nikolai Spiess. 2022. \emph{Onls: Orthogonal Nonlinear
Least-Squares Regression}.
\url{https://CRAN.R-project.org/package=onls}.

\leavevmode\vadjust pre{\hypertarget{ref-bateswatts}{}}%
Bates, D. M., and D. G. Watts. 1988. \emph{Nonlinear Regression Analysis
and Its Applications}. Wiley.

\leavevmode\vadjust pre{\hypertarget{ref-BatesWatts81}{}}%
Bates, Douglas M., and Watts, Donald G. 1981. {``A Relative Offset
Orthogonality Convergence Criterion for Nonlinear Least Squares.''}
\emph{Technometrics} 23 (2): 179--83.

\leavevmode\vadjust pre{\hypertarget{ref-p-lme4}{}}%
Bates, Douglas, Martin Mächler, Ben Bolker, and Steve Walker. 2015.
{``Fitting Linear Mixed-Effects Models Using {lme4}.''} \emph{Journal of
Statistical Software} 67 (1): 1--48.
\url{https://doi.org/10.18637/jss.v067.i01}.

\leavevmode\vadjust pre{\hypertarget{ref-p-nlstools}{}}%
Baty, Florent, and Marie-Laure Delignette-Muller. 2013. \emph{Nlstools:
Tools for Nonlinear Regression Diagnostics}.

\leavevmode\vadjust pre{\hypertarget{ref-nlsMicrobio-manual}{}}%
---------. 2014. \emph{nlsMicrobio: Data Sets and Nonlinear Regression
Models Dedicated to Predictive Microbiology}.

\leavevmode\vadjust pre{\hypertarget{ref-p-bbmle}{}}%
Bolker, Ben, and R Development Core Team. 2013. \emph{Bbmle: Tools for
General Maximum Likelihood Estimation}.
\url{http://CRAN.R-project.org/package=bbmle}.

\leavevmode\vadjust pre{\hypertarget{ref-gslnls23}{}}%
Chau, Joris. 2023. \emph{Gslnls: GSL Nonlinear Least-Squares Fitting}.
\url{https://CRAN.R-project.org/package=gslnls}.

\leavevmode\vadjust pre{\hypertarget{ref-nlshelper-manual}{}}%
Duursma, Remko. 2017. \emph{Nlshelper: Convenient Functions for
Non-Linear Regression}.
\url{https://CRAN.R-project.org/package=nlshelper}.

\leavevmode\vadjust pre{\hypertarget{ref-minpacklm12}{}}%
Elzhov, Timur V., Katharine M. Mullen, Andrej-Nikolai Spiess, and Ben
Bolker. 2012. \emph{Minpack.lm: R Interface to the Levenberg-Marquardt
Nonlinear Least-Squares Algorithm Found in MINPACK, Plus Support for
Bounds}. R Project for Statistical Computing.
\url{http://CRAN.R-project.org/package=minpack.lm}.

\leavevmode\vadjust pre{\hypertarget{ref-FoxHallSchryer1978}{}}%
Fox, P. A., A. P. Hall, and N. L. Schryer. 1978. {``The PORT
Mathematical Subroutine Library.''} \emph{ACM Trans. Math. Softw.} 4
(2): 104--26. \url{https://doi.org/10.1145/355780.355783}.

\leavevmode\vadjust pre{\hypertarget{ref-GSL-manual}{}}%
Galassi, Mark, Jim Davies, James Theiler, Brian Gough, and Gerard
Jungman. 2009. \emph{GNU Scientific Library - Reference Manual, Third
Edition, for GSL Version 1.12 (3. Ed.).}

\leavevmode\vadjust pre{\hypertarget{ref-Bioconductor}{}}%
Gentleman, Robert C., Vincent J. Carey, Douglas M. Bates, Ben Bolstad,
Marcel Dettling, Sandrine Dudoit, Byron Ellis, et al. 2004.
{``Bioconductor: Open Software Development for Computational Biology and
Bioinformatics.''} \emph{{Genome Biology}} 5 (R80).
\url{https://doi.org/10.1186/gb-2004-5-10-r80}.

\leavevmode\vadjust pre{\hypertarget{ref-Golub1973}{}}%
Golub, G. H., and V. Pereyra. 1973. {``The Differentiation of
Pseudo-Inverses and Nonlinear Least Squares Problems Whose Variables
Separate.''} \emph{SIAM Journal of Numerical Analysis} 10 (2): 413--32.

\leavevmode\vadjust pre{\hypertarget{ref-nls2-22}{}}%
Grothendieck, G. 2022. \emph{Nls2: Non-Linear Regression with Brute
Force}. \url{https://CRAN.R-project.org/package=nls2}.

\leavevmode\vadjust pre{\hypertarget{ref-Hartley1961}{}}%
Hartley, H. O. 1961. {``The Modified Gauss-Newton Method for the Fitting
of Non-Linear Regression Functions by Least Squares.''}
\emph{Technometrics} 3: 269-\/-280.

\leavevmode\vadjust pre{\hypertarget{ref-Huet2004}{}}%
Huet, S., A. Bouvier, M.-A. Poursat, and E. Jolivet. 2004.
\emph{Statistical Tools for Nonlinear Regression: A Practical Guide with
{S-PLUS} Examples, 2nd Edition}. Berlin \& New York: Springer-Verlag.

\leavevmode\vadjust pre{\hypertarget{ref-nlsr2023manual}{}}%
John C Nash, and Duncan Murdoch. 2023. \emph{{nlsr: Functions for
Nonlinear Least Squares Solutions}}.

\leavevmode\vadjust pre{\hypertarget{ref-tvrdik2007}{}}%
Josef Tvrdík and Ivan Křivý and Ladislav Mišík. 2007. {``Adaptive
Population-Based Search: Application to Estimation of Nonlinear
Regression Parameters.''} \emph{{Computational Statistics \& Data
Analysis}} 52 (2): 713--24.
\url{https://doi.org/10.1016/j.csda.2006.10.014}.

\leavevmode\vadjust pre{\hypertarget{ref-Levenberg1944}{}}%
Levenberg, Kenneth. 1944. {``A Method for the Solution of Certain
Non-Linear Problems in Least Squares.''} \emph{Quarterly of Applied
Mathematics} 2: 164-\/-168.

\leavevmode\vadjust pre{\hypertarget{ref-Marquardt1963}{}}%
Marquardt, Donald W. 1963. {``{An Algorithm for Least-Squares Estimation
of Nonlinear Parameters}.''} \emph{SIAM Journal on Applied Mathematics}
11 (2): 431--41.

\leavevmode\vadjust pre{\hypertarget{ref-MiguezNLRAA2021}{}}%
Miguez, Fernando. 2021. \emph{{nlraa: Nonlinear Regression for
Agricultural Applications}}.
\url{https://CRAN.R-project.org/package=nlraa}.

\leavevmode\vadjust pre{\hypertarget{ref-More1981TUO}{}}%
Moré, Jorge J., Burton S. Garbow, and Kenneth E. Hillstrom. 1981.
{``Testing Unconstrained Optimization Software.''} \emph{J-Toms} 7 (1):
17--41.

\leavevmode\vadjust pre{\hypertarget{ref-jn77ima}{}}%
Nash, John C. 1977. {``Minimizing a Nonlinear Sum of Squares Function on
a Small Computer.''} \emph{Journal of the Institute for Mathematics and
Its Applications} 19: 231--37.

\leavevmode\vadjust pre{\hypertarget{ref-cnm79}{}}%
---------. 1979. \emph{Compact Numerical Methods for Computers: Linear
Algebra and Function Minimisation}. Bristol: Adam Hilger.

\leavevmode\vadjust pre{\hypertarget{ref-p-optimx}{}}%
Nash, John C, and Ravi Varadhan. 2011. \emph{\emph{Optimx}: A
Replacement and Extension of the {optim()} Function}. Nash Information
Services Inc.; Johns Hopkins University.

\leavevmode\vadjust pre{\hypertarget{ref-nash1996nonlinear}{}}%
Nash, John C., and Paul Velleman. 1996. {``Nonlinear Estimation
Combining Visual Fitting with Optimization Methods.''} In
\emph{Proceedings of the Section on Physical and Engineering Sciences of
the American Statistical Association}, 256--61. American Statistical
Association.

\leavevmode\vadjust pre{\hypertarget{ref-NISTnls-manual}{}}%
National Institutes for Standards, original from, and Technology
http://www.itl.nist.gov/div898/strd/nls/nls\_main.shtml R port by
Douglas Bates. 2012. \emph{NISTnls: Nonlinear Least Squares Examples
from NIST}. \url{https://CRAN.R-project.org/package=NISTnls}.

\leavevmode\vadjust pre{\hypertarget{ref-OlearyRust13}{}}%
O'Leary, Dianne P., and Bert W. Rust. 2013. {``Variable Projection for
Nonlinear Least Squares Problems.''} \emph{Computational Optimization
and Applications} 54 (3): 579--93.

\leavevmode\vadjust pre{\hypertarget{ref-nlsmultstart}{}}%
Padfield, Daniel, and Granville Matheson. 2020. \emph{Nls.multstart:
Robust Non-Linear Regression Using AIC Scores}.
\url{https://CRAN.R-project.org/package=nls.multstart}.

\leavevmode\vadjust pre{\hypertarget{ref-p-nlme}{}}%
Pinheiro, Jose, Douglas Bates, Saikat DebRoy, Deepayan Sarkar, and R
Core Team. 2013. \emph{Nlme: Linear and Nonlinear Mixed Effects Models}.

\leavevmode\vadjust pre{\hypertarget{ref-m-nlsmsn}{}}%
Prates, Marcos, Victor Lachos, and Aldo Garay. 2021. \emph{Nlsmsn:
Fitting Nonlinear Models with Scale Mixture of Skew-Normal
Distributions}. \url{https://CRAN.R-project.org/package=nlsmsn}.

\leavevmode\vadjust pre{\hypertarget{ref-Ratkowsky1983}{}}%
Ratkowsky, David A. 1983. \emph{Nonlinear Regression Modeling: A Unified
Practical Approach}. New York; Basel: Marcel Dekker Inc.

\leavevmode\vadjust pre{\hypertarget{ref-m-nlstac}{}}%
Rodriguez-Arias, Mariano, Juan Antonio Fernandez, Javier Cabello, and
Rafael Benitez. 2020. \emph{Nlstac: An r Package for Fitting Separable
Nonlinear Models}. \url{https://CRAN.R-project.org/package=nlstac}.

\leavevmode\vadjust pre{\hypertarget{ref-Rosenbrock1960}{}}%
Rosenbrock, H. H. 1960. {``An Automatic Method for Finding the Greatest
or Least Value of a Function.''} \emph{The Computer Journal} 3: 175--84.

\leavevmode\vadjust pre{\hypertarget{ref-nlsic-manual}{}}%
Sokol, Serguei. 2022. \emph{Nlsic: Non Linear Least Squares with
Inequality Constraints}. \url{https://CRAN.R-project.org/package=nlsic}.

\leavevmode\vadjust pre{\hypertarget{ref-R-crsnls-manual}{}}%
Tvrdík, Josef. 2016. \emph{Crsnls: Nonlinear Regression Parameters
Estimation by 'CRS4HC' and 'CRS4HCe'}.
\url{https://CRAN.R-project.org/package=crsnls}.

\leavevmode\vadjust pre{\hypertarget{ref-HWtestthat11}{}}%
Wickham, Hadley. 2011. {``{testthat: Get Started with Testing}.''}
\emph{The R Journal} 3: 5--10.
\url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-HWdevtools21}{}}%
Wickham, Hadley, Jim Hester, Winston Chang, and Jennifer Bryan. 2021.
\emph{{devtools: Tools to Make Developing R Packages Easier}}.
\url{https://CRAN.R-project.org/package=devtools}.

\end{CSLReferences}

\end{document}