### 2023-10-18 JNABRev20231018/JNAB2review-response.txt
2nd revision

Response by Nash and Bhattacharjee to second set of referees' reports.

Our responses are prefaced by ">> " on each line.

2-review-2:

I believe the authors have adequately addressed my main comments/questions in the revised version of the paper. I have no other (major) concerns that still need clarification.

>> No response necessary.

2-review-1

Referee report of “The Current State of R Tools for Nonlinear Least Squares Modeling”.

I appreciate the effort that the authors have made in this revision. I believe the paper has
been considerably improved.

>> Thanks. Appreciated.

I think it would useful to expand a bit on some of the comments in the section “An illustrative
example”. In particular, the choice of starting estimates, at least for the third formulation of
the model, doesn’t make sense to me.

>> Added explanatory comment.
>> <!-- 231018: attempt to address comments in 2report1 on modeling vs least squares -->
>> We thank one of the referees for expressing the viewpoint that a nonlinear regression
>> approach would use background knowledge and graphs to obtain an understanding of
>> the general magnitude of the parameters. Indeed, the "visual fitting" approach
>> (@nash1996nonlinear) mentioned later accords with this viewpoint, as does use
>> of the **Logistic3T** variant of the model. 
>> 
>> Our emphasis on the software resilience to starting parameters that are, in the
>> sense of statistical modeling, "silly" comes from over half a century of dealing
>> with users whose interests and understanding are very far from those of statistical
>> modelers. Thus we seek methods and codes that obtain reasonable answers under
>> highly unfavourable conditions. Nevertheless, a proper approach to 
>> nonlinear modeling is to apply all available knowledge to the task.
>> <!-- end 231018 work -->

Purpose of nonlinear least squares fits

In contrast to linear least squares or techniques like generalized additive models (GAMs), which
provide empirical or data-driven models, I feel that the goal of a nonlinear regression analysis,
at least in part, centers on the parameter estimates and the interpretation of the parameters.
That is, the model is usually derived from a mechanistic model of how the data are generated,
this model depends on certain parameters, and the estimates of those parameters are of interest
in and of themselves.

If the parameters are of interest, the analyst should be able to “ball park” the values of the
parameters from the data, probably aided by plotting the data.
I feel there are three stages in a nonlinear least squares fit:

1. Determine starting estimates of the parameters from the data and knowledge of the
interpretation of the parameters.

2. Starting at these estimates use an iterative algorithm that seeks to reduce the sum of
squared residuals at each iteration.

3. Determine if convergence has been achieved at the current iteration and, if not, go to
step 2.

Admittedly, the vast majority of research and software development is focussed on step 2 but
steps 1 and 3 are also important and an analyst performing a nonlinear regression analysis
should be able to formulate starting estimates for a model.

Figure 1: Hobbs weed infestation data with fitted 3-parameter logistic curve (blue) and curve
from the starting estimates in third parameterization of the model (green).
Interpretation of parameters in the three-parameter logistic model
The logistic growth model is based on vertically scaling and possibly shifting the cumulative
distribution function of the logistic distribution. If the lower limit of the predicted response is
set at zero, there is only a vertical scaling constant which determines the asymptotic predicted
response as x → ∞. The three-parameter logistic model incorporates this vertical scaling
parameter with location and scale parameters on the horizontal axis. (The four-parameter
logistic model incorporates both shift and scale parameters on the vertical axis.)
The common way of writing location/scale parameters on the horizontal axis results in the
2location being the value on the x-axis of the inflection point (when the fitted value is midway
between the lower and upper vertical limits) and the scale being the x-axis distance between
the midpoint and the point where the fitted value is roughly one-quarter of the distance from
the lower limit and the upper vertical limits.
It is straightforward to look at a data plot, such as Figure 1, and choose appropriate starting
values for the parameters in this representation of the three-parameter logistic.
However, one does not look at this plot and decide that (1, 1, 1), which would give the green
line in the plot, is a reasonable set of parameter values for this model applied to these data.
As we can see the fitted curve is a very good fit to these data but the curve from the unit
starting estimates is not even close to the observed data. It doesn’t even overlap on the y-axis
with the observed data.

>> We agree the starting points are bad, for the reason given above.

So these are egregiously bad starting estimates for the parameters. The authors state that
fitting this model is very tricky but in my previous report I showed that is was trivial to fit
this model using the self-starting model function SSlogis, because it goes through steps of
using the data to obtain good starting values.

>> Added comment to point out selfStart should be used. 
>> 
>> Users should, of course, be using the `SSlogis` *selfStart* tool, but ignorance
>> of this possibility or a slight variant in the model can easily lead to outcomes
>> similar to those seen here.

A good portion of this paper is studying whether different nonlinear least squares software
implementations can converge to appropriate parameters estimates from these ridiculously
bad starting values and, quite frankly, I don’t see the point.

>> To editor: We are not sure that we can resolve this difference of opinion with the
>>  referee. From a modeling point of view, he/she is definitely correct. From a software
>>  reliability / resilience perspective, we feel it is important that codes keep going
>>  and get decent answers if possible, and we have shown that even for very nasty cases,
>>  software can "work" successfully. Moreover, even if software cannot proceed, it should
>>  provide more guidance if possible than "Singular gradient", which is not meaningful to
>>  most current R users.

It is the author’s prerogative to use whatever criterion for comparison of nonlinear least squares
implementations that they feel is appropriate but I think they should clarify to the reader that
they are using intentionally bad and inappropriate starting estimates as a way of stressing
algorithms.

>> Agreed, and we believe we have now done so. To be honest, the original c(1,1,1) parameters
>> of the 1970s were likely pure laziness and quick to type.

On page 4 the authors state that if a selfStart model is not available nls() sets the starting
values to 1. That may be the case now but the original intention was certainly that a user
calling nls() would provide starting estimates suitable to the model and data.

>> To editor: Referee may be revealing their identity!

Also on page 4 I don’t understand the comment in the hobbs.res function that This variant
uses looping. I don’t see any explicit loops. There will naturally be looping in the C code
that implements the arithmetic operations on vectors but that is inevitable.

>> The comment was a left-over from an earlier variant of the functions and should have been
>> removed. Done. Thanks for spotting this.
>> We also noted that the model is UNscaled, and edited the comment.
>> An explanatory comment has been added
>> 
>> The particular form of these explicit residual and Jacobian functions comes from
>> their translation from BASIC codes of the 1970s, as adapted in @jnmws87. The use
>> of the Laplace form of the residual and the inclusion of data within the functions
>> reflects choices of that era that are at odds with current practice. Some users
>> do still want or need to provide problems as explicit functions, particularly
>> for problems that are NOT regressions. For example, the Rosenbrock banana-valley
>> test problem can be provided this way, where the two "residuals" are different
>> functional forms.

Is it really intended that anyone using the function-based specification of the model will encode
their data into the model function like hobbs.res and hobbs.jac? Generally we would expect
to separate the specification of the data from the model - otherwise there is a great risk of,
say, the residual and the Jacobian function accidentally getting out of sync.

>> In October 2023, one of us (JN) had a query on usage of this formulation where 
>> performance was important. Moreover, the query asked how to extract standard errors
>> and has led to a minor but important addition to the "Intro to nlsr" vignette.
>> Generally we agree that users today are much less likely to use explicit code.

Also, this is the first time that the authors have invoked their non-standard definition of
“residual” as the negative of what most others mean by “residual”, and they haven’t mentioned
this. I would discourage the authors from using an idiosyncratic definition of “residual” simply
to save writing a negative sign in a formula for the Jacobian.

>> We believe our comment above addresses this. We have explained our reasoning for the use
>> of the Laplace residual in other works, but present usage or commentary is a red herring
>> that is irrelevant to today's software and practice. Indeed, the main advance that package
>> 'nlsr' represents is to automatically provide analytic derivatives that caused so much work
>> and introduced so many errors in the 1970s and 1980s.